### 2024-08-07 meetings/MeetingNotes_07082024.Rmd
meeting notes 7 August

---
title: "Meeting Metaviromics 7 August 2024"
author: "laura"
date: "2024-08-07"
output: github_document
---

## Meeting Notes

Attendees: Sophie Gryseels, Magdalini Bletsa, Philippe Selhorst, Jayna Raghwani, Christina Faust, Laura Bergner

Sophie - challenging museum specimens and human samples (FFPE)\
Wildlife trafficking, wild meat sampling\
TE as screening tool

Magda - metagenomics, hepaciviruses\
FFPE - HIV using hybrid capture\
Pathogens carried by migratory birds\
Ancient DNA lab, hybrid capture for Hep B in ancient human samples

Philippe - metagenomic protocols for Nanopore

Less good experience with Twist\
Using depletion instead of enrichment\
Mastomys, positive selection worked less well than depletion, but don't know about what was in the panel or what was done by the company\
SISPA generates short pieces so haven't been using adaptive sampling

Ordering probes family by family\
CoV, ParamyxoV, Flavis\
Tested on known positive samples\
Worked well with Illumina seq\
As you combine probes how does that affect efficiency

RNA viruses mainly of interest\
Balancing depth vs breadth\
38 virus families, most RNA\
Drod vir database & VIRION\
Include viruses from metagenomic surveys as well as some from databases\
460 virus genomes

They are doing ~400 genomes per family\
Flavis started with 8000, after clustering had 800 genomes used to design probes\
Similar for CoV and ParamyxoV

Not sure about combining DNA and RNA viruses\
Better to use nuclease treatment or sensitivity is lost\
Need to split workflow\
Could use DNA transcripts in a RNA workflow but won't get genomes

They are using myBaits, RNA probes\
50-60K probes per family\
Smaller tiling\
Have not combined them\
150K probes in the combined set, unsure how that will affect results

Less genomes more families vs more families less genomes?

Current panels cluster at 90% identity\
Can capture different strains of the same virus (in HIV)

Have only done known positive samples so far\
Real bio controls would be better than something synthetic\
Spike in controls for non rodent pathogen on panel\
They've done spike in control in just pure water\
We are wanting also to quantify variation between samples\
Importance of negative controls for monitoring cross sample contamination

They have also found more viruses with shrews\
They also found a CoV but having difficulty assembling full genome\
They were sequencing from organs but have gotten nearly full genomes from shrew faecal samples in Glasgow\
Combine data from CoVs across teams showing geographical spread

Timeline Glasgow: doing some more metagenomic sequencing\
Start to develop panel in ~December\
First run test in new year\
Test out on samples from Oxford\
Eventually do sequencing in country

Problem with targeting what is already known in rodents, might miss out on surprising/new\
Per family - mammalian all included\
Main question is around host switching so don't want to exclude anything

Could use ancestral reconstruction to look for viruses that don't exist in contemporary dataset, include those in bait design\
Doing wider phylo analysis and ancestral reconstruction to guess at diversity

To make it universal, extra RNA and do RT with overhang instead of random primer. Use template switching RT step then amplify from those overhung sites (similar to S9N)\
Then could design downstream steps including barcodes to be more platform agnostic\
Potential to combine S9N, hybrid barcodes and twist hyb?\
Could potentially be more sensitive

Crispr based negative selection to get rid of host reads\
Can use with template switching\
Can be used to get rid of host RNA\
Developed for Mastomys and humans\
What is the most abundant RNA\
Wanting to make a central repository for Crispr panels so this data can be shared\
Going to make an Aedes/Culex panel\
Could even combined with target enrichment\
Downside is that you need one probe or Cas9 module for each molecule\
Originally used in transcriptomics on cells to deplete rRNA\
More difficult in DNA samples\
Expensive for an initial order but includes a lot\
Order two primers, anneal and elongate\
One ds 200 bp molecule per target

Benchmarking - on the virus level similar but on the depletion level it works better\
Compared to commercial rRNA kits\
More sensitive compared to generic hosts included in commercial panels\
5 euros per sample\
Not yet published

Get away from commercial assays to a more flexible and modular approach\
Discovery funding - apply for something together in the future?

### 2025-01-01 manuscript_figures.Rmd
Finalising manuscript figures and updating figures

---
title: "target_offtarget"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

This is an [R Markdown](http://rmarkdown.rstudio.com) Notebook. When you execute code within the notebook, the results appear beneath the code.

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Cmd+Shift+Enter*.

```{r}
##all viruses - target and off target

library(tidyverse)
library(ggpubr)
library(tidyplots)
library(ggthemes)

```

## Figure 1: plots of TE experiment data - read counts and read depths

### Read-in data
```{r}
#ATCC genomes updated version
#October 2024

#viral load vs read count normalised using different methods, comparing deduplicated and non-deduplicated
#use bowtie2 data

####read counts/normalised read counts####

#metadata

metadata <- read.csv("metadata/sampleIDs_TESpikeIn.csv", header = TRUE)

metadata2 <- metadata |>
  select(Sample.ID, Number.of.read.pairs..quality.adaptor.trimmed.) |>
  rename(Sample_id = Sample.ID) |>
  rename(QC_reads = Number.of.read.pairs..quality.adaptor.trimmed.)

#import and combine read count files
#deduplicated and non-deduplicated

counts_dedup_bt <-
  read.table(
    "data_TE/TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_dedup_atcc_ref.tsv",
    sep = "\t",
    header = TRUE
  )

counts_nodedup_bt <-
  read.table(
    "data_TE/TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv",
    sep = "\t",
    header = TRUE
  )

counts_bt_all <- rbind(counts_dedup_bt, counts_nodedup_bt)

counts_reads <- left_join(counts_bt_all, metadata2, by = "Sample_id")

#normalise by both raw read count and genome length - should be the same as mean read depth

counts_reads_norm <- counts_reads |>
  mutate(norm_counts1 = matched / QC_reads) |>
  mutate(norm_counts2 = matched / QC_reads / length) |>
  mutate(norm_counts3 = matched / length) |>
  mutate(genome_structure = case_when((
    virus == "Human_adenovirus_40" |
      virus == "Human_betaherpesvirus"
  ) ~ "DNA",
  ,
  .default = "RNA"
  ))

```

### Format data
```{r}
#change labels in facet plots

facet_names <- c("dedup_TE" = "Deduplicated",
                 "nodedup_TE" = "Non-Deduplicated")

cols <- c("#4477AA",
          "#66CCEE",
          "#228833",
          "#CCBB44",
          "#EE6677",
          "#AA3377")
```

### Make plots
```{r}

#plot viral read counts (non normalised)

read_count <- 
  counts_reads_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = Viral.load,
    y = log(matched),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12, family = "Helvetica Neue"),
    axis.text.x = element_text(angle = 0, hjust = 1),
    axis.title.y = element_text(size = 12, family = "Helvetica Neue"),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_x_log10() +
  scale_color_manual(
    values = cols,
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  ylab("Log(Viral Reads)")

#ggsave("figures/compare_spike_ins_atcc/read_counts.pdf",width=8,height=6)

#normalise by raw read count

read_count_norm1 <- counts_reads_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = Viral.load,
    y = log(norm_counts1),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12, family = "Helvetica Neue"),
    axis.text.x = element_text(angle = 0, hjust = 1),
    axis.title.y = element_text(size = 12, family = "Helvetica Neue"),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_x_log10() +
  scale_color_manual(
    values = cols,
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  ylab("Log(viral reads/cleaned reads)")

#normalise by raw read count and genome length

read_count_norm2 <- counts_reads_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = Viral.load,
    y = log(norm_counts2),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12, family = "Helvetica Neue"),
    axis.text.x = element_text(angle = 0, hjust = 1),
    axis.title.y = element_text(size = 12, family = "Helvetica Neue"),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_x_log10() +
  scale_color_manual(
    values = cols,
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  ylab("Log(viral reads/cleaned reads/genome length)")

```

### plot just the two lefthand panels from the original figure
```{r, warning=FALSE, echo=FALSE, message=FALSE, fig.height=10, fig.width=8}

ggarrange(read_count,
          read_count_norm2,
          nrow = 2,
          common.legend = TRUE)

#ggsave("figures/compare_spike_ins_atcc/log_read_count_depth_2panels.png",width=10,height=7)

```

### plot viral read counts (non normalised)
```{r}

##deduplicated only

read_count_dedup <- 
  counts_reads_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  filter(type == "dedup_TE") |>
  ggplot(aes(
    x = Viral.load,
    y = log(matched),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12, family = "Helvetica Neue"),
    axis.text.x = element_text(angle = 0, hjust = 1),
    axis.title.y = element_text(size = 12, family = "Helvetica Neue"),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_x_log10() +
  scale_color_manual(
    values = cols,
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  ylab("Log(Viral Reads)")

#normalise by raw read count and genome length

read_count_norm2_dedup <- 
  counts_reads_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  filter(type == "dedup_TE") |>
  ggplot(aes(
    x = Viral.load,
    y = log(norm_counts2),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12, family = "Helvetica Neue"),
    axis.text.x = element_text(angle = 0, hjust = 1),
    axis.title.y = element_text(size = 12, family = "Helvetica Neue"),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_x_log10() +
  scale_color_manual(
    values = cols,
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  ylab("Log(viral reads/cleaned reads/genome length)")

```

### ***Final plot***
```{r, warning=FALSE, message=FALSE, echo=FALSE, fig.height=6, fig.height=8}
#plot just the two lefthand panels from the original figure

ggarrange(
  read_count_dedup,
  read_count_norm2_dedup,
  nrow = 2,
  common.legend = TRUE,
  align = "hv"
)

ggsave("figures/compare_spike_ins_atcc/log_read_count_depth_2panels_dedup.png",
       width=10,
       height=7)

ggsave("figures/manuscript_figures_pdf/Figure1.pdf",
       width=10,
       height=7)

```

## Figure S1

### Read depths
```{r}
####read depths####

#import and combine read depth files

depth_dedup_bt <-
  read_tsv(
    "data_TE/TE_sequencing_experiment_readdepth_per_sample_and_virus_bowtie2_dedup_atcc_ref.tsv"
  )

depth_nodedup_bt <-
  read_tsv(
    "data_TE/TE_sequencing_experiment_readdepth_per_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv"
  )

depths_bt_all <- rbind(depth_dedup_bt, depth_nodedup_bt)

depths_reads <- 
  depths_bt_all |>
  left_join(metadata2, by = "Sample_id") |>
  mutate(genome_structure = case_when((
    virus == "Human_adenovirus_40" |
      virus == "Human_betaherpesvirus"
  ) ~ "DNA",
  .default = "RNA"
  )) |>
  rename(Viral.load = `Viral load`)

```

### Make plots
```{r}

read_depths <- 
  depths_reads |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = Viral.load,
    y = log(mean_depth),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12, family = "Helvetica Neue"),
    axis.text.x = element_text(angle = 0, hjust = 1),
    axis.title.y = element_text(size = 12, family = "Helvetica Neue"),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_color_manual(
    values = cols,
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  scale_x_log10() +
  ylab("Log(mean read depth)")

#ggsave("figures/compare_spike_ins_atcc/mean_depth.pdf",width=8,height=6)

```

### ***Final plot***
```{r}

ggarrange(
  read_count,
  read_count_norm1,
  read_count_norm2,
  read_depths,
  nrow = 2,
  ncol = 2,
  common.legend = TRUE,
  align = "hv"
)

#ggsave("figures/compare_spike_ins_atcc/log_read_count_depth.png",width=10,height=7)

#ggsave("figures/manuscript_figures_pdf/FigureS1.pdf",width=10,height=7)

#apparently normalised read count using method 2 should not actually be the same

```

## Figure S4

### Make plots
```{r}
####read counts/depths split by background####

#change labels in facet plots

facet_names_bg <- c(
  "dedup_TE" = "Deduplicated",
  "nodedup_TE" = "Non-Deduplicated",
  "p2" = "P1",
  "p8" = "P2"
)

#break down by Background x virus

background_counts_reads <- 
  counts_reads_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = Viral.load,
    y = log(matched),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
  facet_grid(Background ~ type, labeller = as_labeller(facet_names_bg)) +
  theme_bw() +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.y = element_text(size = 10),
    legend.title = element_blank()
  ) +
  scale_color_manual(
    values = cols,
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  scale_x_log10() +
  ylab("Log(Viral Reads)")

backgrounds_counts_reads_norm <- 
  counts_reads_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = Viral.load,
    y = log(norm_counts1),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
  facet_grid(Background ~ type, labeller = as_labeller(facet_names_bg)) +
  theme_bw() +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.y = element_text(size = 10),
    legend.title = element_blank()
  ) +
  scale_color_manual(
    values = cols,
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  scale_x_log10() +
  ylab("Log(viral reads/cleaned reads)")

backgrounds_counts_reads_norm2 <- 
  counts_reads_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = Viral.load,
    y = log(norm_counts2),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
  facet_grid(Background ~ type, labeller = as_labeller(facet_names_bg)) +
  theme_bw() +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.y = element_text(size = 10),
    legend.title = element_blank()
  ) +
  scale_color_manual(
    values = cols,
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  scale_x_log10() +
  ylab("Log(viral reads/cleaned reads/genome length)")

background_read_depths <- 
  depths_reads |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = Viral.load,
    y = log(mean_depth),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
  facet_grid(Background ~ type, labeller = as_labeller(facet_names_bg)) +
  theme_bw() +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.y = element_text(size = 10),
    legend.title = element_blank()
  ) +
  scale_color_manual(
    values = cols,
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  scale_x_log10() +
  ylab("Log(mean read depth)")

```

### Final plot 
```{r, warning=FALSE, message=FALSE, echo=FALSE, fig.height=6, fig.height=8}

ggarrange(
  background_counts_reads,
  backgrounds_counts_reads_norm,
  backgrounds_counts_reads_norm2,
  background_read_depths,
  nrow = 2,
  ncol = 2,
  common.legend = TRUE
)

#ggsave("figures/compare_spike_ins_atcc/backgrounds_read_count_depth.png",width=10,height=7)

#ggsave("figures/manuscript_figures_pdf/FigureS4.pdf",width=10,height=7)

```

####extras

```{r}
read_count_raw<-counts_reads_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(x=Viral.load,y=matched,colour = virus))+
  geom_point()+
  geom_smooth(aes(group=virus),se=FALSE)+
  facet_grid(~type,labeller=as_labeller(facet_names))+
  theme(axis.title.x=element_blank(),axis.text.x=element_text(angle=45,hjust=1),axis.title.y=element_text(size=10))+
  scale_x_log10()+
  ylab("Viral Reads")

raw_read_count_norm1<- counts_reads_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(x=Viral.load,y=norm_counts1,colour = virus))+
  geom_point()+
  geom_smooth(aes(group=virus),se=FALSE)+
  facet_grid(~type)+
  theme(axis.title.x=element_blank(),axis.text.x=element_text(angle=45,hjust=1),axis.title.y=element_text(size=10))+
  scale_x_log10()+
  ylab("Viral reads/cleaned reads")

raw_read_count_norm2<- counts_reads_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(x=Viral.load,y=norm_counts2,colour = virus))+
  geom_point()+
  geom_smooth(aes(group=virus),se=FALSE)+
  facet_grid(~type)+
  theme(axis.title.x=element_blank(),axis.text.x=element_text(angle=45,hjust=1),axis.title.y=element_text(size=10))+
  scale_x_log10()+
  ylab("Viral reads/cleaned reads/genome length")

##just genome length

read_count_norm3<- counts_reads_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(x=Viral.load,y=log(norm_counts3),colour = virus))+
  geom_point()+
  geom_smooth(aes(group=virus),se=FALSE)+
  facet_grid(~type,labeller=as_labeller(facet_names))+
  theme_bw()+
  theme(axis.title.x=element_blank(),axis.text.x=element_text(angle=45,hjust=1),axis.title.y=element_text(size=10))+
  scale_x_log10()+
  scale_color_hue(labels=c("Human adenovirus 40","Human betaherpesvirus","Human respiratory syncytial virus","Influenza B virus","Mammalian orthoreovirus 3","Zika virus"))+
  ylab("Log(viral reads/genome length)")

#ggsave("figures/compare_spike_ins_atcc/norm_genome_length.pdf")

raw_read_depths<- depths_reads |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(x=Viral.load,y=mean_depth,colour = virus))+
  geom_point()+
  geom_smooth(aes(group=virus),se=FALSE)+
  facet_grid(~type)+
  theme(axis.title.x=element_blank(),axis.text.x=element_text(angle=45,hjust=1),axis.title.y=element_text(size=10))+
  scale_x_log10()+
  ylab("Mean read depth")

ggarrange(read_count_raw,raw_read_count_norm1,raw_read_count_norm2,raw_read_depths,nrow=2,ncol=2,common.legend=TRUE)

#ggsave("figures/compare_spike_ins_atcc/read_count_depth.pdf")

####split by viruses####

counts_split<-counts_reads_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(x=virus,y=log(matched),colour=type))+
  geom_boxplot()+
  facet_grid(~as.character(Viral.load))+
  ylab("Log(Viral Reads)")+
  theme(axis.title.x=element_blank(),axis.text.x=element_blank(),axis.title.y=element_text(size=10))

counts_norm_split<-counts_reads_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(x=virus,y=log(norm_counts1),colour=type))+
  geom_boxplot()+
  facet_grid(~as.character(Viral.load))+
  ylab("Log(viral reads/cleaned reads)")+
  theme(axis.title.x=element_blank(),axis.text.x=element_blank(),axis.title.y=element_text(size=10))

counts_norm2_split<-counts_reads_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(x=virus,y=log(norm_counts2),colour=type))+
  geom_boxplot()+
  facet_grid(~as.character(Viral.load))+
  ylab("Log(viral reads/cleaned reads/genome length)")+
  theme(axis.title.x=element_blank(),axis.text.x=element_text(angle=45,hjust=1),axis.title.y=element_text(size=10))

depth_split<-depths_pool |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(x=virus,y=log(mean_depth),colour=type))+
  geom_boxplot()+
  facet_grid(~as.character(Viral.load))+
  theme(axis.title.x=element_blank(),axis.text.x=element_text(angle=45,hjust=1),axis.title.y=element_text(size=10))+
  ylab("Log(mean read depth)")+
  theme(axis.title.x=element_blank(),axis.text.x=element_text(angle=45,hjust=1),axis.title.y=element_text(size=10))

ggarrange(counts_split,counts_norm_split,counts_norm2_split,depth_split,nrow=2,ncol=2,heights=c(1,2),common.legend = TRUE)

#ggsave("figures/compare_spike_ins_atcc/compare_viruses_split.pdf")

```

## >>> Figure 2: plots of TE experiment data - genome coverage & per site coverage

### Read-in data
```{r}

#plots of TE experiment data - genome coverage & per site coverage
#also separated by viruses
#ATCC genomes updated version
#October 2024

#use bowtie2 data

####genome coverage####

dedup_per_site <-
  read_tsv(
    "data_TE/TE_sequencing_experiment_readdepth_per_site

    TEdata_TE/TE_sequencing_experiment_readdepth_per_site_sample_and_virus_bowtie2_dedup_atcc_ref.tsv"
  )

nodedup_per_site <-
  read_tsv(
    "data_TE/TE_sequencing_experiment_readdepth_per_site_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv"
  )

#add length column from read depth file to per site file

#metadata

metadata <-
  read.csv(
    "metadata/sampleIDs_TESpikeIn.csv",
    header = TRUE
  )

metadata2 <- metadata |>
  select(Sample.ID, Number.of.read.pairs..quality.adaptor.trimmed.) |>
  rename(Sample_id = Sample.ID) |>
  rename(QC_reads = Number.of.read.pairs..quality.adaptor.trimmed.)

#import and combine read count files
#deduplicated and non-deduplicated

counts_dedup_bt <-
  read.table(
    "data_TE/TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_dedup_atcc_ref.tsv",
    sep = "\t",
    header = TRUE
  )

counts_nodedup_bt <-
  read.table(
    "data_TE/TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv",
    sep = "\t",
    header = TRUE
  )

counts_bt_all <- rbind(counts_dedup_bt, counts_nodedup_bt)

counts_reads <- left_join(counts_bt_all, metadata2, by = "Sample_id")

```

### Format data
```{r}
#normalise by both raw read count and genome length - should be the same as mean read depth

counts_reads_norm<-counts_reads |>
  mutate(norm_counts1 = matched/QC_reads) |>
  mutate(norm_counts2 = matched/QC_reads/length) |>
  mutate(norm_counts3 = matched/length) |>
  mutate(genome_structure = case_when((virus == "Human_adenovirus_40"| virus == "Human_betaherpesvirus") ~ "DNA",
                                      ,.default = "RNA"))

lengths<-counts_reads_norm |>
  select(virus,length) |>
  distinct()

#calculate genome coverage

persite_coverage_dedup<-dedup_per_site |>
  group_by(Sample_id,virus,Viral.load,Background,type) |>
  filter(coverage > 0) |>
  summarise(genome_coverage = n()) |>
  left_join(lengths) |>
  mutate(percent_coverage = genome_coverage/length)

persite_coverage_nodedup<-nodedup_per_site |>
  group_by(Sample_id,virus,Viral.load,Background,type) |>
  filter(coverage > 0) |>
  summarise(genome_coverage = n()) |>
  left_join(lengths) |>
  mutate(percent_coverage = genome_coverage/length)

persite_coverage_both<-rbind(persite_coverage_dedup,persite_coverage_nodedup)

coverage_labels<-c("0" = "Control",
                   "100" = "1e+02",
                   "1000" = "1e+03",
                   "100000" = "1e+05",
                   "dedup_TE" = "Deduplicated",
                   "nodedup_TE"= "Non-Deduplicated")

coverage_labels2<-c("0" = "Control",
                    "100" = "1e+02",
                    "1000" = "1e+03",
                    "100000" = "1e+05")

cols2<-c("#BB5566","#004488")

cols4<-c("#DDAA33","#BB5566","#004488")

persite_coverage_both |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(x=virus,y=percent_coverage,colour = Background))+
  geom_boxplot(outlier.shape=NA)+
  geom_point(position = position_dodge(width = .75))+
  facet_grid(type~as.character(Viral.load),labeller=as_labeller(coverage_labels))+
  theme_bw()+
  theme(axis.title.y=element_text(size=10))+
  ylab("Genome coverage")+
  theme(axis.title.x=element_blank(),axis.text.x=element_text(angle=45,hjust=1),axis.title.y=element_text(size=15),axis.text=element_text(size=12),strip.text=element_text(size=12))+
  scale_color_manual(values=cols2,labels=c("ME_P1","ME_P2"))+
  scale_x_discrete(labels=c("Human adenovirus 40","Human betaherpesvirus","Human respiratory syncytial virus","Influenza B virus","Mammalian orthoreovirus 3","Zika virus"))+
  scale_y_continuous(limits=c(0,1))

#ggsave("figures/compare_spike_ins_atcc/genome_cov_split.png",width=15,height=6)

##are the deduplicated and non deduplicated datasets identical??
##if they are the same can maybe just show one
#show deduplicated

persite_coverage_both |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  filter(type == "dedup_TE") |>
  ggplot(aes(x=virus,y=percent_coverage,colour = Background))+
  geom_boxplot(outlier.shape=NA)+
  geom_point(position = position_dodge(width = .75))+
  facet_grid(~as.character(Viral.load),labeller=as_labeller(coverage_labels2))+
  theme_bw()+
  ylab("Genome coverage")+
  theme(axis.title.x=element_blank(),axis.text.x=element_text(angle=45,hjust=1),axis.title.y=element_text(size=15),axis.text=element_text(size=12),strip.text=element_text(size=12),legend.text = element_text(size=10))+
  scale_color_manual(values=cols2,labels=c("ME_P1","ME_P2"))+
  scale_x_discrete(labels=c("Human adenovirus 40","Human betaherpesvirus","Human respiratory syncytial virus","Influenza B virus","Mammalian orthoreovirus 3","Zika virus"))+
  scale_y_continuous(limits=c(0,1))

#ggsave("figures/compare_spike_ins_atcc/genome_cov_deduponly.png",width=15,height=6)

#ggsave("figures/manuscript_figures_pdf/Figure2.pdf",width=15,height=6)

ggplot(persite_coverage_both,aes(x=as.character(Viral.load),y=percent_coverage))+
  geom_point()+
  geom_boxplot(outlier.shape=NA)+
  facet_wrap(~type)+
  ylab("Genome coverage")+
  xlab("Viral load")

#ggsave("figures/compare_spike_ins_atcc/genome_cov_combined.pdf")

persite_both<-rbind(dedup_per_site,nodedup_per_site)

persite_reads<-left_join(persite_both,metadata2,by="Sample_id")

persite_norm<-persite_reads |>
  mutate(norm_cov = coverage/QC_reads)

coverage_labels3<-c("100" = "1e+02",
                    "1000" = "1e+03",
                    "100000" = "1e+05",
                    "p2" = "ME_P1",
                    "p8"= "ME_P2")

persite_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  filter(type == "dedup_TE") |>
  filter(virus == "Human_respiratory_syncytial_virus") |>
  ggplot(aes(x=site,y=coverage,fill=Background))+
  geom_col()+
  facet_wrap(Viral.load~Sample_id)+
  ylab("Log(Coverage)")+
  ggtitle("Human RSV (deduplicated)")+
  guides(fill=guide_legend(title="Background"))+
  theme_bw()+
  scale_y_log10()+
  scale_fill_manual(values=cols2,labels=c("ME_P1","ME_P2"))

#ggsave("figures/manuscript_figures_pdf/FigureS5.pdf",width=10,height=8)

persite_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  filter(type == "dedup_TE") |>
  filter(virus == "Zika_virus") |>
  ggplot(aes(x=site,y=coverage,fill=Background))+
  geom_col()+
  facet_wrap(Viral.load~Sample_id)+
  ylab("Log(Coverage)")+
  ggtitle("Zika virus (deduplicated)")+
  guides(fill=guide_legend(title="Background"))+
  theme_bw()+
  scale_y_log10()+
  scale_fill_manual(values=cols2,labels=c("ME_P1","ME_P2"))

#ggsave("figures/manuscript_figures_pdf/FigureS6.pdf",width=10,height=8)

persite_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  filter(type == "dedup_TE") |>
  filter(virus == "Human_adenovirus_40") |>
  ggplot(aes(x=site,y=coverage,fill=Background))+
  geom_col()+
  facet_wrap(Viral.load~Sample_id)+
  ylab("Log(Coverage)")+
  ggtitle("Human Adenovirus (deduplicated)")+
  guides(fill=guide_legend(title="Background"))+
  theme_bw()+
  scale_y_log10()+
  scale_fill_manual(values=cols2,labels=c("ME_P1","ME_P2"))

#ggsave("figures/manuscript_figures_pdf/FigureS7.pdf",width=20,height=18)

persite_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  filter(type == "dedup_TE") |>
  filter(virus == "Human_betaherpesvirus") |>
  ggplot(aes(x=site,y=coverage,fill=Background))+
  geom_polygon()+
  facet_wrap(Viral.load~Sample_id)+
  ylab("Log(Coverage)")+
  ggtitle("Human Betaherpesvirus (deduplicated)")+
  guides(fill=guide_legend(title="Background"))+
  theme_bw()+
  scale_y_log10()+
  scale_fill_manual(values=cols2,labels=c("ME_P1","ME_P2"))

#ggsave("figures/compare_spike_ins_atcc/HBHV_coverage_test.pdf",width=20,height=18)

#have to do these differently due to segmentation

FLU_ME1<-persite_norm |>
  filter(virus == "Influenza_B_virus") |>
  filter(type == "dedup_TE") |>
  filter(Background == "p2") |>
  ggplot(aes(x=site,y=coverage,fill=as.character(Viral.load)))+
  geom_col()+
  facet_grid(seg~Sample_id)+
  ylab("Log(Coverage)")+
  ggtitle("Influenza B virus (Background ME_1 deduplicated)")+
  guides(fill=guide_legend(title="Viral load"))+
  theme_bw()+
  scale_y_log10()+
  scale_fill_manual(values=cols4,labels=c("1e+02","1e+03","1e+05"))

FLU_ME2<-persite_norm |>
  filter(virus == "Influenza_B_virus") |>
  filter(type == "dedup_TE") |>
  filter(Background == "p8") |>
  ggplot(aes(x=site,y=coverage,fill=as.character(Viral.load)))+
  geom_col()+
  facet_grid(seg~Sample_id)+
  ylab("Log(Coverage)")+
  ggtitle("Influenza B virus (Background ME_2 deduplicated)")+
  guides(fill=guide_legend(title="Viral load"))+
  theme_bw()+
  scale_y_log10()+
  scale_fill_manual(values=cols4,labels=c("1e+02","1e+03","1e+05"))

ggarrange(FLU_ME1,FLU_ME2,ncol=2,common.legend=TRUE)

#ggsave("figures/manuscript_figures_pdf/FigureS9.pdf",width=20,height=18)

REO_ME1<-persite_norm |>
  filter(virus == "Mammalian_orthoreovirus3") |>
  filter(type == "dedup_TE") |>
  filter(Background == "p2") |>
  ggplot(aes(x=site,y=coverage,fill=as.character(Viral.load)))+
  geom_col()+
  facet_grid(seg~Sample_id)+
  ylab("Log(Coverage)")+
  ggtitle("Mammalian orthoreovirus 3 (Background ME_1 deduplicated)")+
  guides(fill=guide_legend(title="Viral load"))+
  theme_bw()+
  scale_y_log10()+
  scale_fill_manual(values=cols4,labels=c("1e+02","1e+03","1e+05"))

REO_ME2<-persite_norm |>
  filter(virus == "Mammalian_orthoreovirus3") |>
  filter(type == "dedup_TE") |>
  filter(Background == "p8") |>
  ggplot(aes(x=site,y=coverage,fill=as.character(Viral.load)))+
  geom_col()+
  facet_grid(seg~Sample_id)+
  ylab("Log(Coverage)")+
  ggtitle("Mammalian orthoreovirus 3 (Background ME_2 deduplicated)")+
  guides(fill=guide_legend(title="Viral load"))+
  theme_bw()+
  scale_y_log10()+
  scale_fill_manual(values=cols4,labels=c("1e+02","1e+03","1e+05"))

ggarrange(REO_ME1,REO_ME2,ncol=2,common.legend=TRUE)

#ggsave("figures/manuscript_figures_pdf/FigureS10.pdf",width=20,height=18)

```

## Figure 3: all viruses - target and off target

### Read-in data
```{r}

# read in metadata

metadata <-
  read.csv("metadata/sampleIDs_TESpikeIn.csv", header = TRUE)

metadata2 <-
  metadata |>
  select(
    Sample.ID,
    Background.sample,
    Viral.load,
    Number.of.read.pairs..quality.adaptor.trimmed.
  ) |>
  rename(Sample_id = Sample.ID) |>
  rename(QC_reads = Number.of.read.pairs..quality.adaptor.trimmed.)

#import viral reads mapped (to calculate proportions)

viral_reads_dedup <-
  read.csv(
    "data_TE/total_virus_mapped_reads_per_sample_dedup_atcc_ref_20241108.csv",
    header = TRUE
  )

viral_reads_nodedup <-
  read.csv(
    "data_TE/total_virus_mapped_reads_per_sample_nodedup_atcc_ref_20241108.csv",
    header = TRUE
  )

#import contingency tables

contingency_dedup <-
  read.csv(
    "data_TE/contingency_table_mapped_virus_reads_per_family_genus_sample_dedup_atcc_ref_20241106.csv",
    header = TRUE
  )

contingency_nodedup <-
  read.csv(
    "data_TE/contingency_table_mapped_virus_reads_per_family_genus_sample_nodedup_atcc_ref_20241106.csv",
    header = TRUE
  )

contaminants <-
  c("Betacoronavirus", "Alphainfluenzavirus", "Gammaretrovirus")

```

### Format data
```{r}

# pivot longer  
dedup_long <-
  contingency_dedup |>
  filter(!genus %in% contaminants) |>
  filter(genus != "NA") |>
  pivot_longer(cols = A:P,
               names_to = "Sample_id",
               values_to = "count") |>
  mutate(
    target = case_when(
      genus == "Cyclovirus" ~ "off_target",
      genus == "Cardiovirus" ~ "off_target",
      genus == "Kobuvirus" ~ "off_target",
      .default = "on_target"
    )
  ) |>
  mutate(genome_structure = case_when((genus == "Mastadenovirus" |
                                         genus == "Cytomegalovirus") ~ "DNA",
                                      .default = "RNA"
  ))

no_dedup_long <- contingency_nodedup |>
  filter(!genus %in% contaminants) |>
  filter(genus != "NA") |>
  pivot_longer(cols = A:P,
               names_to = "Sample_id",
               values_to = "count") |>
  mutate(
    target = case_when(
      genus == "Cyclovirus" ~ "off_target",
      genus == "Cardiovirus" ~ "off_target",
      genus == "Kobuvirus" ~ "off_target",
      .default = "on_target"
    )
  ) |>
  mutate(genome_structure = case_when((genus == "Mastadenovirus" |
                                         genus == "Cytomegalovirus") ~ "DNA",
                                      .default = "RNA"
  ))

metadata_dedup <- left_join(dedup_long, metadata2, by = "Sample_id")

dedup_viral_reads <-
  left_join(metadata_dedup, viral_reads_dedup, by = "Sample_id") |>
  mutate(total_reads = QC_reads * 2) |>
  mutate(prop_total = count / total_reads) |>
  mutate(prop_viral = count / total_virus_reads)

metadata_no_dedup <- left_join(no_dedup_long, metadata2, by = "Sample_id")

nodedup_viral_reads <-
  left_join(metadata_no_dedup, viral_reads_nodedup, by = "Sample_id") |>
  mutate(total_reads = QC_reads * 2) |>
  mutate(prop_total = count / total_reads) |>
  mutate(prop_viral = count / total_virus_reads)

```

### Make plots
```{r, warning=FALSE, message=FALSE, echo=FALSE}

facet_names <- c(
  "Msp_p2" = "P1",
  "Msp_p8" = "P2",
  "off_target" = "Off target viruses",
  "on_target" = "On target viruses"
)

cols <-
  c(
    "#CCBB44",
    "#332288",
    "#EE7733",
    "#66CCEE",
    "#882255",
    "#4477AA",
    "#AA3377",
    "#228833",
    "#EE6677"
  )

dedup_reads <- 
  dedup_viral_reads |>
  filter(Background.sample != "Neg_control") |>
  filter(Background.sample != "Msp_p6") |>
  ggplot(aes(
    x = Viral.load,
    y = log(count),
    colour = genus
  )) +
  geom_point() +
  geom_smooth(aes(group = genus, linetype = genome_structure), se = FALSE) +
  facet_grid(target ~ Background.sample, labeller = as_labeller(facet_names)) +
  theme_few() + 
  scale_x_log10() +
  ylab("Log (Viral Reads)") + 
  xlab("Log (Viral load)") + 
  scale_color_manual(values = cols) +
  theme(
    axis.title.x = element_text(size = 12, family = "Helvetica Neue"),
    axis.text.x = element_text(angle = 30, hjust = 1),
    axis.title.y = element_text(size = 12, family = "Helvetica Neue"),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  )

dedup_prop_viral <-
  dedup_viral_reads |>
  filter(Background.sample != "Neg_control") |>
  filter(Background.sample != "Msp_p6") |>
  ggplot(aes(x = Viral.load, y = prop_viral, colour = genus)) +
  geom_point() +
  geom_smooth(aes(group = genus, linetype = genome_structure), se = FALSE) +
  facet_grid(target ~ Background.sample, labeller = as_labeller(facet_names)) +
  theme_few() + 
  scale_x_log10() +
  scale_y_log10() +
  ylab("Proportion of viral reads") +
  xlab("Log (Viral load)") + 
  scale_color_manual(values = cols) +
    theme(
    axis.title.x = element_text(size = 12, family = "Helvetica Neue"),
    axis.text.x = element_text(angle = 30, hjust = 1),
    axis.title.y = element_text(size = 12, family = "Helvetica Neue"),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  )

dedup_prop_all <-
  dedup_viral_reads |>
  filter(Background.sample != "Neg_control") |>
  filter(Background.sample != "Msp_p6") |>
  ggplot(aes(x = Viral.load, y = prop_total, colour = genus)) +
  geom_point() +
  geom_smooth(aes(group = genus, linetype = genome_structure), se = FALSE) +
  facet_grid(target ~ Background.sample, labeller = as_labeller(facet_names)) +
  theme_few() + 
  scale_x_log10() +
  scale_y_log10(limits=c(5e-8,1)) +
  ylab("Proportion of total reads") +
  xlab("Log (Viral load)") + 
  scale_color_manual(values = cols) +
    theme(
    axis.title.x = element_text(size = 12, family = "Helvetica Neue"),
    axis.text.x = element_text(angle = 30, hjust = 1),
    axis.title.y = element_text(size = 12, family = "Helvetica Neue"),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  )

```

### Dedup -- viral reads, prop all reads, and prop viral reads
```{r, warning=FALSE, message=FALSE, echo=FALSE, fig.width=8, fig.height=12}

ggarrange(
  dedup_reads,
  dedup_prop_all,
  dedup_prop_viral,
  nrow = 3,
  common.legend = TRUE,
  align = "hv"
)

```

### ***Final plot***

```{r, warning=FALSE, message=FALSE, echo=FALSE, fig.width=7, fig.height=8}

#ggsave("figures/compare_spike_ins_atcc/target_offtarget_dedup.png",width=8,height=10)

##just plot viral reads and prop viral

ggarrange(
  dedup_reads,
  dedup_prop_viral,
  nrow = 2,
  common.legend = TRUE,
  align = "hv"
)

ggsave("figures/compare_spike_ins_atcc/target_offtarget_dedup_2025-01-01.png", width = 10, height = 8)

ggsave("figures/manuscript_figures_pdf/Figure_3_2025-01-01.pdf", width = 10, height = 8)

```

### ----nodedup results ----
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=12, fig.width=6}
##extras - no dedup

nodedup_reads <-
  nodedup_viral_reads |>
  filter(Background.sample != "Neg_control") |>
  filter(Background.sample != "Msp_p6") |>
  ggplot(aes(
    x = Viral.load,
    y = log(count),
    colour = genus
  )) +
  geom_point() +
  geom_smooth(aes(group = genus, linetype = genome_structure), se = FALSE) +
  facet_grid(target ~ Background.sample, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12, family = "Helvetica Neue"),
    axis.text.x = element_text(angle = 0, hjust = 1),
    axis.title.y = element_text(size = 12, family = "Helvetica Neue"),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_x_log10() +
  ylab("Log(Viral Reads)") +
  scale_color_manual(values = cols)

nodedup_prop_viral <- 
  nodedup_viral_reads |>
  filter(Background.sample != "Neg_control") |>
  filter(Background.sample != "Msp_p6") |>
  ggplot(aes(x = Viral.load, y = prop_viral, colour = genus)) +
  geom_point() +
  geom_smooth(aes(group = genus, linetype = genome_structure), se = FALSE) +
  facet_grid(target ~ Background.sample, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12, family = "Helvetica Neue"),
    axis.text.x = element_text(angle = 0, hjust = 1),
    axis.title.y = element_text(size = 12, family = "Helvetica Neue"),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_x_log10() +
  scale_y_log10() +
  ylab("Proportion of viral reads") +
  scale_color_manual(values = cols)

nodedup_prop_all <- nodedup_viral_reads |>
  filter(Background.sample != "Neg_control") |>
  filter(Background.sample != "Msp_p6") |>
  ggplot(aes(x = Viral.load, y = prop_total, colour = genus)) +
  geom_point() +
  geom_smooth(aes(group = genus, linetype = genome_structure), se = FALSE) +
  facet_grid(target ~ Background.sample, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12, family = "Helvetica Neue"),
    axis.text.x = element_text(angle = 0, hjust = 1),
    axis.title.y = element_text(size = 12, family = "Helvetica Neue"),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_x_log10() +
  scale_y_log10() +
  ylab("Proportion of total reads") +
  scale_color_manual(values = cols)

ggarrange(
  nodedup_reads,
  nodedup_prop_all,
  nodedup_prop_viral,
  nrow = 3,
  common.legend = TRUE,
  align = "hv"
)

#ggsave("figures/compare_spike_ins_atcc/target_offtarget_nodedup.pdf",width=8,height=10)

```

### 2025-01-02 manuscript_figures.Rmd
add notebook for manuscript figures only

@@
 
-## Figure 1: plots of TE experiment data - read counts and read depths
+## >>>> Figure 1: plots of TE experiment data - read counts and read depths
 
@@
   theme(
-    axis.title.x = element_text(size = 12, family = "Helvetica Neue"),
+    axis.title.x = element_text(size = 12),
     axis.text.x = element_text(angle = 0, hjust = 1),
-    axis.title.y = element_text(size = 12, family = "Helvetica Neue"),
+    axis.title.y = element_text(size = 12),
     legend.title = element_blank(),
@@
 
+### --------------------
+
 ## Figure S1
@@
 
+### --------------------
+
+## Figure S2
+```{r}
+#metadata
+
+metadata <-
+  read.csv(
+    "/Users/laura/Dropbox/glasgow/github/te_ug_rodents/metadata/sampleIDs_TESpikeIn.csv",
+    header = TRUE
+  )
+
+metadata2 <- metadata %>%
+  select(Sample.ID, Number.of.read.pairs..quality.adaptor.trimmed.) %>%
+  rename(Sample_id = Sample.ID) %>%
+  rename(QC_reads = Number.of.read.pairs..quality.adaptor.trimmed.)
+
+#import and combine read count files
+#deduplicated and non-deduplicated
+
+counts_dedup_bt <-
+  read.table(
+    "/Users/laura/Dropbox/glasgow/github/te_ug_rodents/data_TE/TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_dedup_atcc_ref.tsv",
+    sep = "\t",
+    header = TRUE
+  )
+
+counts_nodedup_bt <-
+  read.table(
+    "/Users/laura/Dropbox/glasgow/github/te_ug_rodents/data_TE/TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv",
+    sep = "\t",
+    header = TRUE
+  )
+
+counts_bt_all <- rbind(counts_dedup_bt, counts_nodedup_bt)
+
+counts_reads <- left_join(counts_bt_all, metadata2, by = "Sample_id")
+
+#normalise by both raw read count and genome length - should be the same as mean read depth
+
+counts_reads_norm <- counts_reads %>%
+  mutate(norm_counts1 = matched / QC_reads) %>%
+  mutate(norm_counts2 = matched / QC_reads / length) %>%
+  mutate(norm_counts3 = matched / length) %>%
+  mutate(genome_structure = case_when((
+    virus == "Human_adenovirus_40" |
+      virus == "Human_betaherpesvirus"
+  ) ~ "DNA",
+  ,
+  .default = "RNA"
+  ))
+
+####compare library capture pool####
+
+cols3 <- c("#228833", "#AA3377")
+
+facet_names <- c("dedup_TE" = "Deduplicated",
+                 "nodedup_TE" = "Non-Deduplicated")
+
+metadata3 <- metadata %>%
+  select(Sample.ID, Pool.for.sequencing) %>%
+  rename(Sample_id = Sample.ID) %>%
+  rename(Pool = Pool.for.sequencing)
+
+counts_pool <- left_join(counts_reads_norm, metadata3, by = "Sample_id")
+
+pool_readcounts <- counts_pool %>%
+  filter(Background != "p6") %>%
+  filter(Background != "control") %>%
+  ggplot(aes(x = virus, y = log(matched))) +
+  geom_boxplot() +
+  facet_grid(Pool ~ type) +
+  theme_bw() +
+  theme(
+    axis.title.x = element_blank(),
+    axis.text.x = element_blank(),
+    axis.title.y = element_text(size = 10)
+  ) +
+  ylab("Log(Viral Reads)") +
+  scale_color_manual(values = cols3, labels = c("Pool 1", "Pool 2"))
+
+pool_counts_sum <- counts_pool %>%
+  filter(Background != "p6") %>%
+  filter(Background != "control") %>%
+  ggplot(aes(
+    x = as.character(Viral.load),
+    y = log(matched),
+    colour = Pool
+  )) +
+  geom_boxplot() +
+  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
+  theme_bw() +
+  theme(axis.title.x = element_blank(), axis.title.y = element_text(size =
+                                                                      10)) +
+  ylab("Log(Viral Reads)") +
+  scale_color_manual(values = cols3, labels = c("Pool 1", "Pool 2"))
+
+pool_readcounts_norm <- counts_pool %>%
+  filter(Background != "p6") %>%
+  filter(Background != "control") %>%
+  ggplot(aes(x = virus, y = log(norm_counts1))) +
+  geom_boxplot() +
+  facet_grid(Pool ~ type) +
+  theme(
+    axis.title.x = element_blank(),
+    axis.text.x = element_blank(),
+    axis.title.y = element_text(size = 10)
+  ) +
+  ylab("Log(viral reads/cleaned reads)")
+
+pool_norm1_sum <- counts_pool %>%
+  filter(Background != "p6") %>%
+  filter(Background != "control") %>%
+  ggplot(aes(
+    x = as.character(Viral.load),
+    y = log(matched),
+    colour = Pool
+  )) +
+  geom_boxplot() +
+  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
+  theme_bw() +
+  theme(axis.title.x = element_blank(), axis.title.y = element_text(size =
+                                                                      10)) +
+  ylab("Log(viral reads/cleaned reads)") +
+  scale_color_manual(values = cols3, labels = c("Pool 1", "Pool 2"))
+
+pool_readcounts_norm2 <- counts_pool %>%
+  filter(Background != "p6") %>%
+  filter(Background != "control") %>%
+  ggplot(aes(x = virus, y = log(norm_counts2))) +
+  geom_boxplot() +
+  facet_grid(Pool ~ type) +
+  theme(
+    axis.title.x = element_blank(),
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    axis.title.y = element_text(size = 10)
+  ) +
+  ylab("Log(viral reads/cleaned reads/genome length)")
+
+pool_norm2_sum <- counts_pool %>%
+  filter(Background != "p6") %>%
+  filter(Background != "control") %>%
+  ggplot(aes(
+    x = as.character(Viral.load),
+    y = log(matched),
+    colour = Pool
+  )) +
+  geom_boxplot() +
+  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
+  theme_bw() +
+  theme(axis.title.x = element_blank(), axis.title.y = element_text(size =
+                                                                      10)) +
+  ylab("Log(viral reads/cleaned reads/genome length)") +
+  scale_color_manual(values = cols3, labels = c("Pool 1", "Pool 2"))
+
+#import and combine read depth files
+
+depth_dedup_bt <-
+  read.table(
+    "/Users/laura/Dropbox/glasgow/github/te_ug_rodents/data_TE/TE_sequencing_experiment_readdepth_per_sample_and_virus_bowtie2_dedup_atcc_ref.tsv",
+    sep = "\t",
+    header = TRUE
+  )
+
+depth_nodedup_bt <-
+  read.table(
+    "/Users/laura/Dropbox/glasgow/github/te_ug_rodents/data_TE/TE_sequencing_experiment_readdepth_per_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv",
+    sep = "\t",
+    header = TRUE
+  )
+
+depths_bt_all <- rbind(depth_dedup_bt, depth_nodedup_bt)
+
+depths_reads <- left_join(depths_bt_all, metadata2, by = "Sample_id") %>%
+  mutate(genome_structure = case_when((
+    virus == "Human_adenovirus_40" |
+      virus == "Human_betaherpesvirus"
+  ) ~ "DNA",
+  ,
+  .default = "RNA"
+  ))
+
+depths_pool <- left_join(depths_reads, metadata3, by = "Sample_id")
+
+pool_readdepths <- depths_pool %>%
+  filter(Background != "p6") %>%
+  filter(Background != "control") %>%
+  ggplot(aes(x = virus, y = log(mean_depth))) +
+  geom_boxplot() +
+  facet_grid(Pool ~ type) +
+  theme(
+    axis.title.x = element_blank(),
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    axis.title.y = element_text(size = 10)
+  ) +
+  ylab("Log(mean read depth)")
+
+pool_depths_sum <- depths_pool %>%
+  filter(Background != "p6") %>%
+  filter(Background != "control") %>%
+  ggplot(aes(
+    x = as.character(Viral.load),
+    y = log(mean_depth),
+    colour = Pool
+  )) +
+  geom_boxplot() +
+  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
+  theme_bw() +
+  theme(axis.title.x = element_blank(), axis.title.y = element_text(size =
+                                                                      10)) +
+  ylab("Log(mean read depth)") +
+  scale_color_manual(values = cols3, labels = c("Pool 1", "Pool 2"))
+
+ggarrange(
+  pool_readcounts,
+  pool_readcounts_norm,
+  pool_readcounts_norm2,
+  pool_readdepths,
+  nrow = 2,
+  ncol = 2,
+  heights = c(1, 2),
+  common.legend = TRUE
+)
+
+#ggsave("/Users/laura/Dropbox/glasgow/github/te_ug_rodents/figures/compare_spike_ins_atcc/pool_compare_viruses.pdf")
+
+ggarrange(
+  pool_counts_sum,
+  pool_norm1_sum,
+  pool_norm2_sum,
+  pool_depths_sum,
+  nrow = 2,
+  ncol = 2,
+  common.legend = TRUE
+)
+
+#ggsave("/Users/laura/Dropbox/glasgow/github/te_ug_rodents/figures/compare_spike_ins_atcc/pool_compare_viruses_sum.png",width=10,height=7)
+
+#ggsave("/Users/laura/Dropbox/glasgow/github/te_ug_rodents/figures/manuscript_figures_pdf/FigureS2.pdf",width=10,height=7)
+
+```
+
+### --------------------
+
+## Figure S3
+```{r}
+
+##plots of read counts and viral read counts
+#November 2024
+
+rm(list = ls())
+
+library(tidyverse)
+library(ggpubr)
+
+#metadata
+
+metadata <-
+  read.csv(
+    "/Users/laura/Dropbox/glasgow/github/te_ug_rodents/metadata/sampleIDs_TESpikeIn.csv",
+    header = TRUE
+  )
+
+metadata2 <- metadata %>%
+  select(Sample.ID, Number.of.read.pairs..quality.adaptor.trimmed.) %>%
+  rename(Sample_id = Sample.ID) %>%
+  rename(QC_reads = Number.of.read.pairs..quality.adaptor.trimmed.)
+
+#import read count files
+
+counts_dedup_bt <-
+  read.table(
+    "/Users/laura/Dropbox/glasgow/github/te_ug_rodents/data_TE/TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_dedup_atcc_ref.tsv",
+    sep = "\t",
+    header = TRUE
+  )
+
+counts_nodedup_bt <-
+  read.table(
+    "/Users/laura/Dropbox/glasgow/github/te_ug_rodents/data_TE/TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv",
+    sep = "\t",
+    header = TRUE
+  )
+
+#import viral reads mapped (to calculate proportions)
+
+viral_reads_dedup <-
+  read.csv(
+    "/Users/laura/Dropbox/glasgow/github/te_ug_rodents/data_TE/total_virus_mapped_reads_per_sample_dedup_atcc_ref_20241108.csv",
+    header = TRUE
+  )
+
+viral_reads_nodedup <-
+  read.csv(
+    "/Users/laura/Dropbox/glasgow/github/te_ug_rodents/data_TE/total_virus_mapped_reads_per_sample_nodedup_atcc_ref_20241108.csv",
+    header = TRUE
+  )
+
+cols2 <- c("#BB5566", "#004488")
+
+facet_names <- c("dedup_TE" = "Deduplicated",
+                 "nodedup_TE" = "Non-Deduplicated")
+
+reads_metadata_dedup <-
+  left_join(counts_dedup_bt, metadata2, by = "Sample_id")
+
+reads_viral_dedup <-
+  left_join(reads_metadata_dedup, viral_reads_dedup, by = "Sample_id")
+
+reads_metadata_nodedup <-
+  left_join(counts_nodedup_bt, metadata2, by = "Sample_id")
+
+reads_viral_nodedup <-
+  left_join(reads_metadata_nodedup, viral_reads_nodedup, by = "Sample_id")
+
+reads_viral_all <- rbind(reads_viral_dedup, reads_viral_nodedup)
+
+reads_plot <- reads_viral_all %>%
+  group_by(Background, Sample_id, Viral.load, type) %>%
+  summarise(
+    total_reads = (QC_reads * 2),
+    viral_reads = total_virus_reads,
+    ATCC_reads = sum(matched),
+    prop_ATCC = ATCC_reads / total_reads,
+    prop_viral = total_virus_reads / total_reads,
+    diff = viral_reads - ATCC_reads
+  ) %>%
+  unique()
+
+total_reads <- reads_plot %>%
+  filter(Background != "p6") %>%
+  filter(Background != "control") %>%
+  ggplot(aes(
+    x = Viral.load,
+    y = log(total_reads),
+    colour = Background
+  )) +
+  geom_point() +
+  geom_smooth(aes(group = Background), se = FALSE) +
+  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
+  theme_bw() +
+  theme(
+    axis.title.x = element_blank(),
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    axis.title.y = element_text(size = 10)
+  ) +
+  scale_x_log10() +
+  ggtitle("Total Reads") +
+  ylab("Log(Total Reads)") +
+  scale_color_manual(values = cols2, labels = c("ME_P1", "ME_P2"))
+
+viral_reads <- reads_plot %>%
+  filter(Background != "p6") %>%
+  filter(Background != "control") %>%
+  ggplot(aes(
+    x = Viral.load,
+    y = log(viral_reads),
+    colour = Background
+  )) +
+  geom_point() +
+  geom_smooth(aes(group = Background), se = FALSE) +
+  theme_bw() +
+  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
+  theme(
+    axis.title.x = element_blank(),
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    axis.title.y = element_text(size = 10)
+  ) +
+  scale_x_log10() +
+  ggtitle("Total Viral Reads") +
+  ylab("Log(Total Viral Reads)") +
+  scale_color_manual(values = cols2, labels = c("ME_P1", "ME_P2"))
+
+ATCC_reads <- reads_plot %>%
+  filter(Background != "p6") %>%
+  filter(Background != "control") %>%
+  ggplot(aes(
+    x = Viral.load,
+    y = log(ATCC_reads),
+    colour = Background
+  )) +
+  geom_point() +
+  geom_smooth(aes(group = Background), se = FALSE) +
+  theme_bw() +
+  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
+  theme(
+    axis.title.x = element_blank(),
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    axis.title.y = element_text(size = 10)
+  ) +
+  scale_x_log10() +
+  ggtitle("Spike In Viral Reads") +
+  ylab("Log(Spike In Viral Reads)") +
+  scale_color_manual(values = cols2, labels = c("ME_P1", "ME_P2"))
+
+prop_ATCC <- reads_plot %>%
+  filter(Background != "p6") %>%
+  filter(Background != "control") %>%
+  ggplot(aes(x = Viral.load, y = prop_ATCC, colour = Background)) +
+  geom_point() +
+  geom_smooth(aes(group = Background), se = FALSE) +
+  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
+  theme_bw() +
+  theme(
+    axis.title.x = element_blank(),
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    axis.title.y = element_text(size = 10)
+  ) +
+  scale_x_log10() +
+  ggtitle("Proportion Spike In Viral") +
+  ylab("Spike In Viral Reads/Total Reads") +
+  scale_color_manual(values = cols2, labels = c("ME_P1", "ME_P2"))
+
+prop_viral <- reads_plot %>%
+  filter(Background != "p6") %>%
+  filter(Background != "control") %>%
+  ggplot(aes(x = Viral.load, y = prop_viral, colour = Background)) +
+  geom_point() +
+  geom_smooth(aes(group = Background), se = FALSE) +
+  theme_bw() +
+  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
+  theme(
+    axis.title.x = element_blank(),
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    axis.title.y = element_text(size = 10)
+  ) +
+  scale_x_log10() +
+  ggtitle("Proportion Viral") +
+  ylab("Viral Reads/Total Reads") +
+  scale_color_manual(values = cols2, labels = c("ME_P1", "ME_P2"))
+
+#ggarrange(total_reads,viral_reads,ATCC_reads,prop_viral,prop_ATCC,nrow=2,ncol=3,common.legend = TRUE)
+
+ggarrange(total_reads,
+          viral_reads,
+          prop_viral,
+          nrow = 3,
+          common.legend = TRUE)
+
+#ggsave("/Users/laura/Dropbox/glasgow/github/te_ug_rodents/figures/compare_spike_ins_atcc/backgrounds_reads.png")
+
+#ggsave("/Users/laura/Dropbox/glasgow/github/te_ug_rodents/figures/manuscript_figures_pdf/FigureS3.pdf")
+
+####compare proportion viral reads per pool with shotgun####
+
+metadata3 <- metadata %>%
+  select(Sample.ID,
+         Number.of.read.pairs..quality.adaptor.trimmed.,
+         Pool.for.sequencing) %>%
+  rename(Sample_id = Sample.ID) %>%
+  rename(QC_reads = Number.of.read.pairs..quality.adaptor.trimmed.)
+
+reads_viral_dedup3 <-
+  full_join(metadata3, viral_reads_dedup, by = "Sample_id")
+
+reads_viral_nodedup3 <-
+  full_join(metadata3, viral_reads_nodedup, by = "Sample_id")
+
+reads_viral_dedup3 %>%
+  group_by(Pool.for.sequencing) %>%
+  summarise(
+    viral_reads = sum(total_virus_reads),
+    total_reads = sum(QC_reads * 2),
+    prop_viral = viral_reads / total_reads
+  )
+
+polyomics <-
+  read.csv(
+    "/Users/laura/Dropbox/glasgow/github/te_ug_rodents/data_polyomics/total_virus_mapped_reads_per_sample_dedup.csv"
+  )
+
+#read data from polyomics_indexes_stefano document
+total_reads <- data.frame(total_reads = c(5942760, 7714275))
+
+keeps <- c("RNA-Msp-p2", "RNA-Msp-p8")
+
+polyomics_read_prop <- polyomics %>%
+  filter(Sample_id %in% keeps)
+
+polyomics_read_prop2 <- cbind(polyomics_read_prop, total_reads)
+
+polyomics_read_prop2 %>%
+  summarise(prop_viral = total_virus_reads / total_reads)
+
+```
+
+### --------------------
+
 ## Figure S4
@@
 
+### --------------------
+
 ## >>> Figure 2: plots of TE experiment data - genome coverage & per site coverage
@@
 
+unzip(zipfile = "data_TE/TE_sequencing_experiment_readdepth_per_site_sample_and_virus_bowtie2_dedup_atcc_ref.tsv.zip", exdir = "data_TE/")
+
+unzip(zipfile = "data_TE/TE_sequencing_experiment_readdepth_per_site_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv.zip", exdir = "data_TE/")
+
 dedup_per_site <-
   read_tsv(
-    "data_TE/TE_sequencing_experiment_readdepth_per_site
-
-    TEdata_TE/TE_sequencing_experiment_readdepth_per_site_sample_and_virus_bowtie2_dedup_atcc_ref.tsv"
+    "data_TE/TE_sequencing_experiment_readdepth_per_site_sample_and_virus_bowtie2_dedup_atcc_ref.tsv"
   )
@@
   )
+
+file.remove("data_TE/TE_sequencing_experiment_readdepth_per_site_sample_and_virus_bowtie2_dedup_atcc_ref.tsv")
+
+file.remove( "data_TE/TE_sequencing_experiment_readdepth_per_site_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv")
 
@@
 
-metadata2 <- metadata |>
+metadata2 <- 
+  metadata |>
   select(Sample.ID, Number.of.read.pairs..quality.adaptor.trimmed.) |>
@@
 #import and combine read count files
+
 #deduplicated and non-deduplicated
@@
 counts_dedup_bt <-
-  read.table(
-    "data_TE/TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_dedup_atcc_ref.tsv",
-    sep = "\t",
-    header = TRUE
+  read_tsv(
+    "data_TE/TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_dedup_atcc_ref.tsv"
   )
@@
 counts_nodedup_bt <-
-  read.table(
-    "data_TE/TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv",
-    sep = "\t",
-    header = TRUE
+  read_tsv(
+    "data_TE/TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv"
   )
@@
 
-counts_reads_norm<-counts_reads |>
-  mutate(norm_counts1 = matched/QC_reads) |>
-  mutate(norm_counts2 = matched/QC_reads/length) |>
-  mutate(norm_counts3 = matched/length) |>
-  mutate(genome_structure = case_when((virus == "Human_adenovirus_40"| virus == "Human_betaherpesvirus") ~ "DNA",
-                                      ,.default = "RNA"))
-
-lengths<-counts_reads_norm |>
-  select(virus,length) |>
+counts_reads_norm <- 
+  counts_reads |>
+  mutate(norm_counts1 = matched / QC_reads) |>
+  mutate(norm_counts2 = matched / QC_reads / length) |>
+  mutate(norm_counts3 = matched / length) |>
+  mutate(genome_structure = case_when((
+    virus == "Human_adenovirus_40" |
+      virus == "Human_betaherpesvirus"
+  ) ~ "DNA",
+  .default = "RNA"
+  ))
+
+lengths <- 
+  counts_reads_norm |>
+  select(virus, length) |>
   distinct()
@@
 
-persite_coverage_dedup<-dedup_per_site |>
-  group_by(Sample_id,virus,Viral.load,Background,type) |>
+persite_coverage_dedup <- 
+  dedup_per_site |>
+  rename(Viral.load = `Viral load`) |>
+  group_by(Sample_id, virus, Viral.load, Background, type) |>
   filter(coverage > 0) |>
@@
   left_join(lengths) |>
-  mutate(percent_coverage = genome_coverage/length)
-
-persite_coverage_nodedup<-nodedup_per_site |>
-  group_by(Sample_id,virus,Viral.load,Background,type) |>
+  mutate(percent_coverage = genome_coverage / length)
+
+persite_coverage_nodedup <- 
+  nodedup_per_site |>
+  rename(Viral.load = `Viral load`) |>
+  group_by(Sample_id, virus, Viral.load, Background, type) |>
   filter(coverage > 0) |>
@@
   left_join(lengths) |>
-  mutate(percent_coverage = genome_coverage/length)
-
-persite_coverage_both<-rbind(persite_coverage_dedup,persite_coverage_nodedup)
-
-coverage_labels<-c("0" = "Control",
-                   "100" = "1e+02",
-                   "1000" = "1e+03",
-                   "100000" = "1e+05",
+  mutate(percent_coverage = genome_coverage / length)
+
+persite_coverage_both <-
+  rbind(persite_coverage_dedup, 
+        persite_coverage_nodedup)
+
+coverage_labels <- c("0" = "Control",
+                     "100" = "10^{2}",
+                     "1000" = "10^{3}",
+                     "10000" = "10^{5}",
                      "dedup_TE" = "Deduplicated",
@@
 
-coverage_labels2<-c("0" = "Control",
-                    "100" = "1e+02",
-                    "1000" = "1e+03",
-                    "100000" = "1e+05")
-
-cols2<-c("#BB5566","#004488")
-
-cols4<-c("#DDAA33","#BB5566","#004488")
-
-persite_coverage_both |>
-  filter(Background != "p6") |>
-  filter(Background != "control") |>
-  ggplot(aes(x=virus,y=percent_coverage,colour = Background))+
-  geom_boxplot(outlier.shape=NA)+
-  geom_point(position = position_dodge(width = .75))+
-  facet_grid(type~as.character(Viral.load),labeller=as_labeller(coverage_labels))+
-  theme_bw()+
-  theme(axis.title.y=element_text(size=10))+
-  ylab("Genome coverage")+
-  theme(axis.title.x=element_blank(),axis.text.x=element_text(angle=45,hjust=1),axis.title.y=element_text(size=15),axis.text=element_text(size=12),strip.text=element_text(size=12))+
-  scale_color_manual(values=cols2,labels=c("ME_P1","ME_P2"))+
-  scale_x_discrete(labels=c("Human adenovirus 40","Human betaherpesvirus","Human respiratory syncytial virus","Influenza B virus","Mammalian orthoreovirus 3","Zika virus"))+
-  scale_y_continuous(limits=c(0,1))
-
-#ggsave("figures/compare_spike_ins_atcc/genome_cov_split.png",width=15,height=6)
+coverage_labels2 <- c("0" = "Control",
+                      "100" = "10^{2}",
+                      "1000" = "10^{3}",
+                      "10000" = "10^{5}")
+```
+
+### Make plots
+```{r, echo = FALSE, warning=FALSE, message=FALSE, fig.height=4, fig.width=6}
+
+cols2 <- c("#BB5566",
+           "#004488")
+
+cols4 <- c("#DDAA33",
+           "#BB5566",
+           "#004488")
+
+persite_coverage_both$Viral.load <- factor(persite_coverage_both$Viral.load, labels = c("0",
+    "10^{2}", 
+    "10^{3}", 
+    "10^{5}"
+    ))
 
@@
   filter(type == "dedup_TE") |>
-  ggplot(aes(x=virus,y=percent_coverage,colour = Background))+
-  geom_boxplot(outlier.shape=NA)+
-  geom_point(position = position_dodge(width = .75))+
-  facet_grid(~as.character(Viral.load),labeller=as_labeller(coverage_labels2))+
-  theme_bw()+
-  ylab("Genome coverage")+
-  theme(axis.title.x=element_blank(),axis.text.x=element_text(angle=45,hjust=1),axis.title.y=element_text(size=15),axis.text=element_text(size=12),strip.text=element_text(size=12),legend.text = element_text(size=10))+
-  scale_color_manual(values=cols2,labels=c("ME_P1","ME_P2"))+
-  scale_x_discrete(labels=c("Human adenovirus 40","Human betaherpesvirus","Human respiratory syncytial virus","Influenza B virus","Mammalian orthoreovirus 3","Zika virus"))+
-  scale_y_continuous(limits=c(0,1))
-
+  ggplot(aes(x = virus, y = percent_coverage, colour = Background)) +
+  geom_boxplot(outlier.shape = NA) +
+  geom_point(position = position_dodge(width = .75)) +
+  facet_grid(~as.character(Viral.load), labeller = label_parsed) +
+  theme_few() +
+  ylab("Genome coverage") +
+  theme(
+    axis.title.x = element_text(size = 12),
+    axis.text.x = element_text(angle = 30, hjust = 1),
+    axis.title.y = element_text(size = 12),
+    legend.title = element_blank(),
+    legend.position = "top", 
+    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
+    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
+  ) +
+  scale_color_manual(values = cols2, labels = c("P1", "P2")) +
+  scale_x_discrete(
+    labels = c(
+      "Human adenovirus 40",
+      "Human betaherpesvirus",
+      "Human respiratory syncytial virus",
+      "Influenza B virus",
+      "Mammalian orthoreovirus 3",
+      "Zika virus"
+    )
+  ) +
+  scale_y_continuous(limits = c(0, 1))
+
+```
+
+### ***Final plot***
+```{r, echo = FALSE, warning=FALSE, message=FALSE}
 #ggsave("figures/compare_spike_ins_atcc/genome_cov_deduponly.png",width=15,height=6)
 
-#ggsave("figures/manuscript_figures_pdf/Figure2.pdf",width=15,height=6)
-
-ggplot(persite_coverage_both,aes(x=as.character(Viral.load),y=percent_coverage))+
-  geom_point()+
-  geom_boxplot(outlier.shape=NA)+
-  facet_wrap(~type)+
-  ylab("Genome coverage")+
-  xlab("Viral load")
-
-#ggsave("figures/compare_spike_ins_atcc/genome_cov_combined.pdf")
-
-persite_both<-rbind(dedup_per_site,nodedup_per_site)
-
-persite_reads<-left_join(persite_both,metadata2,by="Sample_id")
-
-persite_norm<-persite_reads |>
-  mutate(norm_cov = coverage/QC_reads)
-
-coverage_labels3<-c("100" = "1e+02",
+ggsave(
+  "figures/manuscript_figures_pdf/Figure_2.pdf",
+  width = 15,
+  height = 6
+)
+
+```
+
+### --------------------
+
+### Figure S5 - linear model
+```{r}
+#Linear model for TE data - separated by virus
+#November 2024
+
+#viral read depth
+#use bowtie2 data
+
+rm(list = ls())
+
+library(tidyverse)
+library(broom)
+
+####read counts/normalised read counts####
+
+#metadata
+
+metadata <-
+  read.csv(
+    "/Users/laura/Dropbox/glasgow/github/te_ug_rodents/metadata/sampleIDs_TESpikeIn.csv",
+    header = TRUE
+  )
+
+metadata2 <- metadata %>%
+  select(Sample.ID, Number.of.read.pairs..quality.adaptor.trimmed.) %>%
+  rename(Sample_id = Sample.ID) %>%
+  rename(QC_reads = Number.of.read.pairs..quality.adaptor.trimmed.)
+
+#import and combine read count files
+#deduplicated and non-deduplicated
+
+counts_dedup_bt <-
+  read.table(
+    "/Users/laura/Dropbox/glasgow/github/te_ug_rodents/data_TE/TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_dedup_atcc_ref.tsv",
+    sep = "\t",
+    header = TRUE
+  )
+
+counts_nodedup_bt <-
+  read.table(
+    "/Users/laura/Dropbox/glasgow/github/te_ug_rodents/data_TE/TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv",
+    sep = "\t",
+    header = TRUE
+  )
+
+counts_bt_all <- rbind(counts_dedup_bt, counts_nodedup_bt)
+
+counts_reads <- left_join(counts_bt_all, metadata2, by = "Sample_id")
+
+####read depths####
+
+#import and combine read depth files
+
+depth_dedup_bt <-
+  read.table(
+    "/Users/laura/Dropbox/glasgow/github/te_ug_rodents/data_TE/TE_sequencing_experiment_readdepth_per_sample_and_virus_bowtie2_dedup_atcc_ref.tsv",
+    sep = "\t",
+    header = TRUE
+  )
+
+depth_nodedup_bt <-
+  read.table(
+    "/Users/laura/Dropbox/glasgow/github/te_ug_rodents/data_TE/TE_sequencing_experiment_readdepth_per_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv",
+    sep = "\t",
+    header = TRUE
+  )
+
+depths_bt_all <- rbind(depth_dedup_bt, depth_nodedup_bt)
+
+depths_reads <- left_join(depths_bt_all, metadata2, by = "Sample_id") %>%
+  mutate(genome_structure = case_when((
+    virus == "Human_adenovirus_40" |
+      virus == "Human_betaherpesvirus"
+  ) ~ "DNA",
+  ,
+  .default = "RNA"
+  ))
+
+####linear model for spike in viruses - mean read depth####
+
+depths_reads_sub <- depths_reads %>%
+  filter(Background != "p6") %>%
+  filter(Background != "control") %>%
+  filter(type == "dedup_TE") %>%
+  mutate(log_depth = log(mean_depth)) %>%
+  mutate(log_viral_load = log10(Viral.load))
+
+#inspect data
+hist(depths_reads_sub$mean_depth)
+hist(depths_reads_sub$Viral.load)
+hist(depths_reads_sub$log_depth)
+hist(depths_reads_sub$log_viral_load)
+
+summary(depths_reads_sub$log_depth)
+summary(depths_reads_sub$log_viral_load)
+
+#all viruses together
+
+lm1 <-
+  glm(log_depth ~ log_viral_load * virus,
+      data = depths_reads_sub,
+      family = "gaussian")
+
+summary(lm1)
+
+glm.diag.plots(lm1)
+
+pred <- predict(lm1, type = "response")
+
+rsq <- function (x, y)
+  cor(x, y) ^ 2
+
+rsq(depths_reads_sub$log_depth, pred)
+
+##split by viruses
+
+cols <- c("#4477AA",
+          "#66CCEE",
+          "#228833",
+          "#CCBB44",
+          "#EE6677",
+          "#AA3377")
+
+facet_names <- c(
+  "Human_adenovirus_40" = "Human adenovirus 40",
+  "Human_betaherpesvirus" = "Human betaherpesvirus",
+  "Human_respiratory_syncytial_virus" = "Human respiratory syncytial virus",
+  "Influenza_B_virus" = "Influenza B virus",
+  "Mammalian_orthoreovirus3" = "Mammalian orthoreovirus 3",
+  "Zika_virus" = "Zika virus"
+)
+
+virus <- (unique(depths_reads_sub$virus)) %>%
+  rep(., each = 2) %>%
+  data.frame()
+
+list_models <- depths_reads_sub %>%
+  group_split(virus) %>%
+  map( ~ lm(log_depth ~ log_viral_load, data = .))
+
+lm_tidy <- map(list_models, broom::tidy) %>%
+  do.call(rbind.data.frame, .) %>%
+  cbind(virus, .) %>%
+  rename(virus = ".") %>%
+  select(virus, term, estimate) %>%
+  pivot_wider(names_from = "term", values_from = "estimate")
+
+lm_summary <- depths_reads_sub %>%
+  group_by(virus) %>%
+  summarise(
+    Intercept = lm(log_depth ~ log_viral_load)$coefficients[1],
+    Coeff_x1 = lm(log_depth ~ log_viral_load)$coefficients[2],
+    R2 = summary(lm(log_depth ~ log_viral_load))$r.squared,
+    pvalue = summary(lm(log_depth ~ log_viral_load))$coefficients["log_viral_load", 4]
+  )
+
+lm_combined <- left_join(lm_tidy, lm_summary, by = "virus")
+
+ggplot() +
+  geom_point(data = depths_reads_sub,
+             mapping = aes(log_viral_load, log_depth, color = virus)) +
+  geom_abline(data = lm_combined, aes(intercept = `(Intercept)`, slope = log_viral_load)) +
+  geom_text(data = lm_combined,
+            colour = "black",
+            aes(
+              x = 4.7,
+              y = 3,
+              label = round(Coeff_x1, digits = 2)
+            )) +
+  geom_text(data = lm_combined,
+            colour = "black",
+            aes(x = 4, y = 3, label = "Slope =")) +
+  geom_text(data = lm_combined,
+            colour = "black",
+            aes(
+              x = 4.7,
+              y = 2,
+              label = round(R2, digits = 2)
+            )) +
+  geom_text(data = lm_combined,
+            colour = "black",
+            aes(x = 4, y = 2, label = "R2 =")) +
+  facet_wrap( ~ virus, labeller = as_labeller(facet_names)) +
+  theme_bw() +
+  xlab("Log(Viral Load)") +
+  ylab("Log(Mean Read Depth)") +
+  scale_color_manual(values = cols) +
+  guides(color = FALSE)
+
+#ggsave("/Users/laura/Dropbox/glasgow/github/te_ug_rodents/figures/compare_spike_ins_atcc/model_readdepth_dedup.png")
+
+#ggsave("/Users/laura/Dropbox/glasgow/github/te_ug_rodents/figures/manuscript_figures_pdf/FigureS5.pdf")
+
+```
+
+### --------------------
+
+### Figure S6
+```{r}
+
+persite_both <- rbind(dedup_per_site, nodedup_per_site)
+
+persite_reads <- left_join(persite_both, metadata2, by = "Sample_id")
+
+persite_norm <- persite_reads |>
+  mutate(norm_cov = coverage / QC_reads)
+
+coverage_labels3 <-c("100" = "1e+02",
                     "1000" = "1e+03",
@@
 
+persite_norm$`Viral load` <- 
+  factor(persite_norm$`Viral load`, 
+         labels = c("0",
+                    "10^{2}", 
+                    "10^{3}",
+                    "10^{5}"))
+
 persite_norm |>
@@
   filter(virus == "Human_respiratory_syncytial_virus") |>
-  ggplot(aes(x=site,y=coverage,fill=Background))+
-  geom_col()+
-  facet_wrap(Viral.load~Sample_id)+
-  ylab("Log(Coverage)")+
-  ggtitle("Human RSV (deduplicated)")+
-  guides(fill=guide_legend(title="Background"))+
-  theme_bw()+
-  scale_y_log10()+
-  scale_fill_manual(values=cols2,labels=c("ME_P1","ME_P2"))
-
+  rename(Viral.load = `Viral load`) |>
+  ggplot(aes(x = site, y = coverage, fill = Background)) +
+  geom_col() +
+  facet_wrap(Viral.load ~ Sample_id, label = label_parsed) +
+  ylab("Log(Coverage)") +
+  ggtitle("Human RSV (deduplicated)") +
+  guides(fill = guide_legend(title = "Background")) +
+  theme_few() +
+  theme(
+    axis.title.x = element_text(size = 12),
+    axis.text.x = element_text(angle = 30, hjust = 1),
+    axis.title.y = element_text(size = 12),
+    legend.title = element_blank(),
+    legend.position = "top", 
+    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
+    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
+  ) +
+  scale_y_log10() +
+  scale_fill_manual(values = cols2, labels = c("P1", "P2"))
+```
+
+#### ***Final plot***
+```{r}
 #ggsave("figures/manuscript_figures_pdf/FigureS5.pdf",width=10,height=8)
+```
+
+### --------------------
+
+### Figure S7
+```{r}
 
@@
   filter(virus == "Zika_virus") |>
-  ggplot(aes(x=site,y=coverage,fill=Background))+
-  geom_col()+
-  facet_wrap(Viral.load~Sample_id)+
-  ylab("Log(Coverage)")+
-  ggtitle("Zika virus (deduplicated)")+
-  guides(fill=guide_legend(title="Background"))+
-  theme_bw()+
-  scale_y_log10()+
-  scale_fill_manual(values=cols2,labels=c("ME_P1","ME_P2"))
+  ggplot(aes(x = site, y = coverage, fill = Background)) +
+  geom_col() +
+  facet_wrap(Viral.load ~ Sample_id) +
+  ylab("Log(Coverage)") +
+  ggtitle("Zika virus (deduplicated)") +
+  guides(fill = guide_legend(title = "Background")) +
+  theme_bw() +
+  scale_y_log10() +
+  scale_fill_manual(values = cols2, labels = c("ME_P1", "ME_P2"))
 
@@
 
+```
+
+### --------------------
+
+### Figure S8
+```{r}
 persite_norm |>
@@
   filter(virus == "Human_adenovirus_40") |>
-  ggplot(aes(x=site,y=coverage,fill=Background))+
-  geom_col()+
-  facet_wrap(Viral.load~Sample_id)+
-  ylab("Log(Coverage)")+
-  ggtitle("Human Adenovirus (deduplicated)")+
-  guides(fill=guide_legend(title="Background"))+
-  theme_bw()+
-  scale_y_log10()+
-  scale_fill_manual(values=cols2,labels=c("ME_P1","ME_P2"))
+  ggplot(aes(x = site, y = coverage, fill = Background)) +
+  geom_col() +
+  facet_wrap(Viral.load ~ Sample_id) +
+  ylab("Log(Coverage)") +
+  ggtitle("Human Adenovirus (deduplicated)") +
+  guides(fill = guide_legend(title = "Background")) +
+  theme_bw() +
+  scale_y_log10() +
+  scale_fill_manual(values = cols2, labels = c("ME_P1", "ME_P2"))
 
 #ggsave("figures/manuscript_figures_pdf/FigureS7.pdf",width=20,height=18)
-
+```
+
+### --------------------
+
+### Figure S9
+```{r}
 persite_norm |>
@@
   filter(virus == "Human_betaherpesvirus") |>
-  ggplot(aes(x=site,y=coverage,fill=Background))+
-  geom_polygon()+
-  facet_wrap(Viral.load~Sample_id)+
-  ylab("Log(Coverage)")+
-  ggtitle("Human Betaherpesvirus (deduplicated)")+
-  guides(fill=guide_legend(title="Background"))+
-  theme_bw()+
-  scale_y_log10()+
-  scale_fill_manual(values=cols2,labels=c("ME_P1","ME_P2"))
+  ggplot(aes(x = site, y = coverage, fill = Background)) +
+  geom_polygon() +
+  facet_wrap(Viral.load ~ Sample_id) +
+  ylab("Log(Coverage)") +
+  ggtitle("Human Betaherpesvirus (deduplicated)") +
+  guides(fill = guide_legend(title = "Background")) +
+  theme_bw() +
+  scale_y_log10() +
+  scale_fill_manual(values = cols2, labels = c("ME_P1", "ME_P2"))
 
 #ggsave("figures/compare_spike_ins_atcc/HBHV_coverage_test.pdf",width=20,height=18)
-
+```
+
+### --------------------
+
+### Figure S10
+```{r}
 #have to do these differently due to segmentation
 
-FLU_ME1<-persite_norm |>
+FLU_ME1 <- persite_norm |>
   filter(virus == "Influenza_B_virus") |>
@@
   filter(Background == "p2") |>
-  ggplot(aes(x=site,y=coverage,fill=as.character(Viral.load)))+
-  geom_col()+
-  facet_grid(seg~Sample_id)+
-  ylab("Log(Coverage)")+
-  ggtitle("Influenza B virus (Background ME_1 deduplicated)")+
-  guides(fill=guide_legend(title="Viral load"))+
-  theme_bw()+
-  scale_y_log10()+
-  scale_fill_manual(values=cols4,labels=c("1e+02","1e+03","1e+05"))
-
-FLU_ME2<-persite_norm |>
+  ggplot(aes(
+    x = site,
+    y = coverage,
+    fill = as.character(Viral.load)
+  )) +
+  geom_col() +
+  facet_grid(seg ~ Sample_id) +
+  ylab("Log(Coverage)") +
+  ggtitle("Influenza B virus (Background ME_1 deduplicated)") +
+  guides(fill = guide_legend(title = "Viral load")) +
+  theme_bw() +
+  scale_y_log10() +
+  scale_fill_manual(values = cols4,
+                    labels = c("1e+02", "1e+03", "1e+05"))
+
+FLU_ME2 <- persite_norm |>
   filter(virus == "Influenza_B_virus") |>
@@
   filter(Background == "p8") |>
-  ggplot(aes(x=site,y=coverage,fill=as.character(Viral.load)))+
-  geom_col()+
-  facet_grid(seg~Sample_id)+
-  ylab("Log(Coverage)")+
-  ggtitle("Influenza B virus (Background ME_2 deduplicated)")+
-  guides(fill=guide_legend(title="Viral load"))+
-  theme_bw()+
-  scale_y_log10()+
-  scale_fill_manual(values=cols4,labels=c("1e+02","1e+03","1e+05"))
-
-ggarrange(FLU_ME1,FLU_ME2,ncol=2,common.legend=TRUE)
+  ggplot(aes(
+    x = site,
+    y = coverage,
+    fill = as.character(Viral.load)
+  )) +
+  geom_col() +
+  facet_grid(seg ~ Sample_id) +
+  ylab("Log(Coverage)") +
+  ggtitle("Influenza B virus (Background ME_2 deduplicated)") +
+  guides(fill = guide_legend(title = "Viral load")) +
+  theme_bw() +
+  scale_y_log10() +
+  scale_fill_manual(values = cols4,
+                    labels = c("1e+02", "1e+03", "1e+05"))
+
+ggarrange(FLU_ME1, FLU_ME2, ncol = 2, common.legend = TRUE)
 
 #ggsave("figures/manuscript_figures_pdf/FigureS9.pdf",width=20,height=18)
-
-REO_ME1<-persite_norm |>
+```
+
+### --------------------
+
+### Figure S11
+```{r}
+
+REO_ME1 <- persite_norm |>
 filter(virus == "Mammalian_orthoreovirus3") |>
@@
 filter(Background == "p2") |>
-  ggplot(aes(x=site,y=coverage,fill=as.character(Viral.load)))+
-  geom_col()+
-  facet_grid(seg~Sample_id)+
-  ylab("Log(Coverage)")+
-  ggtitle("Mammalian orthoreovirus 3 (Background ME_1 deduplicated)")+
-  guides(fill=guide_legend(title="Viral load"))+
-  theme_bw()+
-  scale_y_log10()+
-  scale_fill_manual(values=cols4,labels=c("1e+02","1e+03","1e+05"))
-
-REO_ME2<-persite_norm |>
+ggplot(aes(
+x = site,
+y = coverage,
+fill = as.character(Viral.load)
+)) +
+geom_col() +
+facet_grid(seg ~ Sample_id) +
+ylab("Log(Coverage)") +
+ggtitle("Mammalian orthoreovirus 3 (Background ME_1 deduplicated)") +
+guides(fill = guide_legend(title = "Viral load")) +
+theme_bw() +
+scale_y_log10() +
+scale_fill_manual(values = cols4,
+labels = c("1e+02", "1e+03", "1e+05"))
+
+REO_ME2 <- persite_norm |>
 filter(virus == "Mammalian_orthoreovirus3") |>
@@
 filter(Background == "p8") |>
-  ggplot(aes(x=site,y=coverage,fill=as.character(Viral.load)))+
-  geom_col()+
-  facet_grid(seg~Sample_id)+
-  ylab("Log(Coverage)")+
-  ggtitle("Mammalian orthoreovirus 3 (Background ME_2 deduplicated)")+
-  guides(fill=guide_legend(title="Viral load"))+
-  theme_bw()+
-  scale_y_log10()+
-  scale_fill_manual(values=cols4,labels=c("1e+02","1e+03","1e+05"))
-
-ggarrange(REO_ME1,REO_ME2,ncol=2,common.legend=TRUE)
+ggplot(aes(
+x = site,
+y = coverage,
+fill = as.character(Viral.load)
+)) +
+geom_col() +
+facet_grid(seg ~ Sample_id) +
+ylab("Log(Coverage)") +
+ggtitle("Mammalian orthoreovirus 3 (Background ME_2 deduplicated)") +
+guides(fill = guide_legend(title = "Viral load")) +
+theme_bw() +
+scale_y_log10() +
+scale_fill_manual(values = cols4,
+labels = c("1e+02", "1e+03", "1e+05"))
+
+ggarrange(REO_ME1, REO_ME2, ncol = 2, common.legend = TRUE)
 
@@
 
-## Figure 3: all viruses - target and off target
+### --------------------
+
+## >>>> Figure 3: all viruses - target and off target
 
@@
   theme(
-    axis.title.x = element_text(size = 12, family = "Helvetica Neue"),
+    axis.title.x = element_text(size = 12),
     axis.text.x = element_text(angle = 30, hjust = 1),
-    axis.title.y = element_text(size = 12, family = "Helvetica Neue"),
+    axis.title.y = element_text(size = 12),
     legend.title = element_blank(),
@@
     theme(
-    axis.title.x = element_text(size = 12, family = "Helvetica Neue"),
+    axis.title.x = element_text(size = 12),
     axis.text.x = element_text(angle = 30, hjust = 1),
-    axis.title.y = element_text(size = 12, family = "Helvetica Neue"),
+    axis.title.y = element_text(size = 12),
     legend.title = element_blank(),
@@
 ```
+
+### --------------------

### 2025-01-03 manuscript_figures.Rmd
update manuscript_figures.Rmd

@@
   ) ~ "DNA",
-  ,
   .default = "RNA"
@@
 ## Figure S2
+
+### Reaed-in data / Make plots
 ```{r}
@@
   read.csv(
-    "/Users/laura/Dropbox/glasgow/github/te_ug_rodents/metadata/sampleIDs_TESpikeIn.csv",
-    header = TRUE
-  )
-
-metadata2 <- metadata %>%
-  select(Sample.ID, Number.of.read.pairs..quality.adaptor.trimmed.) %>%
-  rename(Sample_id = Sample.ID) %>%
+    "metadata/sampleIDs_TESpikeIn.csv",
+    header = TRUE
+  )
+
+metadata2 <- metadata |>
+  select(Sample.ID, Number.of.read.pairs..quality.adaptor.trimmed.) |>
+  rename(Sample_id = Sample.ID) |>
   rename(QC_reads = Number.of.read.pairs..quality.adaptor.trimmed.)
@@
   read.table(
-    "/Users/laura/Dropbox/glasgow/github/te_ug_rodents/data_TE/TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_dedup_atcc_ref.tsv",
+    "data_TE/TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_dedup_atcc_ref.tsv",
     sep = "\t",
@@
   read.table(
-    "/Users/laura/Dropbox/glasgow/github/te_ug_rodents/data_TE/TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv",
+    "data_TE/TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv",
     sep = "\t",
@@
 
-counts_reads_norm <- counts_reads %>%
-  mutate(norm_counts1 = matched / QC_reads) %>%
-  mutate(norm_counts2 = matched / QC_reads / length) %>%
-  mutate(norm_counts3 = matched / length) %>%
+counts_reads_norm <- counts_reads |>
+  mutate(norm_counts1 = matched / QC_reads) |>
+  mutate(norm_counts2 = matched / QC_reads / length) |>
+  mutate(norm_counts3 = matched / length) |>
   mutate(genome_structure = case_when((
@@
 
-metadata3 <- metadata %>%
-  select(Sample.ID, Pool.for.sequencing) %>%
-  rename(Sample_id = Sample.ID) %>%
+metadata3 <- 
+  metadata |>
+  select(Sample.ID, Pool.for.sequencing) |>
+  rename(Sample_id = Sample.ID) |>
   rename(Pool = Pool.for.sequencing)
@@
 
-pool_readcounts <- counts_pool %>%
-  filter(Background != "p6") %>%
-  filter(Background != "control") %>%
+```
+
+### Make plots
+```{r}
+
+pool_readcounts <- 
+  counts_pool |>
+  filter(Background != "p6") |>
+  filter(Background != "control") |>
   ggplot(aes(x = virus, y = log(matched))) +
@@
   facet_grid(Pool ~ type) +
-  theme_bw() +
-  theme(
-    axis.title.x = element_blank(),
-    axis.text.x = element_blank(),
-    axis.title.y = element_text(size = 10)
-  ) +
-  ylab("Log(Viral Reads)") +
+  theme_few() +
+  theme(
+    # axis.title.x = element_blank(),
+    # axis.text.x = element_blank(),
+    axis.title.y = element_text(size = 10),
+    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
+    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
+  ) +
+  ylab("Log (Viral Reads)") + xlab("Viral load (copies / ml)") +
   scale_color_manual(values = cols3, labels = c("Pool 1", "Pool 2"))
 
-pool_counts_sum <- counts_pool %>%
-  filter(Background != "p6") %>%
-  filter(Background != "control") %>%
+pool_counts_sum <- 
+  counts_pool |>
+  filter(Background != "p6") |>
+  filter(Background != "control") |>
   ggplot(aes(
@@
   facet_grid( ~ type, labeller = as_labeller(facet_names)) +
-  theme_bw() +
-  theme(axis.title.x = element_blank(), axis.title.y = element_text(size =
-                                                                      10)) +
-  ylab("Log(Viral Reads)") +
+  theme_few() +
+  theme(
+    # axis.title.x = element_blank(),
+    # axis.text.x = element_blank(),
+    axis.title.y = element_text(size = 10),
+    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
+    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
+  ) +
+  ylab("Log (Viral Reads)") + xlab("Viral load (copies / ml)") +
   scale_color_manual(values = cols3, labels = c("Pool 1", "Pool 2"))
 
-pool_readcounts_norm <- counts_pool %>%
-  filter(Background != "p6") %>%
-  filter(Background != "control") %>%
+pool_readcounts_norm <- 
+  counts_pool |>
+  filter(Background != "p6") |>
+  filter(Background != "control") |>
   ggplot(aes(x = virus, y = log(norm_counts1))) +
@@
   facet_grid(Pool ~ type) +
-  theme(
-    axis.title.x = element_blank(),
-    axis.text.x = element_blank(),
-    axis.title.y = element_text(size = 10)
-  ) +
-  ylab("Log(viral reads/cleaned reads)")
-
-pool_norm1_sum <- counts_pool %>%
-  filter(Background != "p6") %>%
-  filter(Background != "control") %>%
+  theme_few() +
+  theme(
+    # axis.title.x = element_blank(),
+    # axis.text.x = element_blank(),
+    axis.title.y = element_text(size = 10),
+    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
+    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
+  ) +
+  ylab("Log(viral reads/cleaned reads)") +  xlab("Viral load (copies / ml)")
+
+pool_norm1_sum <- 
+  counts_pool |>
+  filter(Background != "p6") |>
+  filter(Background != "control") |>
   ggplot(aes(
@@
   facet_grid( ~ type, labeller = as_labeller(facet_names)) +
-  theme_bw() +
-  theme(axis.title.x = element_blank(), axis.title.y = element_text(size =
-                                                                      10)) +
-  ylab("Log(viral reads/cleaned reads)") +
+  theme_few() +
+  theme(
+    # axis.title.x = element_blank(),
+    # axis.text.x = element_blank(),
+    axis.title.y = element_text(size = 10),
+    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
+    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
+  ) +
+  ylab("Log(viral reads/cleaned reads)") +   xlab("Viral load (copies / ml)") +
   scale_color_manual(values = cols3, labels = c("Pool 1", "Pool 2"))
 
-pool_readcounts_norm2 <- counts_pool %>%
-  filter(Background != "p6") %>%
-  filter(Background != "control") %>%
+pool_readcounts_norm2 <- 
+  counts_pool |>
+  filter(Background != "p6") |>
+  filter(Background != "control") |>
   ggplot(aes(x = virus, y = log(norm_counts2))) +
@@
   facet_grid(Pool ~ type) +
-  theme(
-    axis.title.x = element_blank(),
-    axis.text.x = element_text(angle = 45, hjust = 1),
-    axis.title.y = element_text(size = 10)
-  ) +
-  ylab("Log(viral reads/cleaned reads/genome length)")
-
-pool_norm2_sum <- counts_pool %>%
-  filter(Background != "p6") %>%
-  filter(Background != "control") %>%
+  theme_few() +
+  theme(
+    # axis.title.x = element_blank(),
+    # axis.text.x = element_blank(),
+    axis.title.y = element_text(size = 10),
+    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
+    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
+  ) +
+  ylab("Log(viral reads/cleaned reads/genome length)") + xlab("Viral load (copies / ml)")
+
+pool_norm2_sum <- 
+  counts_pool |>
+  filter(Background != "p6") |>
+  filter(Background != "control") |>
   ggplot(aes(
@@
   facet_grid( ~ type, labeller = as_labeller(facet_names)) +
-  theme_bw() +
-  theme(axis.title.x = element_blank(), axis.title.y = element_text(size =
-                                                                      10)) +
+  theme_few() +
+  theme(
+    # axis.title.x = element_blank(),
+    # axis.text.x = element_blank(),
+    axis.title.y = element_text(size = 10),
+    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
+    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
+  ) +
   ylab("Log(viral reads/cleaned reads/genome length)") +
-  scale_color_manual(values = cols3, labels = c("Pool 1", "Pool 2"))
+  scale_color_manual(values = cols3, labels = c("Pool 1", "Pool 2")) + 
+  xlab("Viral load (copies / ml)")
 
@@
   read.table(
-    "/Users/laura/Dropbox/glasgow/github/te_ug_rodents/data_TE/TE_sequencing_experiment_readdepth_per_sample_and_virus_bowtie2_dedup_atcc_ref.tsv",
+    "data_TE/TE_sequencing_experiment_readdepth_per_sample_and_virus_bowtie2_dedup_atcc_ref.tsv",
     sep = "\t",
@@
   read.table(
-    "/Users/laura/Dropbox/glasgow/github/te_ug_rodents/data_TE/TE_sequencing_experiment_readdepth_per_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv",
+    "data_TE/TE_sequencing_experiment_readdepth_per_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv",
     sep = "\t",
@@
 
-depths_reads <- left_join(depths_bt_all, metadata2, by = "Sample_id") %>%
+depths_reads <- 
+  left_join(depths_bt_all, metadata2, by = "Sample_id") |>
   mutate(genome_structure = case_when((
@@
 
-pool_readdepths <- depths_pool %>%
-  filter(Background != "p6") %>%
-  filter(Background != "control") %>%
+pool_readdepths <- 
+  depths_pool |>
+  filter(Background != "p6") |>
+  filter(Background != "control") |>
   ggplot(aes(x = virus, y = log(mean_depth))) +
@@
   facet_grid(Pool ~ type) +
+  theme_few() +
   theme(
@@
     axis.text.x = element_text(angle = 45, hjust = 1),
-    axis.title.y = element_text(size = 10)
-  ) +
-  ylab("Log(mean read depth)")
-
-pool_depths_sum <- depths_pool %>%
-  filter(Background != "p6") %>%
-  filter(Background != "control") %>%
+    axis.title.y = element_text(size = 10),
+    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
+    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
+  ) +
+  ylab("Log (mean read depth)") + xlab("Viral load (copies / ml)")
+
+pool_depths_sum <- 
+  depths_pool |>
+  filter(Background != "p6") |>
+  filter(Background != "control") |>
   ggplot(aes(
@@
   facet_grid( ~ type, labeller = as_labeller(facet_names)) +
-  theme_bw() +
-  theme(axis.title.x = element_blank(), axis.title.y = element_text(size =
-                                                                      10)) +
-  ylab("Log(mean read depth)") +
+  theme_few() +
+  theme(
+    # axis.title.x = element_blank(),
+    # axis.text.x = element_blank(),
+    axis.title.y = element_text(size = 10),
+    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
+    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
+  ) +
+  ylab("Log (mean read depth)") +
+  xlab("Viral load (copies / ml)") +
   scale_color_manual(values = cols3, labels = c("Pool 1", "Pool 2"))
 
-ggarrange(
-  pool_readcounts,
-  pool_readcounts_norm,
-  pool_readcounts_norm2,
-  pool_readdepths,
-  nrow = 2,
-  ncol = 2,
-  heights = c(1, 2),
-  common.legend = TRUE
-)
-
-#ggsave("/Users/laura/Dropbox/glasgow/github/te_ug_rodents/figures/compare_spike_ins_atcc/pool_compare_viruses.pdf")
-
+```
+
+### Final plot
+```{r}
 ggarrange(
@@
   ncol = 2,
-  common.legend = TRUE
-)
-
-#ggsave("/Users/laura/Dropbox/glasgow/github/te_ug_rodents/figures/compare_spike_ins_atcc/pool_compare_viruses_sum.png",width=10,height=7)
-
-#ggsave("/Users/laura/Dropbox/glasgow/github/te_ug_rodents/figures/manuscript_figures_pdf/FigureS2.pdf",width=10,height=7)
+  common.legend = TRUE,
+  align = "hv"
+)
+
+#ggsave("figures/compare_spike_ins_atcc/pool_compare_viruses_sum.png",width=10,height=7)
+
+#ggsave("figures/manuscript_figures_pdf/FigureS2.pdf",width=10,height=7)
 
@@
 
-rm(list = ls())
-
-library(tidyverse)
-library(ggpubr)
-
 #metadata
@@
   read.csv(
-    "/Users/laura/Dropbox/glasgow/github/te_ug_rodents/data_TE/total_virus_mapped_reads_per_sample_dedup_atcc_ref_20241108.csv",
+    "data_TE/total_virus_mapped_reads_per_sample_dedup_atcc_ref_20241108.csv",
     header = TRUE
@@
   read.csv(
-    "/Users/laura/Dropbox/glasgow/github/te_ug_rodents/data_TE/total_virus_mapped_reads_per_sample_nodedup_atcc_ref_20241108.csv",
+    "data_TE/total_virus_mapped_reads_per_sample_nodedup_atcc_ref_20241108.csv",
     header = TRUE
@@
 
-reads_plot <- reads_viral_all %>%
-  group_by(Background, Sample_id, Viral.load, type) %>%
+reads_plot <- 
+  reads_viral_all |>
+  group_by(Background, Sample_id, Viral.load, type) |>
   summarise(
@@
     diff = viral_reads - ATCC_reads
-  ) %>%
+  ) |>
   unique()
 
-total_reads <- reads_plot %>%
-  filter(Background != "p6") %>%
-  filter(Background != "control") %>%
+```
+
+### Make plots
+```{r, echo=FALSE, warning=FALSE, message=FALSE}
+
+total_reads <- 
+  reads_plot |>
+  filter(Background != "p6") |>
+  filter(Background != "control") |>
   ggplot(aes(
@@
   facet_grid( ~ type, labeller = as_labeller(facet_names)) +
-  theme_bw() +
-  theme(
-    axis.title.x = element_blank(),
-    axis.text.x = element_text(angle = 45, hjust = 1),
-    axis.title.y = element_text(size = 10)
-  ) +
-  scale_x_log10() +
+  theme_few() +
+  theme(
+    axis.title.y = element_text(size = 10),
+    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
+    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
+  ) +
+  scale_x_log10(labels = scales::label_log()) + xlab("Viral load (copies / ml)") +
   ggtitle("Total Reads") +
   ylab("Log(Total Reads)") +
-  scale_color_manual(values = cols2, labels = c("ME_P1", "ME_P2"))
-
-viral_reads <- reads_plot %>%
-  filter(Background != "p6") %>%
-  filter(Background != "control") %>%
+  scale_color_manual(values = cols2, labels = c("P1", "P2"))
+
+viral_reads <- 
+  reads_plot |>
+  filter(Background != "p6") |>
+  filter(Background != "control") |>
   ggplot(aes(
@@
   geom_smooth(aes(group = Background), se = FALSE) +
-  theme_bw() +
   facet_grid( ~ type, labeller = as_labeller(facet_names)) +
-  theme(
-    axis.title.x = element_blank(),
-    axis.text.x = element_text(angle = 45, hjust = 1),
-    axis.title.y = element_text(size = 10)
-  ) +
-  scale_x_log10() +
+  theme_few() +
+  theme(
+    axis.title.y = element_text(size = 10),
+    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
+    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
+  ) +
+  scale_x_log10(labels = scales::label_log()) + xlab("Viral load (copies / ml)") +
   ggtitle("Total Viral Reads") +
   ylab("Log(Total Viral Reads)") +
-  scale_color_manual(values = cols2, labels = c("ME_P1", "ME_P2"))
-
-ATCC_reads <- reads_plot %>%
-  filter(Background != "p6") %>%
-  filter(Background != "control") %>%
+  scale_color_manual(values = cols2, labels = c("P1", "P2"))
+
+ATCC_reads <- 
+  reads_plot |>
+  filter(Background != "p6") |>
+  filter(Background != "control") |>
   ggplot(aes(
@@
   geom_smooth(aes(group = Background), se = FALSE) +
-  theme_bw() +
   facet_grid( ~ type, labeller = as_labeller(facet_names)) +
-  theme(
-    axis.title.x = element_blank(),
-    axis.text.x = element_text(angle = 45, hjust = 1),
-    axis.title.y = element_text(size = 10)
-  ) +
-  scale_x_log10() +
+  theme_few() +
+  theme(
+    axis.title.y = element_text(size = 10),
+    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
+    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
+  ) +
+  scale_x_log10(labels = scales::label_log()) + xlab("Viral load (copies / ml)") +
   ggtitle("Spike In Viral Reads") +
   ylab("Log(Spike In Viral Reads)") +
-  scale_color_manual(values = cols2, labels = c("ME_P1", "ME_P2"))
-
-prop_ATCC <- reads_plot %>%
-  filter(Background != "p6") %>%
-  filter(Background != "control") %>%
+  scale_color_manual(values = cols2, labels = c("P1", "P2"))
+
+prop_ATCC <- reads_plot |>
+  filter(Background != "p6") |>
+  filter(Background != "control") |>
   ggplot(aes(x = Viral.load, y = prop_ATCC, colour = Background)) +
@@
   facet_grid( ~ type, labeller = as_labeller(facet_names)) +
-  theme_bw() +
-  theme(
-    axis.title.x = element_blank(),
-    axis.text.x = element_text(angle = 45, hjust = 1),
-    axis.title.y = element_text(size = 10)
-  ) +
-  scale_x_log10() +
+  theme_few() +
+  theme(
+    axis.title.y = element_text(size = 10),
+    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
+    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
+  ) +
+  scale_x_log10(labels = scales::label_log()) + xlab("Viral load (copies / ml)") +
   ggtitle("Proportion Spike In Viral") +
   ylab("Spike In Viral Reads/Total Reads") +
-  scale_color_manual(values = cols2, labels = c("ME_P1", "ME_P2"))
-
-prop_viral <- reads_plot %>%
-  filter(Background != "p6") %>%
-  filter(Background != "control") %>%
+  scale_color_manual(values = cols2, labels = c("P1", "P2"))
+
+prop_viral <- reads_plot |>
+  filter(Background != "p6") |>
+  filter(Background != "control") |>
   ggplot(aes(x = Viral.load, y = prop_viral, colour = Background)) +
@@
   geom_smooth(aes(group = Background), se = FALSE) +
-  theme_bw() +
   facet_grid( ~ type, labeller = as_labeller(facet_names)) +
-  theme(
-    axis.title.x = element_blank(),
-    axis.text.x = element_text(angle = 45, hjust = 1),
-    axis.title.y = element_text(size = 10)
-  ) +
-  scale_x_log10() +
+  theme_few() +
+  theme(
+    axis.title.y = element_text(size = 10),
+    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
+    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
+  ) +
+  scale_x_log10(labels = scales::label_log()) + xlab("Viral load (copies / ml)") +
   ggtitle("Proportion Viral") +
   ylab("Viral Reads/Total Reads") +
-  scale_color_manual(values = cols2, labels = c("ME_P1", "ME_P2"))
+  scale_color_manual(values = cols2, labels = c("P1", "P2"))
 
@@
 
+```
+
+### ***Final plot***
+```{r}
 ggarrange(total_reads,
@@
           nrow = 3,
-          common.legend = TRUE)
-
-#ggsave("/Users/laura/Dropbox/glasgow/github/te_ug_rodents/figures/compare_spike_ins_atcc/backgrounds_reads.png")
-
-#ggsave("/Users/laura/Dropbox/glasgow/github/te_ug_rodents/figures/manuscript_figures_pdf/FigureS3.pdf")
+          common.legend = TRUE,
+          align = "hv")
+
+#ggsave("figures/compare_spike_ins_atcc/backgrounds_reads.png")
+
+#ggsave("figures/manuscript_figures_pdf/FigureS3.pdf")
 
@@
 
-metadata3 <- metadata %>%
-  select(Sample.ID,
-         Number.of.read.pairs..quality.adaptor.trimmed.,
-         Pool.for.sequencing) %>%
-  rename(Sample_id = Sample.ID) %>%
-  rename(QC_reads = Number.of.read.pairs..quality.adaptor.trimmed.)
-
-reads_viral_dedup3 <-
-  full_join(metadata3, viral_reads_dedup, by = "Sample_id")
-
-reads_viral_nodedup3 <-
-  full_join(metadata3, viral_reads_nodedup, by = "Sample_id")
-
-reads_viral_dedup3 %>%
-  group_by(Pool.for.sequencing) %>%
-  summarise(
-    viral_reads = sum(total_virus_reads),
-    total_reads = sum(QC_reads * 2),
-    prop_viral = viral_reads / total_reads
-  )
-
-polyomics <-
-  read.csv(
-    "/Users/laura/Dropbox/glasgow/github/te_ug_rodents/data_polyomics/total_virus_mapped_reads_per_sample_dedup.csv"
-  )
-
-#read data from polyomics_indexes_stefano document
-total_reads <- data.frame(total_reads = c(5942760, 7714275))
-
-keeps <- c("RNA-Msp-p2", "RNA-Msp-p8")
-
-polyomics_read_prop <- polyomics %>%
-  filter(Sample_id %in% keeps)
-
-polyomics_read_prop2 <- cbind(polyomics_read_prop, total_reads)
-
-polyomics_read_prop2 %>%
-  summarise(prop_viral = total_virus_reads / total_reads)
+# metadata3 <- metadata |>
+#   select(Sample.ID,
+#          Number.of.read.pairs..quality.adaptor.trimmed.,
+#          Pool.for.sequencing) |>
+#   rename(Sample_id = Sample.ID) |>
+#   rename(QC_reads = Number.of.read.pairs..quality.adaptor.trimmed.)
+# 
+# reads_viral_dedup3 <-
+#   full_join(metadata3, viral_reads_dedup, by = "Sample_id")
+# 
+# reads_viral_nodedup3 <-
+#   full_join(metadata3, viral_reads_nodedup, by = "Sample_id")
+# 
+# reads_viral_dedup3 |>
+#   group_by(Pool.for.sequencing) |>
+#   summarise(
+#     viral_reads = sum(total_virus_reads),
+#     total_reads = sum(QC_reads * 2),
+#     prop_viral = viral_reads / total_reads
+#   )
+# 
+# polyomics <-
+#   read.csv(
+#     "data_polyomics/total_virus_mapped_reads_per_sample_dedup.csv"
+#   )
+# 
+# #read data from polyomics_indexes_stefano document
+# total_reads <- data.frame(total_reads = c(5942760, 7714275))
+# 
+# keeps <- c("RNA-Msp-p2", "RNA-Msp-p8")
+# 
+# polyomics_read_prop <- polyomics |>
+#   filter(Sample_id %in% keeps)
+# 
+# polyomics_read_prop2 <- cbind(polyomics_read_prop, total_reads)
+# 
+# polyomics_read_prop2 |>
+#   summarise(prop_viral = total_virus_reads / total_reads)
 
@@
   facet_grid(Background ~ type, labeller = as_labeller(facet_names_bg)) +
+  theme_few() +
+  theme(
+    axis.title.x = element_blank(),
+    axis.title.y = element_text(size = 10),
+    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
+    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
+  ) +
+  scale_color_manual(
+    values = cols,
+    labels =
+      c(
+      "Human adenovirus 40",
+      "Human betaherpesvirus",
+      "Human respiratory syncytial virus",
+      "Influenza B virus",
+      "Mammalian orthoreovirus 3",
+      "Zika virus"
+      )
+  ) +
+  scale_x_log10(labels = scales::label_log()) +
+  ylab("Log(Viral Reads)")
+
+backgrounds_counts_reads_norm <- 
+  counts_reads_norm |>
+  filter(Background != "p6") |>
+  filter(Background != "control") |>
+  ggplot(aes(
+    x = Viral.load,
+    y = log(norm_counts1),
+    colour = virus
+  )) +
+  geom_point() +
+  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
+  facet_grid(Background ~ type, labeller = as_labeller(facet_names_bg)) +
   theme_bw() +
+  theme_few() +
   theme(
     axis.title.x = element_blank(),
-    axis.text.x = element_text(angle = 45, hjust = 1),
     axis.title.y = element_text(size = 10),
-    legend.title = element_blank()
+    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
+    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
   ) +
@@
   ) +
-  scale_x_log10() +
-  ylab("Log(Viral Reads)")
-
-backgrounds_counts_reads_norm <- 
+  scale_x_log10(label = scales::label_log()) +
+  ylab("Log(viral reads/cleaned reads)")
+
+backgrounds_counts_reads_norm2 <- 
   counts_reads_norm |>
@@
     x = Viral.load,
-    y = log(norm_counts1),
+    y = log(norm_counts2),
     colour = virus
@@
   facet_grid(Background ~ type, labeller = as_labeller(facet_names_bg)) +
-  theme_bw() +
+  theme_few() +
   theme(
     axis.title.x = element_blank(),
-    axis.text.x = element_text(angle = 45, hjust = 1),
     axis.title.y = element_text(size = 10),
-    legend.title = element_blank()
+    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
+    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
   ) +
@@
   ) +
-  scale_x_log10() +
-  ylab("Log(viral reads/cleaned reads)")
-
-backgrounds_counts_reads_norm2 <- 
-  counts_reads_norm |>
+  scale_x_log10(labels = scales::label_log()) +
+  ylab("Log(viral reads/cleaned reads/genome length)")
+
+background_read_depths <- 
+  depths_reads |>
   filter(Background != "p6") |>
@@
     x = Viral.load,
-    y = log(norm_counts2),
+    y = log(mean_depth),
     colour = virus
@@
   facet_grid(Background ~ type, labeller = as_labeller(facet_names_bg)) +
-  theme_bw() +
+  theme_few() +
   theme(
     axis.title.x = element_blank(),
-    axis.text.x = element_text(angle = 45, hjust = 1),
+    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
+    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray"),
     axis.title.y = element_text(size = 10),
@@
   ) +
-  scale_x_log10() +
-  ylab("Log(viral reads/cleaned reads/genome length)")
-
-background_read_depths <- 
-  depths_reads |>
-  filter(Background != "p6") |>
-  filter(Background != "control") |>
-  ggplot(aes(
-    x = Viral.load,
-    y = log(mean_depth),
-    colour = virus
-  )) +
-  geom_point() +
-  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
-  facet_grid(Background ~ type, labeller = as_labeller(facet_names_bg)) +
-  theme_bw() +
-  theme(
-    axis.title.x = element_blank(),
-    axis.text.x = element_text(angle = 45, hjust = 1),
-    axis.title.y = element_text(size = 10),
-    legend.title = element_blank()
-  ) +
-  scale_color_manual(
-    values = cols,
-    labels = c(
-      "Human adenovirus 40",
-      "Human betaherpesvirus",
-      "Human respiratory syncytial virus",
-      "Influenza B virus",
-      "Mammalian orthoreovirus 3",
-      "Zika virus"
-    )
-  ) +
-  scale_x_log10() +
+  scale_x_log10(label = scales::label_log()) +
   ylab("Log(mean read depth)")
@@
 
-### Final plot 
+### ***Final plot*** 
 ```{r, warning=FALSE, message=FALSE, echo=FALSE, fig.height=6, fig.height=8}
@@
   ncol = 2,
-  common.legend = TRUE
+  common.legend = TRUE,
+  align = "hv"
 )
@@
 #ggsave("figures/manuscript_figures_pdf/FigureS4.pdf",width=10,height=7)
-
-```
-
-####extras
-
-```{r}
-read_count_raw<-counts_reads_norm |>
-  filter(Background != "p6") |>
-  filter(Background != "control") |>
-  ggplot(aes(x=Viral.load,y=matched,colour = virus))+
-  geom_point()+
-  geom_smooth(aes(group=virus),se=FALSE)+
-  facet_grid(~type,labeller=as_labeller(facet_names))+
-  theme(axis.title.x=element_blank(),axis.text.x=element_text(angle=45,hjust=1),axis.title.y=element_text(size=10))+
-  scale_x_log10()+
-  ylab("Viral Reads")
-
-raw_read_count_norm1<- counts_reads_norm |>
-  filter(Background != "p6") |>
-  filter(Background != "control") |>
-  ggplot(aes(x=Viral.load,y=norm_counts1,colour = virus))+
-  geom_point()+
-  geom_smooth(aes(group=virus),se=FALSE)+
-  facet_grid(~type)+
-  theme(axis.title.x=element_blank(),axis.text.x=element_text(angle=45,hjust=1),axis.title.y=element_text(size=10))+
-  scale_x_log10()+
-  ylab("Viral reads/cleaned reads")
-
-raw_read_count_norm2<- counts_reads_norm |>
-  filter(Background != "p6") |>
-  filter(Background != "control") |>
-  ggplot(aes(x=Viral.load,y=norm_counts2,colour = virus))+
-  geom_point()+
-  geom_smooth(aes(group=virus),se=FALSE)+
-  facet_grid(~type)+
-  theme(axis.title.x=element_blank(),axis.text.x=element_text(angle=45,hjust=1),axis.title.y=element_text(size=10))+
-  scale_x_log10()+
-  ylab("Viral reads/cleaned reads/genome length")
-
-##just genome length
-
-read_count_norm3<- counts_reads_norm |>
-  filter(Background != "p6") |>
-  filter(Background != "control") |>
-  ggplot(aes(x=Viral.load,y=log(norm_counts3),colour = virus))+
-  geom_point()+
-  geom_smooth(aes(group=virus),se=FALSE)+
-  facet_grid(~type,labeller=as_labeller(facet_names))+
-  theme_bw()+
-  theme(axis.title.x=element_blank(),axis.text.x=element_text(angle=45,hjust=1),axis.title.y=element_text(size=10))+
-  scale_x_log10()+
-  scale_color_hue(labels=c("Human adenovirus 40","Human betaherpesvirus","Human respiratory syncytial virus","Influenza B virus","Mammalian orthoreovirus 3","Zika virus"))+
-  ylab("Log(viral reads/genome length)")
-
-#ggsave("figures/compare_spike_ins_atcc/norm_genome_length.pdf")
-
-raw_read_depths<- depths_reads |>
-  filter(Background != "p6") |>
-  filter(Background != "control") |>
-  ggplot(aes(x=Viral.load,y=mean_depth,colour = virus))+
-  geom_point()+
-  geom_smooth(aes(group=virus),se=FALSE)+
-  facet_grid(~type)+
-  theme(axis.title.x=element_blank(),axis.text.x=element_text(angle=45,hjust=1),axis.title.y=element_text(size=10))+
-  scale_x_log10()+
-  ylab("Mean read depth")
-
-ggarrange(read_count_raw,raw_read_count_norm1,raw_read_count_norm2,raw_read_depths,nrow=2,ncol=2,common.legend=TRUE)
-
-#ggsave("figures/compare_spike_ins_atcc/read_count_depth.pdf")
-
-####split by viruses####
-
-counts_split<-counts_reads_norm |>
-  filter(Background != "p6") |>
-  filter(Background != "control") |>
-  ggplot(aes(x=virus,y=log(matched),colour=type))+
-  geom_boxplot()+
-  facet_grid(~as.character(Viral.load))+
-  ylab("Log(Viral Reads)")+
-  theme(axis.title.x=element_blank(),axis.text.x=element_blank(),axis.title.y=element_text(size=10))
-
-counts_norm_split<-counts_reads_norm |>
-  filter(Background != "p6") |>
-  filter(Background != "control") |>
-  ggplot(aes(x=virus,y=log(norm_counts1),colour=type))+
-  geom_boxplot()+
-  facet_grid(~as.character(Viral.load))+
-  ylab("Log(viral reads/cleaned reads)")+
-  theme(axis.title.x=element_blank(),axis.text.x=element_blank(),axis.title.y=element_text(size=10))
-
-counts_norm2_split<-counts_reads_norm |>
-  filter(Background != "p6") |>
-  filter(Background != "control") |>
-  ggplot(aes(x=virus,y=log(norm_counts2),colour=type))+
-  geom_boxplot()+
-  facet_grid(~as.character(Viral.load))+
-  ylab("Log(viral reads/cleaned reads/genome length)")+
-  theme(axis.title.x=element_blank(),axis.text.x=element_text(angle=45,hjust=1),axis.title.y=element_text(size=10))
-
-depth_split<-depths_pool |>
-  filter(Background != "p6") |>
-  filter(Background != "control") |>
-  ggplot(aes(x=virus,y=log(mean_depth),colour=type))+
-  geom_boxplot()+
-  facet_grid(~as.character(Viral.load))+
-  theme(axis.title.x=element_blank(),axis.text.x=element_text(angle=45,hjust=1),axis.title.y=element_text(size=10))+
-  ylab("Log(mean read depth)")+
-  theme(axis.title.x=element_blank(),axis.text.x=element_text(angle=45,hjust=1),axis.title.y=element_text(size=10))
-
-ggarrange(counts_split,counts_norm_split,counts_norm2_split,depth_split,nrow=2,ncol=2,heights=c(1,2),common.legend = TRUE)
-
-#ggsave("figures/compare_spike_ins_atcc/compare_viruses_split.pdf")
 
@@
 
-depths_reads <- left_join(depths_bt_all, metadata2, by = "Sample_id") %>%
+depths_reads <- left_join(depths_bt_all, metadata2, by = "Sample_id") |>
   mutate(genome_structure = case_when((
@@
 
-depths_reads_sub <- depths_reads %>%
-  filter(Background != "p6") %>%
-  filter(Background != "control") %>%
-  filter(type == "dedup_TE") %>%
-  mutate(log_depth = log(mean_depth)) %>%
+depths_reads_sub <- depths_reads |>
+  filter(Background != "p6") |>
+  filter(Background != "control") |>
+  filter(type == "dedup_TE") |>
+  mutate(log_depth = log(mean_depth)) |>
   mutate(log_viral_load = log10(Viral.load))
@@
 
-virus <- (unique(depths_reads_sub$virus)) %>%
-  rep(., each = 2) %>%
+virus <- (unique(depths_reads_sub$virus)) |>
+  rep(., each = 2) |>
   data.frame()
 
-list_models <- depths_reads_sub %>%
-  group_split(virus) %>%
+list_models <- depths_reads_sub |>
+  group_split(virus) |>
   map( ~ lm(log_depth ~ log_viral_load, data = .))
 
-lm_tidy <- map(list_models, broom::tidy) %>%
-  do.call(rbind.data.frame, .) %>%
-  cbind(virus, .) %>%
-  rename(virus = ".") %>%
-  select(virus, term, estimate) %>%
+lm_tidy <- map(list_models, broom::tidy) |>
+  do.call(rbind.data.frame, .) |>
+  cbind(virus, .) |>
+  rename(virus = ".") |>
+  select(virus, term, estimate) |>
   pivot_wider(names_from = "term", values_from = "estimate")
 
-lm_summary <- depths_reads_sub %>%
-  group_by(virus) %>%
+lm_summary <- depths_reads_sub |>
+  group_by(virus) |>
   summarise(
@@
 
-#ggsave("/Users/laura/Dropbox/glasgow/github/te_ug_rodents/figures/compare_spike_ins_atcc/model_readdepth_dedup.png")
-
-#ggsave("/Users/laura/Dropbox/glasgow/github/te_ug_rodents/figures/manuscript_figures_pdf/FigureS5.pdf")
+#ggsave("figures/compare_spike_ins_atcc/model_readdepth_dedup.png")
+
+#ggsave("figures/manuscript_figures_pdf/FigureS5.pdf")
 
@@
                     "100000" = "1e+05",
-                    "p2" = "ME_P1",
-                    "p8"= "ME_P2")
+                    "p2" = "P1",
+                    "p8"= "P2")
 
@@
   scale_y_log10() +
-  scale_fill_manual(values = cols2, labels = c("ME_P1", "ME_P2"))
+  scale_fill_manual(values = cols2, labels = c("P1", "P2"))
 
@@
   ylab("Log(Coverage)") +
-  ggtitle("Influenza B virus (Background ME_1 deduplicated)") +
+  ggtitle("Influenza B virus (Background 1 deduplicated)") +
   guides(fill = guide_legend(title = "Viral load")) +
@@
   ylab("Log(Coverage)") +
-  ggtitle("Influenza B virus (Background ME_2 deduplicated)") +
+  ggtitle("Influenza B virus (Background 2 deduplicated)") +
   guides(fill = guide_legend(title = "Viral load")) +
@@
 ylab("Log(Coverage)") +
-ggtitle("Mammalian orthoreovirus 3 (Background ME_1 deduplicated)") +
+ggtitle("Mammalian orthoreovirus 3 (Background 1 deduplicated)") +
 guides(fill = guide_legend(title = "Viral load")) +
@@
 ylab("Log(Coverage)") +
-ggtitle("Mammalian orthoreovirus 3 (Background ME_2 deduplicated)") +
+ggtitle("Mammalian orthoreovirus 3 (Background 2 deduplicated)") +
 guides(fill = guide_legend(title = "Viral load")) +

### 2025-01-21 manuscript_figures.Rmd
updating figures

@@
   ) +
-  ylab("Log(Viral Reads)")
+  ylab("Log(Viral Reads)") +
+  xlab("Viral load")
 
@@
   ) +
-  ylab("Log(viral reads/cleaned reads/genome length)")
+  ylab("Log(viral reads/cleaned reads/genome length)") +
+  xlab("Viral load")
 
@@
 
-ggsave("figures/compare_spike_ins_atcc/log_read_count_depth_2panels_dedup.png",
-       width=10,
-       height=7)
-
-ggsave("figures/manuscript_figures_pdf/Figure1.pdf",
-       width=10,
-       height=7)
+# ggsave("figures/compare_spike_ins_atcc/log_read_count_depth_2panels_dedup.png",
+#        width=10,
+#        height=7)
+
+ggsave("figures/manuscript_figure_2025/PDF/Figure_1.pdf",
+       width=6,
+       height=8)
+
+ggsave("figures/manuscript_figure_2025/PNG/Figure_1.png",
+       width=6,
+       height=8)
 
@@
 
+### Figure S5 - linear model
+
+#### Read-in data
+```{r}
+#Linear model for TE data - separated by virus
+#November 2024
+
+#viral read depth
+#use bowtie2 data
+
+library(tidyverse)
+library(broom)
+library(boot)
+
+####read counts/normalised read counts####
+
+#metadata
+
+metadata <-
+  read.csv(
+    "metadata/sampleIDs_TESpikeIn.csv",
+    header = TRUE
+  )
+
+metadata2 <- metadata |>
+  select(Sample.ID, Number.of.read.pairs..quality.adaptor.trimmed.) |>
+  rename(Sample_id = Sample.ID) |>
+  rename(QC_reads = Number.of.read.pairs..quality.adaptor.trimmed.)
+
+#import and combine read count files
+#deduplicated and non-deduplicated
+
+counts_dedup_bt <-
+  read.table(
+    "data_TE/TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_dedup_atcc_ref.tsv",
+    sep = "\t",
+    header = TRUE
+  )
+
+counts_nodedup_bt <-
+  read.table(
+    "data_TE/TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv",
+    sep = "\t",
+    header = TRUE
+  )
+
+counts_bt_all <- rbind(counts_dedup_bt, counts_nodedup_bt)
+
+counts_reads <- left_join(counts_bt_all, metadata2, by = "Sample_id")
+
+####read depths####
+
+#import and combine read depth files
+
+depth_dedup_bt <-
+  read.table(
+    "data_TE/TE_sequencing_experiment_readdepth_per_sample_and_virus_bowtie2_dedup_atcc_ref.tsv",
+    sep = "\t",
+    header = TRUE
+  )
+
+depth_nodedup_bt <-
+  read.table(
+    "data_TE/TE_sequencing_experiment_readdepth_per_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv",
+    sep = "\t",
+    header = TRUE
+  )
+
+```
+
+#### Format data
+```{r}
+
+depths_bt_all <- rbind(depth_dedup_bt, depth_nodedup_bt)
+
+depths_reads <- left_join(depths_bt_all, metadata2, by = "Sample_id") |>
+  mutate(genome_structure = case_when((
+    virus == "Human_adenovirus_40" |
+      virus == "Human_betaherpesvirus"
+  ) ~ "DNA",
+  ,
+  .default = "RNA"
+  ))
+
+####linear model for spike in viruses - mean read depth####
+
+depths_reads_sub <- depths_reads |>
+  filter(Background != "p6") |>
+  filter(Background != "control") |>
+  filter(type == "dedup_TE") |>
+  mutate(log_depth = log(mean_depth)) |>
+  mutate(log_viral_load = log10(Viral.load))
+
+#inspect data
+hist(depths_reads_sub$mean_depth)
+hist(depths_reads_sub$Viral.load)
+hist(depths_reads_sub$log_depth)
+hist(depths_reads_sub$log_viral_load)
+
+summary(depths_reads_sub$log_depth)
+summary(depths_reads_sub$log_viral_load)
+
+```
+
+#### model
+```{r}
+#all viruses together
+
+lm1 <-
+  glm(log_depth ~ log_viral_load * virus,
+      data = depths_reads_sub,
+      family = "gaussian")
+
+summary(lm1)
+
+glm.diag.plots(lm1)
+
+pred <- predict(lm1, type = "response")
+
+rsq <- function (x, y) {
+  cor(x, y) ^ 2
+}
+
+rsq(depths_reads_sub$log_depth, pred)
+
+##split by viruses
+
+cols <- c("#4477AA",
+          "#66CCEE",
+          "#228833",
+          "#CCBB44",
+          "#EE6677",
+          "#AA3377")
+
+facet_names <- c(
+  "Human_adenovirus_40" = "Human adenovirus 40",
+  "Human_betaherpesvirus" = "Human betaherpesvirus",
+  "Human_respiratory_syncytial_virus" = "Human respiratory syncytial virus",
+  "Influenza_B_virus" = "Influenza B virus",
+  "Mammalian_orthoreovirus3" = "Mammalian orthoreovirus 3",
+  "Zika_virus" = "Zika virus"
+)
+
+virus <- (unique(depths_reads_sub$virus)) |>
+  rep(each = 2) |>
+  data.frame() |>
+  rename(virus = rep..unique.depths_reads_sub.virus....each...2.)
+
+list_models <- 
+  depths_reads_sub |>
+  group_split(virus) |>
+  map( ~ lm(log_depth ~ log_viral_load, data = .))
+
+lm_tidy <- 
+  map(list_models, broom::tidy) |>
+  do.call(rbind.data.frame, .) |>
+  cbind(virus, .) |>
+  rename(virus = ".") |>
+  select(virus, term, estimate) |>
+  pivot_wider(names_from = "term", values_from = "estimate")
+
+lm_summary <- depths_reads_sub |>
+  group_by(virus) |>
+  summarise(
+    Intercept = lm(log_depth ~ log_viral_load)$coefficients[1],
+    Coeff_x1 = lm(log_depth ~ log_viral_load)$coefficients[2],
+    R2 = summary(lm(log_depth ~ log_viral_load))$r.squared,
+    pvalue = summary(lm(log_depth ~ log_viral_load))$coefficients["log_viral_load", 4]
+  )
+
+lm_combined <- left_join(lm_tidy, lm_summary, by = "virus")
+
+ggplot() +
+  geom_point(data = depths_reads_sub,
+             mapping = aes(log_viral_load, log_depth, color = virus)) +
+  geom_abline(data = lm_combined, aes(intercept = `(Intercept)`, slope = log_viral_load)) +
+  geom_text(data = lm_combined,
+            colour = "black",
+            aes(
+              x = 4.7,
+              y = 3,
+              label = round(Coeff_x1, digits = 2)
+            )) +
+  geom_text(data = lm_combined,
+            colour = "black",
+            aes(x = 4, y = 3, label = "Slope =")) +
+  geom_text(data = lm_combined,
+            colour = "black",
+            aes(
+              x = 4.7,
+              y = 2,
+              label = round(R2, digits = 2)
+            )) +
+  geom_text(data = lm_combined,
+            colour = "black",
+            aes(x = 4, y = 2, label = "R2 =")) +
+  facet_wrap( ~ virus, labeller = as_labeller(facet_names)) +
+  theme_bw() +
+  xlab("Log(Viral Load)") +
+  ylab("Log(Mean Read Depth)") +
+  scale_color_manual(values = cols) +
+  guides(color = FALSE)
+
+#ggsave("figures/compare_spike_ins_atcc/model_readdepth_dedup.png")
+
+#ggsave("figures/manuscript_figures_pdf/FigureS5.pdf")
+
+```
+
+### --------------------
+
 ## >>> Figure 2: plots of TE experiment data - genome coverage & per site coverage
@@
 ggsave(
-  "figures/manuscript_figures_pdf/Figure_2.pdf",
+  "figures/manuscript_figure_2025/PDF/Figure_2.pdf",
   width = 15,
@@
 
+ggsave(
+  "figures/manuscript_figure_2025/PNG/Figure_2.png",
+  width = 15,
+  height = 6
+)
+
 ```
@@
 ### --------------------
-
-### Figure S5 - linear model
-```{r}
-#Linear model for TE data - separated by virus
-#November 2024
-
-#viral read depth
-#use bowtie2 data
-
-rm(list = ls())
-
-library(tidyverse)
-library(broom)
-
-####read counts/normalised read counts####
-
-#metadata
-
-metadata <-
-  read.csv(
-    "metadata/sampleIDs_TESpikeIn.csv",
-    header = TRUE
-  )
-
-metadata2 <- metadata |>
-  select(Sample.ID, Number.of.read.pairs..quality.adaptor.trimmed.) |>
-  rename(Sample_id = Sample.ID) |>
-  rename(QC_reads = Number.of.read.pairs..quality.adaptor.trimmed.)
-
-#import and combine read count files
-#deduplicated and non-deduplicated
-
-counts_dedup_bt <-
-  read.table(
-    "data_TE/TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_dedup_atcc_ref.tsv",
-    sep = "\t",
-    header = TRUE
-  )
-
-counts_nodedup_bt <-
-  read.table(
-    "data_TE/TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv",
-    sep = "\t",
-    header = TRUE
-  )
-
-counts_bt_all <- rbind(counts_dedup_bt, counts_nodedup_bt)
-
-counts_reads <- left_join(counts_bt_all, metadata2, by = "Sample_id")
-
-####read depths####
-
-#import and combine read depth files
-
-depth_dedup_bt <-
-  read.table(
-    "data_TE/TE_sequencing_experiment_readdepth_per_sample_and_virus_bowtie2_dedup_atcc_ref.tsv",
-    sep = "\t",
-    header = TRUE
-  )
-
-depth_nodedup_bt <-
-  read.table(
-    "data_TE/TE_sequencing_experiment_readdepth_per_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv",
-    sep = "\t",
-    header = TRUE
-  )
-
-depths_bt_all <- rbind(depth_dedup_bt, depth_nodedup_bt)
-
-depths_reads <- left_join(depths_bt_all, metadata2, by = "Sample_id") |>
-  mutate(genome_structure = case_when((
-    virus == "Human_adenovirus_40" |
-      virus == "Human_betaherpesvirus"
-  ) ~ "DNA",
-  ,
-  .default = "RNA"
-  ))
-
-####linear model for spike in viruses - mean read depth####
-
-depths_reads_sub <- depths_reads |>
-  filter(Background != "p6") |>
-  filter(Background != "control") |>
-  filter(type == "dedup_TE") |>
-  mutate(log_depth = log(mean_depth)) |>
-  mutate(log_viral_load = log10(Viral.load))
-
-#inspect data
-hist(depths_reads_sub$mean_depth)
-hist(depths_reads_sub$Viral.load)
-hist(depths_reads_sub$log_depth)
-hist(depths_reads_sub$log_viral_load)
-
-summary(depths_reads_sub$log_depth)
-summary(depths_reads_sub$log_viral_load)
-
-#all viruses together
-
-lm1 <-
-  glm(log_depth ~ log_viral_load * virus,
-      data = depths_reads_sub,
-      family = "gaussian")
-
-summary(lm1)
-
-glm.diag.plots(lm1)
-
-pred <- predict(lm1, type = "response")
-
-rsq <- function (x, y)
-  cor(x, y) ^ 2
-
-rsq(depths_reads_sub$log_depth, pred)
-
-##split by viruses
-
-cols <- c("#4477AA",
-          "#66CCEE",
-          "#228833",
-          "#CCBB44",
-          "#EE6677",
-          "#AA3377")
-
-facet_names <- c(
-  "Human_adenovirus_40" = "Human adenovirus 40",
-  "Human_betaherpesvirus" = "Human betaherpesvirus",
-  "Human_respiratory_syncytial_virus" = "Human respiratory syncytial virus",
-  "Influenza_B_virus" = "Influenza B virus",
-  "Mammalian_orthoreovirus3" = "Mammalian orthoreovirus 3",
-  "Zika_virus" = "Zika virus"
-)
-
-virus <- (unique(depths_reads_sub$virus)) |>
-  rep(., each = 2) |>
-  data.frame()
-
-list_models <- depths_reads_sub |>
-  group_split(virus) |>
-  map( ~ lm(log_depth ~ log_viral_load, data = .))
-
-lm_tidy <- map(list_models, broom::tidy) |>
-  do.call(rbind.data.frame, .) |>
-  cbind(virus, .) |>
-  rename(virus = ".") |>
-  select(virus, term, estimate) |>
-  pivot_wider(names_from = "term", values_from = "estimate")
-
-lm_summary <- depths_reads_sub |>
-  group_by(virus) |>
-  summarise(
-    Intercept = lm(log_depth ~ log_viral_load)$coefficients[1],
-    Coeff_x1 = lm(log_depth ~ log_viral_load)$coefficients[2],
-    R2 = summary(lm(log_depth ~ log_viral_load))$r.squared,
-    pvalue = summary(lm(log_depth ~ log_viral_load))$coefficients["log_viral_load", 4]
-  )
-
-lm_combined <- left_join(lm_tidy, lm_summary, by = "virus")
-
-ggplot() +
-  geom_point(data = depths_reads_sub,
-             mapping = aes(log_viral_load, log_depth, color = virus)) +
-  geom_abline(data = lm_combined, aes(intercept = `(Intercept)`, slope = log_viral_load)) +
-  geom_text(data = lm_combined,
-            colour = "black",
-            aes(
-              x = 4.7,
-              y = 3,
-              label = round(Coeff_x1, digits = 2)
-            )) +
-  geom_text(data = lm_combined,
-            colour = "black",
-            aes(x = 4, y = 3, label = "Slope =")) +
-  geom_text(data = lm_combined,
-            colour = "black",
-            aes(
-              x = 4.7,
-              y = 2,
-              label = round(R2, digits = 2)
-            )) +
-  geom_text(data = lm_combined,
-            colour = "black",
-            aes(x = 4, y = 2, label = "R2 =")) +
-  facet_wrap( ~ virus, labeller = as_labeller(facet_names)) +
-  theme_bw() +
-  xlab("Log(Viral Load)") +
-  ylab("Log(Mean Read Depth)") +
-  scale_color_manual(values = cols) +
-  guides(color = FALSE)
-
-#ggsave("figures/compare_spike_ins_atcc/model_readdepth_dedup.png")
-
-#ggsave("figures/manuscript_figures_pdf/FigureS5.pdf")
-
-```
 

### 2025-01-21 manuscript_figures.Rmd
updating figures

@@
 
-## >>>> Figure 1: plots of TE experiment data - read counts and read depths
+### Figure 1 - linear model
+
+#### Read-in data
+```{r}
+#Linear model for TE data - separated by virus
+#November 2024
+
+#viral read depth
+#use bowtie2 data
+
+library(tidyverse)
+library(broom)
+library(boot)
+
+####read counts/normalised read counts####
+
+#metadata
+
+metadata <-
+  read.csv(
+    "metadata/sampleIDs_TESpikeIn.csv",
+    header = TRUE
+  )
+
+metadata2 <- metadata |>
+  select(Sample.ID, Number.of.read.pairs..quality.adaptor.trimmed.) |>
+  rename(Sample_id = Sample.ID) |>
+  rename(QC_reads = Number.of.read.pairs..quality.adaptor.trimmed.)
+
+#import and combine read count files
+#deduplicated and non-deduplicated
+
+counts_dedup_bt <-
+  read.table(
+    "data_TE/TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_dedup_atcc_ref.tsv",
+    sep = "\t",
+    header = TRUE
+  )
+
+counts_nodedup_bt <-
+  read.table(
+    "data_TE/TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv",
+    sep = "\t",
+    header = TRUE
+  )
+
+counts_bt_all <- rbind(counts_dedup_bt, counts_nodedup_bt)
+
+counts_reads <- left_join(counts_bt_all, metadata2, by = "Sample_id")
+
+####read depths####
+
+#import and combine read depth files
+
+depth_dedup_bt <-
+  read.table(
+    "data_TE/TE_sequencing_experiment_readdepth_per_sample_and_virus_bowtie2_dedup_atcc_ref.tsv",
+    sep = "\t",
+    header = TRUE
+  )
+
+depth_nodedup_bt <-
+  read.table(
+    "data_TE/TE_sequencing_experiment_readdepth_per_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv",
+    sep = "\t",
+    header = TRUE
+  )
+
+```
+
+#### Format data
+```{r}
+
+depths_bt_all <- rbind(depth_dedup_bt, depth_nodedup_bt)
+
+depths_reads <- left_join(depths_bt_all, metadata2, by = "Sample_id") |>
+  mutate(genome_structure = case_when((
+    virus == "Human_adenovirus_40" |
+      virus == "Human_betaherpesvirus"
+  ) ~ "DNA",
+  ,
+  .default = "RNA"
+  ))
+
+####linear model for spike in viruses - mean read depth####
+
+depths_reads_sub <- depths_reads |>
+  filter(Background != "p6") |>
+  filter(Background != "control") |>
+  filter(type == "dedup_TE") |>
+  mutate(log_depth = log10(mean_depth)) |>
+  mutate(log_viral_load = log10(Viral.load))
+
+#inspect data
+hist(depths_reads_sub$mean_depth)
+hist(depths_reads_sub$Viral.load)
+hist(depths_reads_sub$log_depth)
+hist(depths_reads_sub$log_viral_load)
+
+summary(depths_reads_sub$log_depth)
+summary(depths_reads_sub$log_viral_load)
+
+```
+
+#### model
+```{r}
+#all viruses together
+
+lm1 <-
+  glm(log_depth ~ log_viral_load * virus,
+      data = depths_reads_sub,
+      family = "gaussian")
+
+summary(lm1)
+
+glm.diag.plots(lm1)
+
+pred <- predict(lm1, type = "response")
+
+rsq <- function (x, y) {
+  cor(x, y) ^ 2
+}
+
+rsq(depths_reads_sub$log_depth, pred)
+
+##split by viruses
+
+cols <- c("#4477AA",
+          "#66CCEE",
+          "#228833",
+          "#CCBB44",
+          "#EE6677",
+          "#AA3377")
+
+facet_names <- c(
+  "Human_adenovirus_40" = "Human adenovirus 40",
+  "Human_betaherpesvirus" = "Human betaherpesvirus",
+  "Human_respiratory_syncytial_virus" = "Human respiratory syncytial virus",
+  "Influenza_B_virus" = "Influenza B virus",
+  "Mammalian_orthoreovirus3" = "Mammalian orthoreovirus 3",
+  "Zika_virus" = "Zika virus"
+)
+
+virus <- (unique(depths_reads_sub$virus)) %>%
+  rep(., each = 2) %>%
+  data.frame()
+
+list_models <- 
+  depths_reads_sub |>
+  group_split(virus) |>
+  map( ~ lm(log_depth ~ log_viral_load, data = .))
+
+lm_tidy <- 
+  map(list_models, broom::tidy) %>%
+  do.call(rbind.data.frame, .) %>%
+  cbind(virus, .) %>%
+  rename(virus = ".") %>%
+  select(virus, term, estimate) %>%
+  pivot_wider(names_from = "term", values_from = "estimate")
+
+lm_summary <- depths_reads_sub |>
+  group_by(virus) |>
+  summarise(
+    Intercept = lm(log_depth ~ log_viral_load)$coefficients[1],
+    Coeff_x1 = lm(log_depth ~ log_viral_load)$coefficients[2],
+    R2 = summary(lm(log_depth ~ log_viral_load))$r.squared,
+    pvalue = summary(lm(log_depth ~ log_viral_load))$coefficients["log_viral_load", 4]
+  )
+
+lm_combined <- left_join(lm_tidy, lm_summary, by = "virus")
+
+```
+
+#### Make plot
+```{r}
+ggplot() +
+  geom_point(data = depths_reads_sub,
+             mapping = aes(log_viral_load, log_depth, color = virus)) +
+  geom_abline(data = lm_combined, aes(intercept = `(Intercept)`, slope = log_viral_load), linetype="dotted") +
+  geom_text(data = lm_combined,
+            colour = "black",
+            aes(
+              x = 4.7,
+              y = 3,
+              label = round(Coeff_x1, digits = 2)),
+            size = 3) +
+  geom_text(data = lm_combined,
+            colour = "black",
+            aes(x = 4, 
+                y = 3, 
+                label = "Slope ="),
+            size = 3) +
+  geom_text(data = lm_combined,
+            colour = "black",
+            aes(
+              x = 4.7,
+              y = 2,
+              label = round(R2, digits = 2)),
+            size = 3) +
+  geom_text(data = lm_combined,
+            colour = "black",
+            aes(x = 4, 
+                y = 2, 
+                label = "R2 ="),
+            size = 3) +
+  facet_wrap( ~ virus, labeller = as_labeller(facet_names)) +
+  theme_few() +
+  theme(
+    axis.title.x = element_text(size = 10),
+    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
+    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray"),
+    axis.title.y = element_text(size = 10),
+    legend.title = element_blank()
+  ) +
+  xlab("log10(Viral Load)") +
+  ylab("log10(Mean Read Depth)") +
+  scale_color_manual(values = cols) +
+  guides(color = FALSE)
+
+```
+
+#### Final plot
+```{r}
+#ggsave("figures/compare_spike_ins_atcc/model_readdepth_dedup.png")
+
+ggsave("figures/manuscript_figure_2025/PDF/Figure_1.pdf",
+       width = 7.5,
+       height = 4)
+
+ggsave("figures/manuscript_figure_2025/PNG/Figure_1.png",
+       width = 7.5,
+       height = 4)
+
+```
+
+### --------------------
+
+## >>>> now redundant? - Figure 1: plots of TE experiment data - read counts and read depths
 
@@
     x = Viral.load,
-    y = log(matched),
+    y = log10(matched),
     colour = virus
@@
   ) +
-  scale_x_log10() +
   scale_color_manual(
@@
   ) +
-  ylab("Log(Viral Reads)")
+  ylab("log10(Viral Reads)") +
+  xlab("Log (Viral load)") + 
+  scale_x_log10(labels = scales::trans_format("log10", scales::label_math()))
 
@@
     x = Viral.load,
-    y = log(norm_counts1),
+    y = log10(norm_counts1),
     colour = virus
@@
   ) +
-  ylab("Log(viral reads/cleaned reads)")
+  ylab("log10(viral reads/cleaned reads)") +
+  xlab("Log (Viral load)") + 
+  scale_x_log10(labels = scales::trans_format("log10", scales::label_math()))
 
@@
 
-read_count_norm2 <- counts_reads_norm |>
-  filter(Background != "p6") |>
-  filter(Background != "control") |>
-  ggplot(aes(
-    x = Viral.load,
-    y = log(norm_counts2),
+read_count_norm2 <- 
+  counts_reads_norm |>
+  filter(Background != "p6") |>
+  filter(Background != "control") |>
+  ggplot(aes(
+    x = (Viral.load),
+    y = log10(norm_counts2),
     colour = virus
@@
   ) +
-  ylab("Log(viral reads/cleaned reads/genome length)")
+  ylab("Log (viral reads/cleaned reads/genome length)") +
+  xlab("Log (Viral load)") + 
+  scale_x_log10(labels = scales::trans_format("log10", scales::label_math()))
 
@@
   geom_point() +
-  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
+  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE, method = "lm") +
   theme_few() +
@@
   ) +
-  ylab("Log(Viral Reads)") +
-  xlab("Viral load")
+  ylab("Log (Viral Reads)") +
+  xlab("Log (Viral load)") + 
+  scale_x_log10(labels = scales::trans_format("log10", scales::label_math()))
 
@@
     x = Viral.load,
-    y = log(norm_counts2),
+    y = log10(norm_counts2),
     colour = virus
@@
   ) +
-  ylab("Log(viral reads/cleaned reads/genome length)") +
-  xlab("Viral load")
-
+  ylab("Log (viral reads/cleaned reads/genome length)") +
+  xlab("Log (Viral load)") + 
+  scale_x_log10(labels = scales::trans_format("log10", scales::label_math()))
 ```
@@
 ### ***Final plot***
-```{r, warning=FALSE, message=FALSE, echo=FALSE, fig.height=6, fig.height=8}
+```{r, warning=FALSE, message=FALSE, echo=FALSE, fig.width=8, fig.height=10}
 #plot just the two lefthand panels from the original figure
@@
 
-ggsave("figures/manuscript_figure_2025/PDF/Figure_1.pdf",
-       width=6,
+ggsave("figures/manuscript_figure_2025/PDF/Figure_S1.pdf",
+       width=7,
        height=8)
 
-ggsave("figures/manuscript_figure_2025/PNG/Figure_1.png",
-       width=6,
+ggsave("figures/manuscript_figure_2025/PNG/Figure_S1.png",
+       width=7,
        height=8)
@@
     x = Viral.load,
-    y = log(mean_depth),
+    y = log10(mean_depth),
     colour = virus
@@
   ) +
-  scale_x_log10() +
-  ylab("Log(mean read depth)")
+  ylab("Log (mean read depth)") +
+  xlab("Log (Viral load)") + 
+  scale_x_log10(labels = scales::trans_format("log10", scales::label_math()))
 
@@
 ### ***Final plot***
-```{r}
+```{r, warning = FALSE, echo = FALSE, message = FALSE}
 
@@
 
+ggsave("figures/manuscript_figure_2025/PDF/Figure_S1.pdf",
+       width = 12,
+       height = 8)
+
+ggsave("figures/manuscript_figure_2025/PNG/Figure_S1.png",
+       width = 12,
+       height = 8)
+
 ```
@@
 
-### Reaed-in data / Make plots
+### Read-in data / Make plots
 ```{r}
@@
 
-pool_readcounts <- 
-  counts_pool |>
-  filter(Background != "p6") |>
-  filter(Background != "control") |>
-  ggplot(aes(x = virus, y = log(matched))) +
-  geom_boxplot() +
-  facet_grid(Pool ~ type) +
-  theme_few() +
-  theme(
-    # axis.title.x = element_blank(),
-    # axis.text.x = element_blank(),
-    axis.title.y = element_text(size = 10),
-    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
-    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
-  ) +
-  ylab("Log (Viral Reads)") + xlab("Viral load (copies / ml)") +
-  scale_color_manual(values = cols3, labels = c("Pool 1", "Pool 2"))
-
 pool_counts_sum <- 
@@
     x = as.character(Viral.load),
-    y = log(matched),
+    y = log10(matched),
     colour = Pool
@@
   theme(
+    axis.title.x = element_blank(),
+    # axis.text.x = element_blank(),
+    legend.title = element_blank(),
+    axis.title.y = element_text(size = 10),
+    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
+    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
+  ) +
+  ylab("Log (Viral Reads)") + xlab("Log (Viral load)") +
+  scale_color_manual(values = cols3, labels = c("Pool 1", "Pool 2")) + 
+  scale_x_discrete(labels = c(bquote(10^{2}), bquote(10^{3}), bquote(10^{5})))
+
+# pool_readcounts_norm <- 
+#   counts_pool |>
+#   filter(Background != "p6") |>
+#   filter(Background != "control") |>
+#   ggplot(aes(x = virus, y = log10(norm_counts1))) +
+#   geom_boxplot() +
+#   facet_grid(Pool ~ type) +
+#   theme_few() +
+#   theme(
 #     axis.title.x = element_blank(),
-    # axis.text.x = element_blank(),
-    axis.title.y = element_text(size = 10),
-    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
-    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
-  ) +
-  ylab("Log (Viral Reads)") + xlab("Viral load (copies / ml)") +
-  scale_color_manual(values = cols3, labels = c("Pool 1", "Pool 2"))
-
-pool_readcounts_norm <- 
-  counts_pool |>
-  filter(Background != "p6") |>
-  filter(Background != "control") |>
-  ggplot(aes(x = virus, y = log(norm_counts1))) +
-  geom_boxplot() +
-  facet_grid(Pool ~ type) +
-  theme_few() +
-  theme(
-    # axis.title.x = element_blank(),
-    # axis.text.x = element_blank(),
-    axis.title.y = element_text(size = 10),
-    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
-    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
-  ) +
-  ylab("Log(viral reads/cleaned reads)") +  xlab("Viral load (copies / ml)")
+#     # axis.text.x = element_blank(),
+#     axis.title.y = element_text(size = 10),
+#     panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
+#     panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
+#   ) +
+#   ylab("log10(viral reads/cleaned reads)")
 
@@
   theme(
+    axis.title.x = element_blank(),
+    # axis.text.x = element_blank(),
+    legend.title = element_blank(),
+    axis.title.y = element_text(size = 10),
+    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
+    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray") 
+  ) +
+  ylab("log10(viral reads/cleaned reads)") + xlab("Log (Viral load)") +
+  scale_color_manual(values = cols3, labels = c("Pool 1", "Pool 2")) + 
+    scale_x_discrete(labels = c(bquote(10^{2}), bquote(10^{3}), bquote(10^{5})))
+
+# pool_readcounts_norm2 <- 
+#   counts_pool |>
+#   filter(Background != "p6") |>
+#   filter(Background != "control") |>
+#   ggplot(aes(x = virus, y = log10(norm_counts2))) +
+#   geom_boxplot() +
+#   facet_grid(Pool ~ type) +
+#   theme_few() +
+#   theme(
 #     axis.title.x = element_blank(),
-    # axis.text.x = element_blank(),
-    axis.title.y = element_text(size = 10),
-    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
-    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
-  ) +
-  ylab("Log(viral reads/cleaned reads)") +   xlab("Viral load (copies / ml)") +
-  scale_color_manual(values = cols3, labels = c("Pool 1", "Pool 2"))
-
-pool_readcounts_norm2 <- 
-  counts_pool |>
-  filter(Background != "p6") |>
-  filter(Background != "control") |>
-  ggplot(aes(x = virus, y = log(norm_counts2))) +
-  geom_boxplot() +
-  facet_grid(Pool ~ type) +
-  theme_few() +
-  theme(
-    # axis.title.x = element_blank(),
-    # axis.text.x = element_blank(),
-    axis.title.y = element_text(size = 10),
-    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
-    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
-  ) +
-  ylab("Log(viral reads/cleaned reads/genome length)") + xlab("Viral load (copies / ml)")
+#     # axis.text.x = element_blank(),
+#     axis.title.y = element_text(size = 10),
+#     panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
+#     panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
+#   ) +
+#   ylab("log10(viral reads/cleaned reads/genome length)")
 
@@
   theme(
-    # axis.title.x = element_blank(),
+    axis.title.x = element_blank(),
     # axis.text.x = element_blank(),
+    legend.title = element_blank(),
     axis.title.y = element_text(size = 10),
@@
   ) +
-  ylab("Log(viral reads/cleaned reads/genome length)") +
+  ylab("log10(viral reads/cleaned reads/genome length)") + xlab("Log (Viral load)") +
   scale_color_manual(values = cols3, labels = c("Pool 1", "Pool 2")) + 
-  xlab("Viral load (copies / ml)")
+    scale_x_discrete(labels = c(bquote(10^{2}), bquote(10^{3}), bquote(10^{5})))
 
@@
 
-pool_readdepths <- 
-  depths_pool |>
-  filter(Background != "p6") |>
-  filter(Background != "control") |>
-  ggplot(aes(x = virus, y = log(mean_depth))) +
-  geom_boxplot() +
-  facet_grid(Pool ~ type) +
-  theme_few() +
-  theme(
-    axis.title.x = element_blank(),
-    axis.text.x = element_text(angle = 45, hjust = 1),
-    axis.title.y = element_text(size = 10),
-    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
-    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
-  ) +
-  ylab("Log (mean read depth)") + xlab("Viral load (copies / ml)")
+# pool_readdepths <- 
+#   depths_pool |>
+#   filter(Background != "p6") |>
+#   filter(Background != "control") |>
+#   ggplot(aes(x = virus, y = log10(mean_depth))) +
+#   geom_boxplot() +
+#   facet_grid(Pool ~ type) +
+#   theme_few() +
+#   theme(
+#     axis.title.x = element_blank(),
+#     axis.text.x = element_text(angle = 45, hjust = 1),
+#     axis.title.y = element_text(size = 10),
+#     panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
+#     panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
+#   ) +
+#   ylab("Log (mean read depth)")
 
@@
     x = as.character(Viral.load),
-    y = log(mean_depth),
+    y = log10(mean_depth),
     colour = Pool
@@
   ylab("Log (mean read depth)") +
-  xlab("Viral load (copies / ml)") +
+  xlab("Log (Viral load)") + 
+  scale_x_discrete(labels = c(bquote(10^{2}), bquote(10^{3}), bquote(10^{5}))) +
   scale_color_manual(values = cols3, labels = c("Pool 1", "Pool 2"))
@@
 
-#ggsave("figures/manuscript_figures_pdf/FigureS2.pdf",width=10,height=7)
+ggsave("figures/manuscript_figure_2025/PDF/Figure_S2.pdf",
+       width = 10,
+       height = 7)
+
+ggsave("figures/manuscript_figure_2025/PNG/Figure_S2.png",
+       width = 10,
+       height = 7)
 
@@
     x = Viral.load,
-    y = log(total_reads),
+    y = log10(total_reads),
     colour = Background
@@
   ) +
-  scale_x_log10(labels = scales::label_log()) + xlab("Viral load (copies / ml)") +
+  scale_x_log10(labels = scales::label_log10()) + xlab("Viral load (copies / ml)") +
   ggtitle("Total Reads") +
-  ylab("Log(Total Reads)") +
+  ylab("log10(Total Reads)") +
   scale_color_manual(values = cols2, labels = c("P1", "P2"))
@@
     x = Viral.load,
-    y = log(viral_reads),
+    y = log10(viral_reads),
     colour = Background
@@
   ) +
-  scale_x_log10(labels = scales::label_log()) + xlab("Viral load (copies / ml)") +
+  scale_x_log10(labels = scales::label_log10()) + xlab("Viral load (copies / ml)") +
   ggtitle("Total Viral Reads") +
-  ylab("Log(Total Viral Reads)") +
+  ylab("log10(Total Viral Reads)") +
   scale_color_manual(values = cols2, labels = c("P1", "P2"))
@@
     x = Viral.load,
-    y = log(ATCC_reads),
+    y = log10(ATCC_reads),
     colour = Background
@@
   ) +
-  scale_x_log10(labels = scales::label_log()) + xlab("Viral load (copies / ml)") +
+  scale_x_log10(labels = scales::label_log10()) + xlab("Viral load (copies / ml)") +
   ggtitle("Spike In Viral Reads") +
-  ylab("Log(Spike In Viral Reads)") +
+  ylab("log10(Spike In Viral Reads)") +
   scale_color_manual(values = cols2, labels = c("P1", "P2"))
@@
   ) +
-  scale_x_log10(labels = scales::label_log()) + xlab("Viral load (copies / ml)") +
+  scale_x_log10(labels = scales::label_log10()) + xlab("Viral load (copies / ml)") +
   ggtitle("Proportion Spike In Viral") +
@@
   ) +
-  scale_x_log10(labels = scales::label_log()) + xlab("Viral load (copies / ml)") +
+  scale_x_log10(labels = scales::label_log10()) + xlab("Viral load (copies / ml)") +
   ggtitle("Proportion Viral") +
@@
   ) +
-  scale_x_log10(labels = scales::label_log()) +
-  ylab("Log(Viral Reads)")
+  scale_x_log10(labels = scales::label_log10()) +
+  ylab("log10(Viral Reads)")
 
@@
   ) +
-  scale_x_log10(label = scales::label_log()) +
-  ylab("Log(viral reads/cleaned reads)")
+  scale_x_log10(label = scales::label_log10()) +
+  ylab("log10(viral reads/cleaned reads)")
 
@@
   ) +
-  scale_x_log10(labels = scales::label_log()) +
-  ylab("Log(viral reads/cleaned reads/genome length)")
+  scale_x_log10(labels = scales::label_log10()) +
+  ylab("log10(viral reads/cleaned reads/genome length)")
 
@@
   ) +
-  scale_x_log10(label = scales::label_log()) +
-  ylab("Log(mean read depth)")
+  scale_x_log10(label = scales::label_log10()) +
+  ylab("log10(mean read depth)")
 
@@
 ### --------------------
-
-### Figure S5 - linear model
-
-#### Read-in data
-```{r}
-#Linear model for TE data - separated by virus
-#November 2024
-
-#viral read depth
-#use bowtie2 data
-
-library(tidyverse)
-library(broom)
-library(boot)
-
-####read counts/normalised read counts####
-
-#metadata
-
-metadata <-
-  read.csv(
-    "metadata/sampleIDs_TESpikeIn.csv",
-    header = TRUE
-  )
-
-metadata2 <- metadata |>
-  select(Sample.ID, Number.of.read.pairs..quality.adaptor.trimmed.) |>
-  rename(Sample_id = Sample.ID) |>
-  rename(QC_reads = Number.of.read.pairs..quality.adaptor.trimmed.)
-
-#import and combine read count files
-#deduplicated and non-deduplicated
-
-counts_dedup_bt <-
-  read.table(
-    "data_TE/TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_dedup_atcc_ref.tsv",
-    sep = "\t",
-    header = TRUE
-  )
-
-counts_nodedup_bt <-
-  read.table(
-    "data_TE/TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv",
-    sep = "\t",
-    header = TRUE
-  )
-
-counts_bt_all <- rbind(counts_dedup_bt, counts_nodedup_bt)
-
-counts_reads <- left_join(counts_bt_all, metadata2, by = "Sample_id")
-
-####read depths####
-
-#import and combine read depth files
-
-depth_dedup_bt <-
-  read.table(
-    "data_TE/TE_sequencing_experiment_readdepth_per_sample_and_virus_bowtie2_dedup_atcc_ref.tsv",
-    sep = "\t",
-    header = TRUE
-  )
-
-depth_nodedup_bt <-
-  read.table(
-    "data_TE/TE_sequencing_experiment_readdepth_per_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv",
-    sep = "\t",
-    header = TRUE
-  )
-
-```
-
-#### Format data
-```{r}
-
-depths_bt_all <- rbind(depth_dedup_bt, depth_nodedup_bt)
-
-depths_reads <- left_join(depths_bt_all, metadata2, by = "Sample_id") |>
-  mutate(genome_structure = case_when((
-    virus == "Human_adenovirus_40" |
-      virus == "Human_betaherpesvirus"
-  ) ~ "DNA",
-  ,
-  .default = "RNA"
-  ))
-
-####linear model for spike in viruses - mean read depth####
-
-depths_reads_sub <- depths_reads |>
-  filter(Background != "p6") |>
-  filter(Background != "control") |>
-  filter(type == "dedup_TE") |>
-  mutate(log_depth = log(mean_depth)) |>
-  mutate(log_viral_load = log10(Viral.load))
-
-#inspect data
-hist(depths_reads_sub$mean_depth)
-hist(depths_reads_sub$Viral.load)
-hist(depths_reads_sub$log_depth)
-hist(depths_reads_sub$log_viral_load)
-
-summary(depths_reads_sub$log_depth)
-summary(depths_reads_sub$log_viral_load)
-
-```
-
-#### model
-```{r}
-#all viruses together
-
-lm1 <-
-  glm(log_depth ~ log_viral_load * virus,
-      data = depths_reads_sub,
-      family = "gaussian")
-
-summary(lm1)
-
-glm.diag.plots(lm1)
-
-pred <- predict(lm1, type = "response")
-
-rsq <- function (x, y) {
-  cor(x, y) ^ 2
-}
-
-rsq(depths_reads_sub$log_depth, pred)
-
-##split by viruses
-
-cols <- c("#4477AA",
-          "#66CCEE",
-          "#228833",
-          "#CCBB44",
-          "#EE6677",
-          "#AA3377")
-
-facet_names <- c(
-  "Human_adenovirus_40" = "Human adenovirus 40",
-  "Human_betaherpesvirus" = "Human betaherpesvirus",
-  "Human_respiratory_syncytial_virus" = "Human respiratory syncytial virus",
-  "Influenza_B_virus" = "Influenza B virus",
-  "Mammalian_orthoreovirus3" = "Mammalian orthoreovirus 3",
-  "Zika_virus" = "Zika virus"
-)
-
-virus <- (unique(depths_reads_sub$virus)) |>
-  rep(each = 2) |>
-  data.frame() |>
-  rename(virus = rep..unique.depths_reads_sub.virus....each...2.)
-
-list_models <- 
-  depths_reads_sub |>
-  group_split(virus) |>
-  map( ~ lm(log_depth ~ log_viral_load, data = .))
-
-lm_tidy <- 
-  map(list_models, broom::tidy) |>
-  do.call(rbind.data.frame, .) |>
-  cbind(virus, .) |>
-  rename(virus = ".") |>
-  select(virus, term, estimate) |>
-  pivot_wider(names_from = "term", values_from = "estimate")
-
-lm_summary <- depths_reads_sub |>
-  group_by(virus) |>
-  summarise(
-    Intercept = lm(log_depth ~ log_viral_load)$coefficients[1],
-    Coeff_x1 = lm(log_depth ~ log_viral_load)$coefficients[2],
-    R2 = summary(lm(log_depth ~ log_viral_load))$r.squared,
-    pvalue = summary(lm(log_depth ~ log_viral_load))$coefficients["log_viral_load", 4]
-  )
-
-lm_combined <- left_join(lm_tidy, lm_summary, by = "virus")
-
-ggplot() +
-  geom_point(data = depths_reads_sub,
-             mapping = aes(log_viral_load, log_depth, color = virus)) +
-  geom_abline(data = lm_combined, aes(intercept = `(Intercept)`, slope = log_viral_load)) +
-  geom_text(data = lm_combined,
-            colour = "black",
-            aes(
-              x = 4.7,
-              y = 3,
-              label = round(Coeff_x1, digits = 2)
-            )) +
-  geom_text(data = lm_combined,
-            colour = "black",
-            aes(x = 4, y = 3, label = "Slope =")) +
-  geom_text(data = lm_combined,
-            colour = "black",
-            aes(
-              x = 4.7,
-              y = 2,
-              label = round(R2, digits = 2)
-            )) +
-  geom_text(data = lm_combined,
-            colour = "black",
-            aes(x = 4, y = 2, label = "R2 =")) +
-  facet_wrap( ~ virus, labeller = as_labeller(facet_names)) +
-  theme_bw() +
-  xlab("Log(Viral Load)") +
-  ylab("Log(Mean Read Depth)") +
-  scale_color_manual(values = cols) +
-  guides(color = FALSE)
-
-#ggsave("figures/compare_spike_ins_atcc/model_readdepth_dedup.png")
-
-#ggsave("figures/manuscript_figures_pdf/FigureS5.pdf")
-
-```
-
 ### --------------------
@@
   facet_wrap(Viral.load ~ Sample_id, label = label_parsed) +
-  ylab("Log(Coverage)") +
+  ylab("log10(Coverage)") +
   ggtitle("Human RSV (deduplicated)") +
@@
   facet_wrap(Viral.load ~ Sample_id) +
-  ylab("Log(Coverage)") +
+  ylab("log10(Coverage)") +
   ggtitle("Zika virus (deduplicated)") +
@@
   facet_wrap(Viral.load ~ Sample_id) +
-  ylab("Log(Coverage)") +
+  ylab("log10(Coverage)") +
   ggtitle("Human Adenovirus (deduplicated)") +
@@
   facet_wrap(Viral.load ~ Sample_id) +
-  ylab("Log(Coverage)") +
+  ylab("log10(Coverage)") +
   ggtitle("Human Betaherpesvirus (deduplicated)") +
@@
   facet_grid(seg ~ Sample_id) +
-  ylab("Log(Coverage)") +
+  ylab("log10(Coverage)") +
   ggtitle("Influenza B virus (Background 1 deduplicated)") +
@@
   facet_grid(seg ~ Sample_id) +
-  ylab("Log(Coverage)") +
+  ylab("log10(Coverage)") +
   ggtitle("Influenza B virus (Background 2 deduplicated)") +
@@
 facet_grid(seg ~ Sample_id) +
-ylab("Log(Coverage)") +
+ylab("log10(Coverage)") +
 ggtitle("Mammalian orthoreovirus 3 (Background 1 deduplicated)") +
@@
 facet_grid(seg ~ Sample_id) +
-ylab("Log(Coverage)") +
+ylab("log10(Coverage)") +
 ggtitle("Mammalian orthoreovirus 3 (Background 2 deduplicated)") +
@@
     x = Viral.load,
-    y = log(count),
+    y = log10(count),
     colour = genus
@@
   scale_x_log10() +
-  ylab("Log(Viral Reads)") +
+  ylab("log10(Viral Reads)") +
   scale_color_manual(values = cols)

### 2025-01-21 manuscript_figures.Rmd
updating figures

@@
 
-lm_combined <- left_join(lm_tidy, lm_summary, by = "virus")
+lm_combined <- 
+  left_join(lm_tidy, lm_summary, by = "virus")
 
@@
 #### Make plot
-```{r}
+```{r, message=FALSE, warning=FALSE, echo=FALSE}
 ggplot() +
   geom_point(data = depths_reads_sub,
-             mapping = aes(log_viral_load, log_depth, color = virus)) +
+             mapping = aes(Viral.load, mean_depth, color = virus)) +
   geom_abline(data = lm_combined, aes(intercept = `(Intercept)`, slope = log_viral_load), linetype="dotted") +
@@
             aes(
-              x = 4.7,
-              y = 3,
+              x = 6*10000,
+              y = 2,
               label = round(Coeff_x1, digits = 2)),
-            size = 3) +
+            size = 3.5) +
   geom_text(data = lm_combined,
             colour = "black",
-            aes(x = 4, 
-                y = 3, 
+            aes(x = 1.5*10000, 
+                y = 2, 
                 label = "Slope ="),
-            size = 3) +
+            size = 3.5) +
   geom_text(data = lm_combined,
@@
             aes(
-              x = 4.7,
-              y = 2,
+              x = 6*10000,
+              y = 6,
               label = round(R2, digits = 2)),
-            size = 3) +
+            size = 3.5) +
   geom_text(data = lm_combined,
             colour = "black",
-            aes(x = 4, 
-                y = 2, 
+            aes(x = 2*10000, 
+                y = 6, 
                 label = "R2 ="),
-            size = 3) +
+            size = 3.5) +
   facet_wrap( ~ virus, labeller = as_labeller(facet_names)) +
@@
   ) +
-  xlab("log10(Viral Load)") +
-  ylab("log10(Mean Read Depth)") +
+    xlab("Log (Viral load)") + 
+  ylab("Log (Mean Read Depth)") +
   scale_color_manual(values = cols) +
-  guides(color = FALSE)
+  guides(color = FALSE) +
+  scale_x_log10(labels = scales::trans_format("log10", scales::label_math())) + 
+    scale_y_log10(labels = scales::trans_format("log10", scales::label_math()))
 
@@
 ggsave("figures/manuscript_figure_2025/PDF/Figure_1.pdf",
-       width = 7.5,
-       height = 4)
+       width = 8,
+       height = 5)
 
 ggsave("figures/manuscript_figure_2025/PNG/Figure_1.png",
-       width = 7.5,
-       height = 4)
+       width = 8,
+       height = 5)
 
@@
 
-### Final plot
+### ***Final plot***
 ```{r}

### 2025-01-22 manuscript_figures.Rmd
Figure S3 update

@@
 library(ggthemes)
+library(scales)
 
@@
     x = Viral.load,
-    y = log10(total_reads),
+    y = (total_reads),
     colour = Background
@@
   ) +
-  scale_x_log10(labels = scales::label_log10()) + xlab("Viral load (copies / ml)") +
-  ggtitle("Total Reads") +
-  ylab("log10(Total Reads)") +
+  scale_y_log10(labels = scales::trans_format("log10", scales::label_math())) + 
+    scale_x_log10(labels = scales::trans_format("log10", scales::label_math())) + xlab("Viral load") +
+  # ggtitle("All") +
+  ylab("Total number of reads") +
   scale_color_manual(values = cols2, labels = c("P1", "P2"))
@@
     x = Viral.load,
-    y = log10(viral_reads),
+    y = (viral_reads),
     colour = Background
@@
   ) +
-  scale_x_log10(labels = scales::label_log10()) + xlab("Viral load (copies / ml)") +
-  ggtitle("Total Viral Reads") +
-  ylab("log10(Total Viral Reads)") +
+  scale_y_log10(labels = scales::trans_format("log10", scales::label_math())) + 
+    scale_x_log10(labels = scales::trans_format("log10", scales::label_math())) + xlab("Viral load") +
+  # ggtitle("Viral reads") +
+  ylab("Number of viral reads") +
   scale_color_manual(values = cols2, labels = c("P1", "P2"))
@@
     x = Viral.load,
-    y = log10(ATCC_reads),
+    y = (ATCC_reads),
     colour = Background
@@
   ) +
-  scale_x_log10(labels = scales::label_log10()) + xlab("Viral load (copies / ml)") +
-  ggtitle("Spike In Viral Reads") +
-  ylab("log10(Spike In Viral Reads)") +
+  scale_y_log10(labels = scales::trans_format("log10", scales::label_math())) + 
+    scale_x_log10(labels = scales::trans_format("log10", scales::label_math())) + xlab("Viral load") +
+  ggtitle("Spike-in Viral Reads") +
+  ylab("Spike-in Viral Reads") +
   scale_color_manual(values = cols2, labels = c("P1", "P2"))
@@
   ) +
-  scale_x_log10(labels = scales::label_log10()) + xlab("Viral load (copies / ml)") +
-  ggtitle("Proportion Spike In Viral") +
-  ylab("Spike In Viral Reads/Total Reads") +
+    scale_x_log10(labels = scales::trans_format("log10", scales::label_math())) + xlab("Viral load") +
+  ggtitle("Spike-in viral reads") +
+  ylab("Proportion") +
   scale_color_manual(values = cols2, labels = c("P1", "P2"))
@@
   ) +
-  scale_x_log10(labels = scales::label_log10()) + xlab("Viral load (copies / ml)") +
-  ggtitle("Proportion Viral") +
-  ylab("Viral Reads/Total Reads") +
-  scale_color_manual(values = cols2, labels = c("P1", "P2"))
+ scale_x_log10(labels = scales::trans_format("log10", scales::label_math())) + xlab("Viral load") +
+  # ggtitle("Viral reads") +
+  ylab("Proportion of viral reads") +
+  scale_color_manual(values = cols2, labels = c("P1", "P2")) + scale_y_continuous(limits=c(0,1.0))
 
@@
 ### ***Final plot***
-```{r}
+```{r, warning=FALSE, echo=FALSE, message=FALSE}
+
 ggarrange(total_reads,
@@
 
-#ggsave("figures/manuscript_figures_pdf/FigureS3.pdf")
+ggsave("figures/manuscript_figure_2025/PDF/Figure_S3.pdf", 
+       width = 8,
+       height = 8)
 
@@
 
-## Figure S4
+## Not required ?Figure S4
 

### 2025-01-22 manuscript_figures.Rmd
Figure S4 update

@@
 
-## Not required ?Figure S4
+## No longer required? Figure S4
 
@@
   facet_wrap(Viral.load ~ Sample_id, label = label_parsed) +
-  ylab("log10(Coverage)") +
+  ylab("Coverage") +
   ggtitle("Human RSV (deduplicated)") +
@@
 ```{r}
-#ggsave("figures/manuscript_figures_pdf/FigureS5.pdf",width=10,height=8)
+ggsave("figures/manuscript_figure_2025/PDF/Figure_S4.pdf",
+       width=9,
+       height=8)
+
+ggsave("figures/manuscript_figure_2025/PNG/Figure_S4.png",
+       width=9,
+       height=8)
 ```

### 2025-01-22 manuscript_figures.Rmd
Updating supp figures

@@
 ggsave("figures/manuscript_figure_2025/PDF/Figure_S3.pdf", 
+       width = 8,
+       height = 8)
+
+ggsave("figures/manuscript_figure_2025/PNG/Figure_S3.png", 
        width = 8,
@@
 
-### Figure S6
+### Figure S4
 ```{r}
@@
 ggsave("figures/manuscript_figure_2025/PDF/Figure_S4.pdf",
-       width=9,
-       height=8)
+       width=15,
+       height=10)
 
 ggsave("figures/manuscript_figure_2025/PNG/Figure_S4.png",
-       width=9,
-       height=8)
+       width=15,
+       height=10)
 ```
@@
 
-### Figure S7
-```{r}
+### Figure S5
+```{r, warning=FALSE, message=FALSE, echo=FALSE}
 
@@
   filter(virus == "Zika_virus") |>
+  rename(Viral.load = `Viral load`) |>
   ggplot(aes(x = site, y = coverage, fill = Background)) +
   geom_col() +
-  facet_wrap(Viral.load ~ Sample_id) +
-  ylab("log10(Coverage)") +
+  facet_wrap(Viral.load ~ Sample_id, label = label_parsed) +
+  ylab("Coverage") +
   ggtitle("Zika virus (deduplicated)") +
   guides(fill = guide_legend(title = "Background")) +
-  theme_bw() +
+  theme_few() +
+  theme(
+    axis.title.x = element_text(size = 12),
+    axis.text.x = element_text(angle = 30, hjust = 1),
+    axis.title.y = element_text(size = 12),
+    legend.title = element_blank(),
+    legend.position = "top", 
+    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
+    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
+  ) +
   scale_y_log10() +
@@
 
-#ggsave("figures/manuscript_figures_pdf/FigureS6.pdf",width=10,height=8)
+```
+
+####***Final plot***
+```{r}
+
+ggsave("figures/manuscript_figure_2025/PDF/Figure_S5.pdf",
+       width=15,
+       height=10)
+
+ggsave("figures/manuscript_figure_2025/PNG/Figure_S5.png",
+       width=15,
+       height=10)
 
@@
 
-### Figure S8
-```{r}
+### Figure S6
+```{r, warning=FALSE, message=FALSE, echo=FALSE}
+
 persite_norm |>
@@
   filter(virus == "Human_adenovirus_40") |>
+  rename(Viral.load = `Viral load`) |>
   ggplot(aes(x = site, y = coverage, fill = Background)) +
   geom_col() +
-  facet_wrap(Viral.load ~ Sample_id) +
-  ylab("log10(Coverage)") +
+  facet_wrap(Viral.load ~ Sample_id, label = label_parsed) +
+  ylab("Coverage") +
   ggtitle("Human Adenovirus (deduplicated)") +
   guides(fill = guide_legend(title = "Background")) +
-  theme_bw() +
+ theme_few() +
+  theme(
+    axis.title.x = element_text(size = 12),
+    axis.text.x = element_text(angle = 30, hjust = 1),
+    axis.title.y = element_text(size = 12),
+    legend.title = element_blank(),
+    legend.position = "top", 
+    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
+    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
+  ) +
   scale_y_log10() +
@@
 
-#ggsave("figures/manuscript_figures_pdf/FigureS7.pdf",width=20,height=18)
+```
+
+####***Final plot***
+```{r}
+
+ggsave("figures/manuscript_figure_2025/PDF/Figure_S6.pdf",
+       width=15,
+       height=10)
+
+ggsave("figures/manuscript_figure_2025/PNG/Figure_S6.png",
+       width=15,
+       height=10)
+
 ```
@@
 
-### Figure S9
-```{r}
+### Figure S7
+```{r}
+HBV <- 
 persite_norm |>
@@
   filter(virus == "Human_betaherpesvirus") |>
+  rename(Viral.load = `Viral load`) |>
   ggplot(aes(x = site, y = coverage, fill = Background)) +
-  geom_polygon() +
-  facet_wrap(Viral.load ~ Sample_id) +
-  ylab("log10(Coverage)") +
+  geom_col() +
+  facet_wrap(Viral.load ~ Sample_id, label = label_parsed) +
+  ylab("Coverage") +
   ggtitle("Human Betaherpesvirus (deduplicated)") +
   guides(fill = guide_legend(title = "Background")) +
-  theme_bw() +
+ theme_few() +
+  theme(
+    axis.title.x = element_text(size = 12),
+    axis.text.x = element_text(angle = 30, hjust = 1),
+    axis.title.y = element_text(size = 12),
+    legend.title = element_blank(),
+    legend.position = "top", 
+    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
+    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
+  ) +
   scale_y_log10() +
@@
 
-#ggsave("figures/compare_spike_ins_atcc/HBHV_coverage_test.pdf",width=20,height=18)
+```
+
+####***Final plot***
+```{r}
+
+ggsave(plot = HBV, file="figures/manuscript_figure_2025/PDF/Figure_S7.pdf",
+       width=15,
+       height=10,
+       dpi = 600)
+
+ggsave(plot = HBV, file="figures/manuscript_figure_2025/PNG/Figure_S7.png",
+       width=15,
+       height=10)
+
 ```
@@
 
-### Figure S10
+### Figure S8
 ```{r}
@@
   filter(Background == "p2") |>
+  rename(Viral.load = `Viral load`) |>
   ggplot(aes(
@@
   facet_grid(seg ~ Sample_id) +
-  ylab("log10(Coverage)") +
+  ylab("Coverage") +
   ggtitle("Influenza B virus (Background 1 deduplicated)") +
   guides(fill = guide_legend(title = "Viral load")) +
-  theme_bw() +
+ theme_few() +
+  theme(
+    axis.title.x = element_text(size = 12),
+    axis.text.x = element_text(angle = 30, hjust = 1),
+    axis.title.y = element_text(size = 12),
+    legend.title = element_blank(),
+    legend.position = "top", 
+    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
+    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
+  ) +
   scale_y_log10() +
-  scale_fill_manual(values = cols4,
-                    labels = c("1e+02", "1e+03", "1e+05"))
+  scale_fill_manual(values = cols4, label = c(expression("10"^"2"), expression("10"^"3"), expression("10"^"5")))
 
@@
   filter(type == "dedup_TE") |>
-  filter(Background == "p8") |>
+  filter(Background == "p2") |>
+  rename(Viral.load = `Viral load`) |>
   ggplot(aes(
@@
   facet_grid(seg ~ Sample_id) +
-  ylab("log10(Coverage)") +
+  ylab("Coverage") +
   ggtitle("Influenza B virus (Background 2 deduplicated)") +
   guides(fill = guide_legend(title = "Viral load")) +
-  theme_bw() +
+ theme_few() +
+  theme(
+    axis.title.x = element_text(size = 12),
+    axis.text.x = element_text(angle = 30, hjust = 1),
+    axis.title.y = element_text(size = 12),
+    legend.title = element_blank(),
+    legend.position = "top", 
+    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
+    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
+  ) +
   scale_y_log10() +
-  scale_fill_manual(values = cols4,
-                    labels = c("1e+02", "1e+03", "1e+05"))
+  scale_fill_manual(values = cols4, label = c(expression("10"^"2"), expression("10"^"3"), expression("10"^"5")))
 
@@
 
-#ggsave("figures/manuscript_figures_pdf/FigureS9.pdf",width=20,height=18)
+```
+
+####***Final plot***
+```{r}
+
+ggsave("figures/manuscript_figure_2025/PDF/Figure_S8.pdf",
+       width=20,
+       height=12,
+       dpi=600)
+
+ggsave("figures/manuscript_figure_2025/PNG/Figure_S8.png",
+       width=20,
+       height=12)
 ```
@@
 
-### Figure S11
-```{r}
-
-REO_ME1 <- persite_norm |>
+### Figure S9
+```{r}
+
+REO_ME1 <- 
+  persite_norm |>
   filter(virus == "Mammalian_orthoreovirus3") |>
@@
   guides(fill = guide_legend(title = "Viral load")) +
-theme_bw() +
+  theme_few() +
+  theme(
+    axis.title.x = element_text(size = 12),
+    axis.text.x = element_text(angle = 30, hjust = 1),
+    axis.title.y = element_text(size = 12),
+    legend.title = element_blank(),
+    legend.position = "top",
+    panel.grid.major = element_line(linewidth = 0.1, color = "gray60"),
+    panel.grid.minor = element_line(linewidth = 0.1, color = "lightgray")
+  ) +
   scale_y_log10() +
   scale_fill_manual(values = cols4,
-labels = c("1e+02", "1e+03", "1e+05"))
-
-REO_ME2 <- persite_norm |>
+                    label = c(
+                      expression("10" ^ "2"),
+                      expression("10" ^ "3"),
+                      expression("10" ^ "5")
+                    ))
+
+REO_ME2 <- 
+  persite_norm |>
   filter(virus == "Mammalian_orthoreovirus3") |>
@@
   filter(Background == "p8") |>
+  rename(Viral.load = `Viral load`) |>
   ggplot(aes(
@@
   guides(fill = guide_legend(title = "Viral load")) +
-theme_bw() +
-scale_y_log10() +
+  theme_few() +
+  theme(
+    axis.title.x = element_text(size = 12),
+    axis.text.x = element_text(angle = 30, hjust = 1),
+    axis.title.y = element_text(size = 12),
+    legend.title = element_blank(),
+    legend.position = "top",
+    panel.grid.major = element_line(linewidth = 0.1, color = "gray60"),
+    panel.grid.minor = element_line(linewidth = 0.1, color = "lightgray")
+  ) + scale_y_log10() +
   scale_fill_manual(values = cols4,
-labels = c("1e+02", "1e+03", "1e+05"))
+                    label = c(
+                      expression("10" ^ "2"),
+                      expression("10" ^ "3"),
+                      expression("10" ^ "5")
+                    ))
 
@@
 
-#ggsave("figures/manuscript_figures_pdf/FigureS10.pdf",width=20,height=18)
-
+```
+
+####***Final plot***
+```{r}
+ggsave("figures/manuscript_figure_2025/PDF/Figure_S9.pdf",
+       width=20,
+       height=12,
+       dpi=600)
+
+ggsave("figures/manuscript_figure_2025/PNG/Figure_S9.png",
+       width=20,
+       height=12,
+       dpi=600)
 ```

### 2025-01-30 manuscript_figures.Rmd
update Figure 3

@@
   "Msp_p8" = "P2",
-  "off_target" = "Off target viruses",
-  "on_target" = "On target viruses"
+  "off_target" = "Background",
+  "on_target" = "Spike-in"
 )
@@
     x = Viral.load,
-    y = log10(count),
+    y = (count),
     colour = genus
@@
   scale_x_log10() +
-  ylab("Log (Viral Reads)") + 
-  xlab("Log (Viral load)") + 
+  scale_y_log10() +
+  ylab("Viral Reads") + 
+  xlab("Viral load (gc/ml)") + 
   scale_color_manual(values = cols) +
+  scale_linetype_manual(values=c("solid", "dotted")) + 
+  scale_x_log10(labels = scales::trans_format("log10", scales::label_math())) + 
+    scale_y_log10(labels = scales::trans_format("log10", scales::label_math())) +
   theme(
     axis.title.x = element_text(size = 12),
-    axis.text.x = element_text(angle = 30, hjust = 1),
+    # axis.text.x = element_text(hjust = 1),
     axis.title.y = element_text(size = 12),
@@
   scale_y_log10() +
+  scale_x_log10(labels = scales::trans_format("log10",
+                                              scales::label_math())) + 
+  scale_y_log10(labels = scales::trans_format("log10",
+                                              scales::label_math())) +
+    scale_linetype_manual(values=c("solid", "dotted"))+ 
   ylab("Proportion of viral reads") +
-  xlab("Log (Viral load)") + 
+  xlab("Viral load (gc/ml)") + 
   scale_color_manual(values = cols) +
@@
     axis.title.x = element_text(size = 12),
-    axis.text.x = element_text(angle = 30, hjust = 1),
+    # axis.text.x = element_text(angle = 30, hjust = 1),
     axis.title.y = element_text(size = 12),
@@
   theme_few() + 
-  scale_x_log10() +
-  scale_y_log10(limits=c(5e-8,1)) +
+  scale_x_log10(labels = scales::trans_format("log10",
+                                              scales::label_math())) + 
+  scale_y_log10(limits = c(5e-8, 1), breaks = 10**(seq(-6,0,2)),labels = scales::trans_format("log10",
+                                              scales::label_math())) +
+    scale_linetype_manual(values=c("solid", "dotted"))+ 
   ylab("Proportion of total reads") +
-  xlab("Log (Viral load)") + 
+  xlab("Viral load (gc/ml)") + 
   scale_color_manual(values = cols) +
@@
     axis.title.x = element_text(size = 12),
-    axis.text.x = element_text(angle = 30, hjust = 1),
+    # axis.text.x = element_text(angle = 30, hjust = 1),
     axis.title.y = element_text(size = 12),
+    axis.text.y = element_text(hjust = 1),
     legend.title = element_blank(),
@@
 
-#ggsave("figures/compare_spike_ins_atcc/target_offtarget_dedup.png",width=8,height=10)
-
 ##just plot viral reads and prop viral
@@
 
-ggsave("figures/compare_spike_ins_atcc/target_offtarget_dedup_2025-01-01.png", width = 10, height = 8)
-
-ggsave("figures/manuscript_figures_pdf/Figure_3_2025-01-01.pdf", width = 10, height = 8)
-
+# ggsave("figures/compare_spike_ins_atcc/target_offtarget_dedup_2025-01-01.png", width = 10, height = 8)
+
+ggsave("figures/manuscript_figure_2025/PDF/Figure_3.pdf", 
+       width = 10, height = 8)
+
+ggsave("figures/manuscript_figure_2025/PNG/Figure_3.png", 
+       width = 10, height = 8)
 ```

### 2025-01-30 manuscript_figures.Rmd
update Figure 3

--- 2025-01-22 manuscript_figures.Rmd
@@
   scale_x_log10() +
-  ylab("Log (Viral Reads)") + 
-  xlab("Log (Viral load)") + 
+  ylab("Viral Reads") + 
+  xlab("Viral load (gc/ml)") + 
   scale_color_manual(values = cols) +
+  scale_linetype_manual(values=c("solid", "dotted"))+ 
   theme(
     axis.title.x = element_text(size = 12),
-    axis.text.x = element_text(angle = 30, hjust = 1),
+    # axis.text.x = element_text(angle = 30, hjust = 1),
     axis.title.y = element_text(size = 12),
@@
   scale_y_log10() +
+    scale_linetype_manual(values=c("solid", "dotted"))+ 
   ylab("Proportion of viral reads") +
-  xlab("Log (Viral load)") + 
+  xlab("Viral load (gc/ml)") + 
   scale_color_manual(values = cols) +
@@
   ylab("Proportion of total reads") +
-  xlab("Log (Viral load)") + 
+  xlab("Viral load (gc/ml)") + 
+  scale_linetype_manual(values=c("solid", "dotted"))+ 
   scale_color_manual(values = cols) +

### 2025-01-30 manuscript_figures.Rmd
Update figure 1

--- 2025-01-30 manuscript_figures.Rmd
@@
 ```{r, message=FALSE, warning=FALSE, echo=FALSE}
+
 ggplot() +
@@
   ) +
-    xlab("Log (Viral load)") + 
-  ylab("Log (Mean Read Depth)") +
+    xlab("Viral load (gc/ml)") + 
+  ylab("Mean Read Depth") +
   scale_color_manual(values = cols) +

### 2025-01-30 manuscript_figures.Rmd
Figures S10-S12 from genom med paper

@@
 ```
-
-### ----nodedup results ----
-```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=12, fig.width=6}
-##extras - no dedup
-
-nodedup_reads <-
-  nodedup_viral_reads |>
-  filter(Background.sample != "Neg_control") |>
-  filter(Background.sample != "Msp_p6") |>
+### --------------------
+
+### Figure S10
+
+#### Read-in data
+```{r}
+
+####read counts/normalised read counts####
+
+#import and combine read count files
+
+gm_readcount_dedup <-
+  read.table(
+    "data_genome_medicine/genome_medicine_TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_dedup_atcc_ref.tsv",
+    sep = "\t",
+    header = TRUE
+  )
+
+gm_readcount_nodedup <-
+  read.table(
+    "data_genome_medicine/genome_medicine_TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv",
+    sep = "\t",
+    header = TRUE
+  )
+
+gm_counts_all <- rbind(gm_readcount_dedup, gm_readcount_nodedup)
+
+```
+
+#### read counts 
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+#normalise by both raw read count and genome length - should be the same as mean read depth
+
+facet_names <- c("dedup_GM_TE" = "Deduplicated",
+                 "nodedup_GM_TE" = "Non-Deduplicated")
+
+cols <- c("#4477AA",
+          "#66CCEE",
+          "#228833",
+          "#CCBB44",
+          "#EE6677",
+          "#AA3377")
+
+gm_counts_norm <- 
+  gm_counts_all |>
+  mutate(norm_counts3 = matched / length)
+
+read_counts <- 
+  gm_counts_norm |>
+  filter(Viral.load != 0) |>
   ggplot(aes(
     x = Viral.load,
-    y = log10(count),
-    colour = genus
+    y = (matched),
+    colour = virus
   )) +
   geom_point() +
-  geom_smooth(aes(group = genus, linetype = genome_structure), se = FALSE) +
-  facet_grid(target ~ Background.sample, labeller = as_labeller(facet_names)) +
+  geom_smooth(aes(group = virus), se = FALSE, span = 0.85) +
+  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
   theme_few() +
@@
     axis.title.x = element_text(size = 12),
-    axis.text.x = element_text(angle = 0, hjust = 1),
+    # axis.text.x = element_text(angle = 30, hjust = 1),
     axis.title.y = element_text(size = 12),
@@
     legend.position = "top",
-    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
-    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
-  ) +
-  scale_x_log10() +
-  ylab("log10(Viral Reads)") +
-  scale_color_manual(values = cols)
-
-nodedup_prop_viral <- 
-  nodedup_viral_reads |>
-  filter(Background.sample != "Neg_control") |>
-  filter(Background.sample != "Msp_p6") |>
-  ggplot(aes(x = Viral.load, y = prop_viral, colour = genus)) +
+    panel.grid.major = element_line(linewidth = 0.1, color = "gray60"),
+    panel.grid.minor = element_line(linewidth = 0.1, color = "lightgray")
+  ) +
+  scale_color_manual(
+    values = cols,
+    labels = c(
+      "Human adenovirus 40",
+      "Human betaherpesvirus",
+      "Human respiratory syncytial virus",
+      "Influenza B virus",
+      "Mammalian orthoreovirus 3",
+      "Zika virus"
+    )
+  ) +
+  ylab("Viral Reads") +
+  xlab("Viral load (gc/ml)") + 
+  scale_x_log10(labels = scales::trans_format("log10",
+                                              scales::label_math())) + 
+  scale_y_log10(labels = scales::trans_format("log10",
+                                              scales::label_math()))
+
+```
+
+#### read depth
+```{r}
+#import and combine read depth files
+
+depth_dedup_gm <-
+  read.table(
+    "data_genome_medicine/genome_medicine_TE_sequencing_experiment_readdepth_per_sample_and_virus_bowtie2_dedup_atcc_ref.tsv",
+    sep = "\t",
+    header = TRUE
+  )
+
+depth_nodedup_gm <-
+  read.table(
+    "data_genome_medicine/genome_medicine_TE_sequencing_experiment_readdepth_per_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv",
+    sep = "\t",
+    header = TRUE
+  )
+
+depths_gm_all <- rbind(depth_dedup_gm, depth_nodedup_gm)
+
+```
+
+```{r, echo=FALSE, warning=FALSE, message=FALSE}
+
+read_depths <- 
+  depths_gm_all |>
+  filter(Viral.load != 0) |>
+  ggplot(aes(
+    x = Viral.load,
+    y = mean_depth,
+    colour = virus
+  )) +
   geom_point() +
-  geom_smooth(aes(group = genus, linetype = genome_structure), se = FALSE) +
-  facet_grid(target ~ Background.sample, labeller = as_labeller(facet_names)) +
+  geom_smooth(aes(group = virus), se = FALSE, span = 0.9) +
+  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
   theme_few() +
@@
     legend.position = "top",
-    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
-    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
-  ) +
-  scale_x_log10() +
-  scale_y_log10() +
-  ylab("Proportion of viral reads") +
-  scale_color_manual(values = cols)
-
-nodedup_prop_all <- nodedup_viral_reads |>
-  filter(Background.sample != "Neg_control") |>
-  filter(Background.sample != "Msp_p6") |>
-  ggplot(aes(x = Viral.load, y = prop_total, colour = genus)) +
-  geom_point() +
-  geom_smooth(aes(group = genus, linetype = genome_structure), se = FALSE) +
-  facet_grid(target ~ Background.sample, labeller = as_labeller(facet_names)) +
-  theme_few() +
-  theme(
-    axis.title.x = element_text(size = 12),
-    axis.text.x = element_text(angle = 0, hjust = 1),
-    axis.title.y = element_text(size = 12),
+    panel.grid.major = element_line(linewidth = 0.1, color = "gray60"),
+    panel.grid.minor = element_line(linewidth = 0.1, color = "lightgray")
+  ) +
+  scale_color_manual(
+    values = cols,
+    labels = c(
+      "Human adenovirus 40",
+      "Human betaherpesvirus",
+      "Human respiratory syncytial virus",
+      "Influenza B virus",
+      "Mammalian orthoreovirus 3",
+      "Zika virus"
+    )
+  ) +
+  ylab("Mean read depth") +
+  xlab("Viral load (gc/ml)") + 
+  scale_x_log10(labels = scales::trans_format("log10",
+                                              scales::label_math())) + 
+  scale_y_log10(labels = scales::trans_format("log10",
+                                              scales::label_math()))
+
+```
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+
+ggarrange(read_counts,
+          read_depths,
+          nrow = 2,
+          common.legend = TRUE)
+
+#PDF
+ggsave("figures/manuscript_figure_2025/PDF/Figure_S10.pdf", 
+       width = 7,
+       height = 7)
+
+#PNG
+ggsave("figures/manuscript_figure_2025/PNG/Figure_S10.png", 
+       width = 7,
+       height = 7)
+
+```
+
+### Figure S11
+
+#### Read-in data
+```{r}
+####genome coverage####
+
+unzip(zipfile = "data_genome_medicine/genome_medicine_TE_sequencing_experiment_readdepth_per_site_sample_and_virus_bowtie2_dedup_atcc_ref.tsv.zip", exdir = "data_genome_medicine/")
+
+dedup_per_site <-
+  read.table(
+    "data_genome_medicine/genome_medicine_TE_sequencing_experiment_readdepth_per_site_sample_and_virus_bowtie2_dedup_atcc_ref.tsv",
+    sep = "\t",
+    header = TRUE
+  )
+
+file.remove("data_genome_medicine/genome_medicine_TE_sequencing_experiment_readdepth_per_site_sample_and_virus_bowtie2_dedup_atcc_ref.tsv")
+
+#View(dedup_per_site)
+
+counts_dedup_bt <-
+  read.table(
+    "data_TE/TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_dedup_atcc_ref.tsv",
+    sep = "\t",
+    header = TRUE
+  )
+
+#add length column from read depth file to per site file
+
+lengths <- 
+  counts_dedup_bt |>
+  select(virus,length) |>
+  distinct()
+
+#calculate genome coverage
+
+coverage <- 
+  dedup_per_site |>
+  group_by(sample_id, virus, Viral.load, Replicate) |>
+  filter(coverage > 0) |>
+  summarise(genome_coverage = n()) |>
+  left_join(lengths) |>
+  mutate(percent_coverage = genome_coverage / length)
+
+coverage |>
+  filter(Viral.load != 0) |>
+  ggplot(aes(x = virus, y = percent_coverage)) +
+  geom_boxplot(outlier.shape = NA) +
+  geom_point(position = position_dodge(width = .75)) +
+  facet_grid( ~ as.character(Viral.load)) +
+  theme_bw() +
+  theme(axis.title.y = element_text(size = 10)) +
+  ylab("Genome coverage") +
+  theme_few() +
+  theme(
+    axis.title.x = element_blank(),
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    axis.title.y = element_text(size = 15),
+    axis.text = element_text(size = 12),
+    strip.text = element_text(size = 12),
     legend.title = element_blank(),
     legend.position = "top",
-    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
-    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
-  ) +
-  scale_x_log10() +
-  scale_y_log10() +
-  ylab("Proportion of total reads") +
-  scale_color_manual(values = cols)
-
-ggarrange(
-  nodedup_reads,
-  nodedup_prop_all,
-  nodedup_prop_viral,
-  nrow = 3,
-  common.legend = TRUE,
-  align = "hv"
-)
-
-#ggsave("figures/compare_spike_ins_atcc/target_offtarget_nodedup.pdf",width=8,height=10)
-
-```
-
-### --------------------
+    panel.grid.major = element_line(linewidth = 0.1, color = "gray60"),
+    panel.grid.minor = element_line(linewidth = 0.1, color = "lightgray")
+  ) +
+  scale_y_continuous(limits = c(0, 1)) +
+    scale_x_discrete(
+    labels = c(
+      "Human adenovirus 40",
+      "Human betaherpesvirus",
+      "Human respiratory syncytial virus",
+      "Influenza B virus",
+      "Mammalian orthoreovirus 3",
+      "Zika virus"
+    )
+  )
+
+#ggsave("/Users/laura/Dropbox/glasgow/github/te_ug_rodents/figures/genome_medicine/all_genome_coverage.png",width=14,height=6)
+
+#ggsave("/Users/laura/Dropbox/glasgow/github/te_ug_rodents/figures/manuscript_figures_pdf/FigureS12.pdf",width=20,height=6)
+```
+
+#### *** Final plot ***
+```{r}
+#PDF
+ggsave(
+  "figures/manuscript_figure_2025/PDF/Figure_S11.pdf",
+  width = 14,
+  height = 6
+)
+
+#PNG
+ggsave(
+  "figures/manuscript_figure_2025/PNG/Figure_S11.png",
+  width = 14,
+  height = 6
+)
+
+```
+
+### Figure S12
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+####per site plot AdV only####
+
+dedup_per_site |>
+  filter(virus == "Human_adenovirus_40") |>
+  filter(Viral.load != 0) |>
+  ggplot(aes(x=site,y=coverage))+
+  geom_col()+
+  facet_grid(Viral.load~Replicate)+
+  ylab("Coverage per site")+
+  ggtitle("Human Adenovirus (deduplicated)")+
+  theme_few() +
+  theme(
+    # axis.title.x = element_text(size = 12),
+    # # axis.text.x = element_text(angle = 30, hjust = 1),
+    # axis.title.y = element_text(size = 12),
+    # legend.title = element_blank(),
+    # legend.position = "top",
+    panel.grid.major = element_line(linewidth = 0.1, color = "gray60"),
+    panel.grid.minor = element_line(linewidth = 0.1, color = "lightgray")
+  ) +  
+  scale_y_log10(labels = scales::trans_format("log10",
+                                              scales::label_math())) +
+
+  xlab("site position")
+
+```
+
+#### *** Final plot ***
+```{r}
+ggsave("figures/manuscript_figure_2025/PDF/Figure_S12.pdf",
+       width=12,
+       height=10)
+
+ggsave("figures/manuscript_figure_2025/PNG/Figure_S12.png",
+       width=12,
+       height=10)
+```

### 2025-01-31 manuscript_figures.Rmd
Update kobu and cardio supp figs

@@
 ```
-### --------------------
-
+### .........................
+### .........................
+
+### ----Genome medicine results----
 ### Figure S10
@@
 ```
+### --------------------
 
@@
 ```
+### .........................
+
+### ----Kobuvirus / Cardiovirus results----
+
+#### Figure S13
+
+```{r}
+
+kobu_depth <-
+  read.table(
+    "data_kobuvirus/kobuvirus_TE_polyomics_readdepth_20241007.tsv",
+    sep = "\t",
+    header = TRUE
+  ) %>%
+  select(!expt) %>%
+  select(!Number.of.paired.end.reads..QT.)
+
+cardio_depth <-
+  read.table(
+    "data_cardiovirus/cardiovirus_denovo_bowtie2_read_depth_per_sample_dedup.tsv",
+    sep = ",",
+    header = TRUE
+  )
+
+depth_both <- rbind(kobu_depth, cardio_depth)
+
+depth_both$type <- depth_both$type %>%
+  case_match("dedup_TE" ~ "dedup",
+             .default = depth_both$type)
+
+cols2 <- c("#BB5566", "#004488")
+
+coverage_labels <- c(
+  "k97_19971" = "Kobuvirus",
+  "cardiovirus" = "Cardiovirus",
+  "p2" = "P1",
+  "p8" = "P2"
+)
+
+```
+
+#### *** Final plot ***
+```{r}
+
+depth_both |>
+  filter(Background != "p6") |>
+  filter(type == "dedup") |>
+  filter(mapper == "bowtie2") |>
+  filter(Viral.load != "NA") |>
+  ggplot(aes(
+    x = Viral.load,
+    y = (mean_depth),
+    colour = Background
+  )) +
+  geom_point() +
+ theme_few() +
+  theme(
+    axis.title.x = element_text(size = 12),
+    axis.title.y = element_text(size = 12),
+    legend.title = element_blank(),
+    legend.position = "none", 
+    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
+    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
+  ) +  
+  facet_grid(Background ~ virus, labeller = as_labeller(coverage_labels)) +
+  scale_color_manual(values = cols2) +
+  ylab("Mean read depth") +
+    xlab("Spike-in viral load (gc/ml)") +
+
+   scale_x_log10(labels = scales::trans_format("log10",
+                                              scales::label_math())) + 
+  scale_y_log10(labels = scales::trans_format("log10",
+                                              scales::label_math()))
+
+ggsave("figures/manuscript_figure_2025/PDF/Figure_S13.pdf",
+       width=8,
+       height=4)
+
+ggsave("figures/manuscript_figure_2025/PNG/Figure_S13.png",
+       width=8,
+       height=4)
+
+```
+
+### --------------------
+
+#### Figure S14
+
+```{r}
+
+kobu_per_site <-
+  read.table(
+    "data_kobuvirus/kobuvirus_TE_polyomics_readdepth_persite_20241007.tsv",
+    sep = "\t",
+    header = TRUE
+  ) %>%
+  select(virus,
+         seg,
+         site,
+         coverage,
+         n_sites,
+         Sample_id,
+         Background,
+         Viral.load,
+         mapper,
+         type)
+
+kobu_per_site$Viral.load <- 
+  kobu_per_site$Viral.load %>%
+  replace_na(., 0)
+
+kobu_per_site$Sample_id <- kobu_per_site$Sample_id %>%
+  case_match("RNA-Msp-p2" ~ "P1",
+             "RNA-Msp-p8" ~ "P2",
+             .default = kobu_per_site$Sample_id)
+
+kobu_per_site$Background <- kobu_per_site$Background %>%
+  case_match("p2" ~ "P1",
+             "p8" ~ "P2",
+             .default = kobu_per_site$Background)
+
+cardio_per_site <-
+  read.table(
+    "data_cardiovirus/cardiovirus_denovo_bowtie2_read_depth_per_site_and_sample_dedup.tsv",
+    sep = ",",
+    header = TRUE
+  )
+
+cardio_polyomics_per_site <-
+  read.table(
+    "data_cardiovirus/cardiovirus_denovo_bowtie2_readdepth_per_site_sample_dedup_polyomics.tsv",
+    sep = ",",
+    header = TRUE
+  ) %>%
+  select(!contig_name)
+
+#View(cardio_polyomics_per_site)
+
+cardio_TE_polyomics <-
+  full_join(
+    cardio_per_site,
+    cardio_polyomics_per_site,
+    by = c(
+      "virus",
+      "seg",
+      "site",
+      "coverage",
+      "n_sites",
+      "Sample_id",
+      "Background",
+      "Viral.load",
+      "mapper",
+      "type"
+    )
+  )
+
+cardio_TE_polyomics$Viral.load <- cardio_TE_polyomics$Viral.load %>%
+  replace_na(., 0)
+
+drops <-
+  c(
+    "RNA-Msp-p1",
+    "RNA-Msp-p3",
+    "RNA-Msp-p4",
+    "RNA-Msp-p5",
+    "RNA-Msp-p6",
+    "RNA-Msp-p7",
+    "RNA-Msp-p9"
+  )
+
+cardio_TE_polyomics <- cardio_TE_polyomics %>%
+  filter(!Sample_id %in% drops)
+
+cardio_TE_polyomics$Sample_id <- cardio_TE_polyomics$Sample_id %>%
+  case_match("RNA-Msp-p2" ~ "P1",
+             "RNA-Msp-p8" ~ "P2",
+             .default = cardio_TE_polyomics$Sample_id)
+
+cardio_TE_polyomics$Background <- cardio_TE_polyomics$Background %>%
+  case_match("S18" ~ "P1",
+             "S43" ~ "P2",
+             "p2" ~ "P1",
+             "p8" ~ "P2",
+             .default = cardio_TE_polyomics$Background)
+
+persite_both <- rbind(kobu_per_site, cardio_TE_polyomics)
+
+persite_both$type <- persite_both$type %>%
+  case_match("dedup_TE" ~ "dedup",
+             "dedup_polyomics" ~ "dedup",
+             .default = persite_both$type)
+
+#TE data only - compare coverage between backgrounds/viral loads
+
+cols2 <- c("#BB5566", "#004488")
+
+coverage_labels <- c(
+  "100" = "10^{2}",
+  "1000" = "10^{3}",
+  "100000" = "10^{5}",
+  "0" = "Shotgun",
+  "Kobuvirus" = "Kobuvirus",
+  "Cardiovirus" = "Cardiovirus"
+)
+
+coverage_labels2 <- c(
+  "100" = "10^{2}",
+  "1000" = "10^{3}",
+  "100000" = "10^{5}",
+  "0" = "Shotgun",
+  "A" = "A",
+  "B" = "B",
+  "C" = "C",
+  "D" = "D",
+  "F" = "F",
+  "G" = "G",
+  "H" = "H",
+  "I" = "I",
+  "J" = "J",
+  "K" = "K",
+  "L" = "L",
+  "M" = "M",
+  "N" = "N",
+  "O" = "O",
+  "P" = "P",
+  "ME_P1" = "P1",
+  "ME_P2" = "P2"
+)
+
+kobu_coverage <- kobu_per_site %>%
+  filter(mapper == "bowtie2") %>%
+  filter(type == "dedup") %>%
+  filter(Background != "p6") %>%
+  group_by(Sample_id, Viral.load, Background) %>%
+  filter(coverage > 0) %>%
+  summarise(genome_coverage = n()) %>%
+  mutate(percent_coverage = genome_coverage / 8467) %>%
+  mutate(virus = "Kobuvirus")
+
+cardio_coverage <- cardio_TE_polyomics %>%
+  filter(Background != "p6") %>%
+  group_by(Sample_id, Viral.load, Background) %>%
+  filter(coverage > 0) %>%
+  summarise(genome_coverage = n()) %>%
+  mutate(percent_coverage = genome_coverage / 6945) %>%
+  mutate(virus = "Cardiovirus")
+
+coverage_both <- rbind(kobu_coverage, cardio_coverage)
+
+coverage_both$Viral.load <- factor(coverage_both$Viral.load, labels = c("Shotgun",
+    "10^{2}", 
+    "10^{3}", 
+    "10^{5}"
+    ))
+
+```
+
+#### *** Final plot ***
+```{r}
+
+coverage_both %>%
+  ggplot(aes(x = Sample_id, y = percent_coverage, color = Background)) +
+  geom_point() +
+  ylab("Genome coverage") +
+  xlab("Sample ID") +
+  facet_grid(virus ~ (Viral.load),
+             scales = "free_x",
+             label = label_parsed) +
+  scale_color_manual(values = cols2, labels = c("P1", "P2")) +
+ theme_few() +
+  theme(
+    axis.title.x = element_text(size = 12),
+    axis.title.y = element_text(size = 12),
+    legend.title = element_blank(),
+    legend.position = "right", 
+    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
+    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
+  ) + 
+  scale_y_continuous(limits = c(0, 1))
+
+#PDF
+ggsave("figures/manuscript_figure_2025/PDF/Figure_S14.pdf", 
+       width = 8, 
+       height = 4)
+
+#PNG
+ggsave("figures/manuscript_figure_2025/PNG/Figure_S14.png", 
+       width = 8, 
+       height = 4)
+
+```
+
+### --------------------
+
+#### Figure S15
+
+```{r, echo=FALSE, warning=FALSE, message=FALSE}
+
+#per site genome coverage - Kobu
+
+persite_both$Viral.load <- factor(persite_both$Viral.load, labels = c("Shotgun",
+    "10^{2}", 
+    "10^{3}", 
+    "10^{5}"
+    ))
+
+persite_both$Viral.load <- factor(persite_both$Viral.load, levels = c(
+    "10^{2}", 
+    "10^{3}", 
+    "10^{5}",
+    "Shotgun"
+    ))
+
+persite_both |>
+  filter(virus == "k97_19971") |>
+  filter(mapper == "bowtie2") |>
+  filter(type == "dedup") |>
+  filter(Background != "p6") |>
+  ggplot(aes(
+    x = site,
+    y = (coverage),
+    fill = Background
+  )) +
+  geom_col() +
+  theme_few() +
+  theme(
+    axis.title.x = element_text(size = 12),
+    axis.title.y = element_text(size = 12),
+    legend.title = element_blank(),
+    legend.position = "right",
+    panel.grid.major = element_line(linewidth = 0.1, color = "gray60"),
+    panel.grid.minor = element_line(linewidth = 0.1, color = "lightgray")
+  ) +   
+  facet_wrap(Viral.load ~ Sample_id, labeller = label_parsed) +
+  scale_fill_manual(values = cols2, labels = c("P1", "P2")) +
+  ylab("Coverage") +
+  xlab("") +
+  ggtitle("Kobuvirus") +
+  scale_y_log10(labels = scales::trans_format("log10",
+                                              scales::label_math()))
+
+```
+
+#### *** Final plot ***
+```{r}
+#PDF
+ggsave("figures/manuscript_figure_2025/PDF/Figure_S15.pdf",
+       width=12,
+       height=7)
+
+#PNG
+ggsave("figures/manuscript_figure_2025/Png/Figure_S15.png",
+       width=12,
+       height=7)
+
+```
+
+#### Figure S16
+```{r, echo=FALSE, warning=FALSE, message=FALSE}
+
+#per site genome coverage - Cardio
+
+persite_both |>
+  filter(virus == "cardiovirus") |>
+  filter(mapper == "bowtie2") |>
+  filter(type == "dedup") |>
+  filter(Background != "p6") |>
+  ggplot(aes(
+    x = site,
+    y = coverage,
+    fill = Background
+  )) +
+  geom_col() +
+  facet_wrap(Viral.load ~ Sample_id, labeller = label_parsed) +
+  scale_fill_manual(values = cols2, labels = c("P1", "P2")) +
+  theme_few() +
+  theme(
+    axis.title.x = element_text(size = 12),
+    axis.title.y = element_text(size = 12),
+    legend.title = element_blank(),
+    legend.position = "right",
+    panel.grid.major = element_line(linewidth = 0.1, color = "gray60"),
+    panel.grid.minor = element_line(linewidth = 0.1, color = "lightgray")
+  ) +
+  ylab("Coverage") +
+  xlab("") +
+  ggtitle("Cardiovirus") +
+  scale_y_log10(labels = scales::trans_format("log10",
+                                              scales::label_math()))
+```
+
+#### *** Final plot ***
+```{r}
+
+#PDF
+ggsave("figures/manuscript_figure_2025/PDF/Figure_S16.pdf",
+       width=12,
+       height=7)
+
+#PNG
+ggsave("figures/manuscript_figure_2025/PNG/Figure_S16.png",
+       width=12,
+       height=7)
+```

### 2025-02-04 manuscript_figures.Rmd
Alternative plot for Fig S13

@@
 
-depth_both <- rbind(kobu_depth, cardio_depth)
-
-depth_both$type <- depth_both$type %>%
+cardio_depth_poly <- 
+  read.table(
+    "data_cardiovirus/cardiovirus_denovo_bowtie2_readdepth_per_sample_dedup_polyomics.tsv",
+    sep = ",",
+    header = TRUE
+  ) |>
+  select(-contig_name) |>
+  filter(Background %in% c("S18", "S19", "S43")) |>
+  select(-Background) |>
+  separate(Sample_id, into = c(NA, NA, "Background"), sep = "-", remove = FALSE)
+
+depth_both <- rbind(kobu_depth, cardio_depth, cardio_depth_poly) |>
+  mutate(Viral.load = ifelse(is.na(Viral.load), 0, Viral.load))
+
+depth_both$type <- depth_both$type |>
   case_match("dedup_TE" ~ "dedup",
+             "dedup_polyomics" ~ "dedup",
              .default = depth_both$type)
+
+depth_both$Sample_id <- depth_both$Sample_id |>
+  case_match("RNA-Msp-p2" ~ "P1",
+             "RNA-Msp-p8" ~ "P2",
+             "RNA-Msp-p6" ~ "P3",
+             .default = depth_both$Sample_id
+             )
+
+depth_both$virus <- depth_both$virus |>
+  case_match("cardiovirus" ~ "Cardiovirus",
+             "k97_19971" ~ "Kobuvirus",
+             .default = depth_both$virus
+             )
 
@@
 
-depth_both |>
+depth_both$Viral.load <- factor(depth_both$Viral.load, labels = c("Shotgun",
+    "10^{2}", 
+    "10^{3}", 
+    "10^{5}"
+    ))
+
+# depth_both |>
+#   filter(Background != "p6") |>
+#   filter(type == "dedup") |>
+#   filter(mapper == "bowtie2") |>
+#   filter(Viral.load != "NA") |>
+#   ggplot(aes(
+#     x = Viral.load,
+#     y = (mean_depth),
+#     colour = Background
+#   )) +
+#   geom_point() +
+#  theme_few() +
+#   theme(
+#     axis.title.x = element_text(size = 12),
+#     axis.title.y = element_text(size = 12),
+#     legend.title = element_blank(),
+#     legend.position = "none", 
+#     panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
+#     panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
+#   ) +  
+#   facet_grid(Background ~ virus, labeller = as_labeller(coverage_labels)) +
+#   scale_color_manual(values = cols2) +
+#   ylab("Mean read depth") +
+#     xlab("Spike-in viral load (gc/ml)") +
+#    scale_x_log10(labels = scales::trans_format("log10",
+#                                         scales::label_math())) + 
+#   scale_y_log10(labels = scales::trans_format("log10",
+#                                         scales::label_math()))
+
+# replotting S13 in similar way to S14
+depth_both %>%
   filter(Background != "p6") |>
@@
   filter(mapper == "bowtie2") |>
-  filter(Viral.load != "NA") |>
-  ggplot(aes(
-    x = Viral.load,
-    y = (mean_depth),
-    colour = Background
-  )) +
+  # filter(Viral.load != "NA") |>
+  ggplot(aes(x = Sample_id, y = mean_depth, color = Background)) +
   geom_point() +
+  ylab("Mean read depth") +
+  xlab("Sample ID") +
+  facet_grid(virus ~ Viral.load,
+             scales = "free_x",
+             label = label_parsed) +
+  scale_color_manual(values = cols2, labels = c("P1", "P2")) +
  theme_few() +
@@
     legend.title = element_blank(),
-    legend.position = "none", 
+    legend.position = "right", 
     panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
     panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
-  ) +  
-  facet_grid(Background ~ virus, labeller = as_labeller(coverage_labels)) +
-  scale_color_manual(values = cols2) +
-  ylab("Mean read depth") +
-    xlab("Spike-in viral load (gc/ml)") +
-
-   scale_x_log10(labels = scales::trans_format("log10",
-                                              scales::label_math())) + 
+  )+
   scale_y_log10(labels = scales::trans_format("log10",
@@
 
-ggsave("figures/manuscript_figure_2025/PDF/Figure_S13.pdf",
+ggsave("figures/manuscript_figure_2025/PDF/Figure_S13_v2.pdf",
        width=8,
@@
 
-ggsave("figures/manuscript_figure_2025/PNG/Figure_S13.png",
+ggsave("figures/manuscript_figure_2025/PNG/Figure_S13_v2.png",
        width=8,

### 2025-03-29 manuscript_figures.Rmd
Changing virus load axis label to genome copies

@@
   ) +
-    xlab("Viral load (gc/ml)") + 
+    xlab("Genome copies") + 
   ylab("Mean Read Depth") +
@@
 
-## >>>> now redundant? - Figure 1: plots of TE experiment data - read counts and read depths
-
+## Figure S1
+
+### Read counts
+```{r}
 ### Read-in data
-```{r}
+
 #ATCC genomes updated version
@@
   ))
+
+```
+
+### Read depths
+```{r}
+####read depths####
+
+#import and combine read depth files
+
+depth_dedup_bt <-
+  read_tsv(
+    "data_TE/TE_sequencing_experiment_readdepth_per_sample_and_virus_bowtie2_dedup_atcc_ref.tsv"
+  )
+
+depth_nodedup_bt <-
+  read_tsv(
+    "data_TE/TE_sequencing_experiment_readdepth_per_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv"
+  )
+
+depths_bt_all <- rbind(depth_dedup_bt, depth_nodedup_bt)
+
+depths_reads <- 
+  depths_bt_all |>
+  left_join(metadata2, by = "Sample_id") |>
+  mutate(genome_structure = case_when((
+    virus == "Human_adenovirus_40" |
+      virus == "Human_betaherpesvirus"
+  ) ~ "DNA",
+  .default = "RNA"
+  )) |>
+  rename(Viral.load = `Viral load`)
 
@@
   ) +
+  scale_linetype_manual(values=c(2, 1)) +
   ylab("log10(Viral Reads)") +
-  xlab("Log (Viral load)") + 
+  xlab("Genome copies") + 
   scale_x_log10(labels = scales::trans_format("log10", scales::label_math()))
 
-#ggsave("figures/compare_spike_ins_atcc/read_counts.pdf",width=8,height=6)
-
-#normalise by raw read count
-
-read_count_norm1 <- counts_reads_norm |>
+read_count_norm1 <- 
+  counts_reads_norm |>
   filter(Background != "p6") |>
@@
   ) +
+  scale_linetype_manual(values=c(2, 1)) +
   ylab("log10(viral reads/cleaned reads)") +
-  xlab("Log (Viral load)") + 
+  xlab("Genome copies") + 
   scale_x_log10(labels = scales::trans_format("log10", scales::label_math()))
@@
   ) +
+  scale_linetype_manual(values=c(2, 1)) +
   ylab("Log (viral reads/cleaned reads/genome length)") +
-  xlab("Log (Viral load)") + 
+  xlab("Genome copies") + 
   scale_x_log10(labels = scales::trans_format("log10", scales::label_math()))
-
-```
-
-### plot just the two lefthand panels from the original figure
-```{r, warning=FALSE, echo=FALSE, message=FALSE, fig.height=10, fig.width=8}
-
-ggarrange(read_count,
-          read_count_norm2,
-          nrow = 2,
-          common.legend = TRUE)
-
-#ggsave("figures/compare_spike_ins_atcc/log_read_count_depth_2panels.png",width=10,height=7)
-
-```
-
-### plot viral read counts (non normalised)
-```{r}
-
-##deduplicated only
-
-read_count_dedup <- 
-  counts_reads_norm |>
-  filter(Background != "p6") |>
-  filter(Background != "control") |>
-  filter(type == "dedup_TE") |>
-  ggplot(aes(
-    x = Viral.load,
-    y = log10(matched),
-    colour = virus
-  )) +
-  geom_point() +
-  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE, method = "lm") +
-  theme_few() +
-  theme(
-    axis.title.x = element_text(size = 12),
-    axis.text.x = element_text(angle = 0, hjust = 1),
-    axis.title.y = element_text(size = 12),
-    legend.title = element_blank(),
-    legend.position = "top", 
-    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
-    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
-  ) +
-  scale_color_manual(
-    values = cols,
-    labels = c(
-      "Human adenovirus 40",
-      "Human betaherpesvirus",
-      "Human respiratory syncytial virus",
-      "Influenza B virus",
-      "Mammalian orthoreovirus 3",
-      "Zika virus"
-    )
-  ) +
-  ylab("Log (Viral Reads)") +
-  xlab("Log (Viral load)") + 
-  scale_x_log10(labels = scales::trans_format("log10", scales::label_math()))
-
-#normalise by raw read count and genome length
-
-read_count_norm2_dedup <- 
-  counts_reads_norm |>
-  filter(Background != "p6") |>
-  filter(Background != "control") |>
-  filter(type == "dedup_TE") |>
-  ggplot(aes(
-    x = Viral.load,
-    y = log10(norm_counts2),
-    colour = virus
-  )) +
-  geom_point() +
-  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE, method = "lm") +
-  theme_few() +
-  theme(
-    axis.title.x = element_text(size = 12),
-    axis.text.x = element_text(angle = 0, hjust = 1),
-    axis.title.y = element_text(size = 12),
-    legend.title = element_blank(),
-    legend.position = "top", 
-    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
-    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
-  ) +
-  scale_color_manual(
-    values = cols,
-    labels = c(
-      "Human adenovirus 40",
-      "Human betaherpesvirus",
-      "Human respiratory syncytial virus",
-      "Influenza B virus",
-      "Mammalian orthoreovirus 3",
-      "Zika virus"
-    )
-  ) +
-  ylab("Log (viral reads/cleaned reads/genome length)") +
-  xlab("Log (Viral load)") + 
-  scale_x_log10(labels = scales::trans_format("log10", scales::label_math()))
-```
-
-### ***Final plot***
-```{r, warning=FALSE, message=FALSE, echo=FALSE, fig.width=8, fig.height=10}
-#plot just the two lefthand panels from the original figure
-
-ggarrange(
-  read_count_dedup,
-  read_count_norm2_dedup,
-  nrow = 2,
-  common.legend = TRUE,
-  align = "hv"
-)
-
-# ggsave("figures/compare_spike_ins_atcc/log_read_count_depth_2panels_dedup.png",
-#        width=10,
-#        height=7)
-
-ggsave("figures/manuscript_figure_2025/PDF/Figure_S1.pdf",
-       width=7,
-       height=8)
-
-ggsave("figures/manuscript_figure_2025/PNG/Figure_S1.png",
-       width=7,
-       height=8)
-
-```
-
-### --------------------
-
-## Figure S1
-
-### Read depths
-```{r}
-####read depths####
-
-#import and combine read depth files
-
-depth_dedup_bt <-
-  read_tsv(
-    "data_TE/TE_sequencing_experiment_readdepth_per_sample_and_virus_bowtie2_dedup_atcc_ref.tsv"
-  )
-
-depth_nodedup_bt <-
-  read_tsv(
-    "data_TE/TE_sequencing_experiment_readdepth_per_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv"
-  )
-
-depths_bt_all <- rbind(depth_dedup_bt, depth_nodedup_bt)
-
-depths_reads <- 
-  depths_bt_all |>
-  left_join(metadata2, by = "Sample_id") |>
-  mutate(genome_structure = case_when((
-    virus == "Human_adenovirus_40" |
-      virus == "Human_betaherpesvirus"
-  ) ~ "DNA",
-  .default = "RNA"
-  )) |>
-  rename(Viral.load = `Viral load`)
-
-```
-
-### Make plots
-```{r}
 
@@
   ) +
+  scale_linetype_manual(values=c(2, 1)) +
   ylab("Log (mean read depth)") +
-  xlab("Log (Viral load)") + 
+  xlab("Genome copies") + 
   scale_x_log10(labels = scales::trans_format("log10", scales::label_math()))
@@
   theme(
-    axis.title.x = element_blank(),
+    # axis.title.x = element_blank(),
     # axis.text.x = element_blank(),
@@
   ) +
-  ylab("Log (Viral Reads)") + xlab("Log (Viral load)") +
+  ylab("Log (Viral Reads)") + xlab("Genome copies") +
   scale_color_manual(values = cols3, labels = c("Pool 1", "Pool 2")) + 
@@
   ) +
-  ylab("log10(viral reads/cleaned reads)") + xlab("Log (Viral load)") +
+  ylab("Log (viral reads/cleaned reads)") + xlab("Genome copies") +
   scale_color_manual(values = cols3, labels = c("Pool 1", "Pool 2")) + 
@@
   ) +
-  ylab("log10(viral reads/cleaned reads/genome length)") + xlab("Log (Viral load)") +
+  ylab("Log (viral reads/cleaned reads/genome length)") + xlab("Genome copies") +
   scale_color_manual(values = cols3, labels = c("Pool 1", "Pool 2")) + 
@@
   ylab("Log (mean read depth)") +
-  xlab("Log (Viral load)") + 
+  xlab("Genome copies") + 
   scale_x_discrete(labels = c(bquote(10^{2}), bquote(10^{3}), bquote(10^{5}))) +
@@
 ggsave("figures/manuscript_figure_2025/PDF/Figure_S2.pdf",
-       width = 10,
-       height = 7)
+       width = 12,
+       height = 8)
 
 ggsave("figures/manuscript_figure_2025/PNG/Figure_S2.png",
-       width = 10,
-       height = 7)
+       width = 12,
+       height = 8)
 
@@
   scale_y_log10(labels = scales::trans_format("log10", scales::label_math())) + 
-    scale_x_log10(labels = scales::trans_format("log10", scales::label_math())) + xlab("Viral load") +
+    scale_x_log10(labels = scales::trans_format("log10", scales::label_math())) + xlab("Genome copies") +
   # ggtitle("All") +
@@
   scale_y_log10(labels = scales::trans_format("log10", scales::label_math())) + 
-    scale_x_log10(labels = scales::trans_format("log10", scales::label_math())) + xlab("Viral load") +
+    scale_x_log10(labels = scales::trans_format("log10", scales::label_math())) + xlab("Genome copies") +
   # ggtitle("Viral reads") +
@@
   scale_y_log10(labels = scales::trans_format("log10", scales::label_math())) + 
-    scale_x_log10(labels = scales::trans_format("log10", scales::label_math())) + xlab("Viral load") +
+    scale_x_log10(labels = scales::trans_format("log10", scales::label_math())) + xlab("Genome copies") +
   ggtitle("Spike-in Viral Reads") +
@@
   ) +
-    scale_x_log10(labels = scales::trans_format("log10", scales::label_math())) + xlab("Viral load") +
+    scale_x_log10(labels = scales::trans_format("log10", scales::label_math())) + xlab("Genome copies") +
   ggtitle("Spike-in viral reads") +
@@
   ) +
- scale_x_log10(labels = scales::trans_format("log10", scales::label_math())) + xlab("Viral load") +
+ scale_x_log10(labels = scales::trans_format("log10", scales::label_math())) + xlab("Genome copies") +
   # ggtitle("Viral reads") +

### 2025-03-29 manuscript_figures.Rmd
Changing virus load axis label to genome copies and background to M1 and M2

@@
   scale_linetype_manual(values=c(2, 1)) +
-  ylab("log10(Viral Reads)") +
+  ylab(" Log (Viral Reads)") +
+  scale_linetype_manual(values=c("dotted", "solid")) + 
+  ylab("Log (Viral Reads)") +
   xlab("Genome copies") + 
@@
   ) +
-  scale_linetype_manual(values=c(2, 1)) +
-  ylab("log10(viral reads/cleaned reads)") +
+  scale_linetype_manual(values=c(3, 1)) +
+  ylab("Log (viral reads/cleaned reads)") +
   xlab("Genome copies") + 
@@
   ) +
-  scale_linetype_manual(values=c(2, 1)) +
+  scale_linetype_manual(values=c(3, 1)) +
   ylab("Log (viral reads/cleaned reads/genome length)") +
@@
   ) +
-  scale_linetype_manual(values=c(2, 1)) +
+  scale_linetype_manual(values=c(3, 1)) +
   ylab("Log (mean read depth)") +
@@
   ylab("Total number of reads") +
-  scale_color_manual(values = cols2, labels = c("P1", "P2"))
+  scale_color_manual(values = cols2, labels = c("M1", "M2"))
 
@@
   ylab("Number of viral reads") +
-  scale_color_manual(values = cols2, labels = c("P1", "P2"))
+  scale_color_manual(values = cols2, labels = c("M1", "M2"))
 
@@
   ylab("Spike-in Viral Reads") +
-  scale_color_manual(values = cols2, labels = c("P1", "P2"))
+  scale_color_manual(values = cols2, labels = c("M1", "M2"))
 
@@
   ylab("Proportion") +
-  scale_color_manual(values = cols2, labels = c("P1", "P2"))
+  scale_color_manual(values = cols2, labels = c("M1", "M2"))
 
@@
   ylab("Proportion of viral reads") +
-  scale_color_manual(values = cols2, labels = c("P1", "P2")) + scale_y_continuous(limits=c(0,1.0))
+  scale_color_manual(values = cols2, labels = c("M1", "M2")) + scale_y_continuous(limits=c(0,1.0))
 
@@
   "nodedup_TE" = "Non-Deduplicated",
-  "p2" = "P1",
-  "p8" = "P2"
+  "p2" = "M1",
+  "p8" = "M2"
 )
@@
   scale_x_log10(labels = scales::label_log10()) +
-  ylab("log10(Viral Reads)")
+  ylab("Log (Viral Reads)")
 
@@
   scale_x_log10(label = scales::label_log10()) +
-  ylab("log10(viral reads/cleaned reads)")
+  ylab("Log (viral reads/cleaned reads)")
 
@@
   scale_x_log10(labels = scales::label_log10()) +
-  ylab("log10(viral reads/cleaned reads/genome length)")
+  ylab("Log (viral reads/cleaned reads/genome length)")
 
@@
   scale_x_log10(label = scales::label_log10()) +
-  ylab("log10(mean read depth)")
+  ylab("Log (mean read depth)")
 
@@
   ) +
-  scale_color_manual(values = cols2, labels = c("P1", "P2")) +
+  scale_color_manual(values = cols2, labels = c("M1", "M2")) +
   scale_x_discrete(
@@
                     "100000" = "1e+05",
-                    "p2" = "P1",
-                    "p8"= "P2")
+                    "p2" = "M1",
+                    "p8"= "M2")
 
@@
   scale_y_log10() +
-  scale_fill_manual(values = cols2, labels = c("P1", "P2"))
+  scale_fill_manual(values = cols2, labels = c("M1", "M2"))
 ```
@@
   scale_y_log10() +
-  scale_fill_manual(values = cols2, labels = c("P1", "P2"))
+  scale_fill_manual(values = cols2, labels = c("M1", "M2"))
 
@@
   ylab("Coverage") +
-  ggtitle("Influenza B virus (Background 1 deduplicated)") +
+  ggtitle("Influenza B virus (Background M1 deduplicated)") +
   guides(fill = guide_legend(title = "Viral load")) +
@@
   ylab("Coverage") +
-  ggtitle("Influenza B virus (Background 2 deduplicated)") +
+  ggtitle("Influenza B virus (Background M2 deduplicated)") +
   guides(fill = guide_legend(title = "Viral load")) +
@@
   facet_grid(seg ~ Sample_id) +
-  ylab("log10(Coverage)") +
-  ggtitle("Mammalian orthoreovirus 3 (Background 1 deduplicated)") +
+  ylab("Coverage") +
+  ggtitle("Mammalian orthoreovirus 3 (Background M1 deduplicated)") +
   guides(fill = guide_legend(title = "Viral load")) +
@@
   facet_grid(seg ~ Sample_id) +
-  ylab("log10(Coverage)") +
-  ggtitle("Mammalian orthoreovirus 3 (Background 2 deduplicated)") +
+  ylab("Coverage") +
+  ggtitle("Mammalian orthoreovirus 3 (Background M2 deduplicated)") +
   guides(fill = guide_legend(title = "Viral load")) +
@@
 facet_names <- c(
-  "Msp_p2" = "P1",
-  "Msp_p8" = "P2",
+  "Msp_p2" = "M1",
+  "Msp_p8" = "M2",
   "off_target" = "Background",
@@
   ylab("Viral Reads") + 
-  xlab("Viral load (gc/ml)") + 
+  xlab("Genome copies") + 
   scale_color_manual(values = cols) +
-  scale_linetype_manual(values=c("solid", "dotted")) + 
+  scale_linetype_manual(values=c("dotted", "solid")) + 
   scale_x_log10(labels = scales::trans_format("log10", scales::label_math())) + 
@@
                                               scales::label_math())) +
-    scale_linetype_manual(values=c("solid", "dotted"))+ 
+  scale_linetype_manual(values=c("dotted", "solid")) + 
   ylab("Proportion of viral reads") +
-  xlab("Viral load (gc/ml)") + 
+  xlab("Genome copies") + 
   scale_color_manual(values = cols) +
@@
                                               scales::label_math())) +
-    scale_linetype_manual(values=c("solid", "dotted"))+ 
+  scale_linetype_manual(values=c("dotted", "solid")) + 
   ylab("Proportion of total reads") +
-  xlab("Viral load (gc/ml)") + 
+  xlab("Genome copies") + 
   scale_color_manual(values = cols) +
@@
   ylab("Viral Reads") +
-  xlab("Viral load (gc/ml)") + 
+  xlab("Genome copies") + 
   scale_x_log10(labels = scales::trans_format("log10",
@@
   ylab("Mean read depth") +
-  xlab("Viral load (gc/ml)") + 
+  xlab("Genome copies") + 
   scale_x_log10(labels = scales::trans_format("log10",
@@
 depth_both$Sample_id <- depth_both$Sample_id |>
-  case_match("RNA-Msp-p2" ~ "P1",
-             "RNA-Msp-p8" ~ "P2",
-             "RNA-Msp-p6" ~ "P3",
+  case_match("RNA-Msp-p2" ~ "M1",
+             "RNA-Msp-p8" ~ "M2",
+             "RNA-Msp-p6" ~ "M3",
              .default = depth_both$Sample_id
@@
   "cardiovirus" = "Cardiovirus",
-  "p2" = "P1",
-  "p8" = "P2"
+  "p2" = "M1",
+  "p8" = "M2"
 )
@@
              label = label_parsed) +
-  scale_color_manual(values = cols2, labels = c("P1", "P2")) +
+  scale_color_manual(values = cols2, labels = c("M1", "M2")) +
  theme_few() +
@@
 kobu_per_site$Sample_id <- kobu_per_site$Sample_id %>%
-  case_match("RNA-Msp-p2" ~ "P1",
-             "RNA-Msp-p8" ~ "P2",
+  case_match("RNA-Msp-p2" ~ "M1",
+             "RNA-Msp-p8" ~ "M2",
              .default = kobu_per_site$Sample_id)
@@
 kobu_per_site$Background <- kobu_per_site$Background %>%
-  case_match("p2" ~ "P1",
-             "p8" ~ "P2",
+  case_match("p2" ~ "M1",
+             "p8" ~ "M2",
              .default = kobu_per_site$Background)
@@
 cardio_TE_polyomics$Sample_id <- cardio_TE_polyomics$Sample_id %>%
-  case_match("RNA-Msp-p2" ~ "P1",
-             "RNA-Msp-p8" ~ "P2",
+  case_match("RNA-Msp-p2" ~ "M1",
+             "RNA-Msp-p8" ~ "M2",
              .default = cardio_TE_polyomics$Sample_id)
@@
 cardio_TE_polyomics$Background <- cardio_TE_polyomics$Background %>%
-  case_match("S18" ~ "P1",
-             "S43" ~ "P2",
-             "p2" ~ "P1",
-             "p8" ~ "P2",
+  case_match("S18" ~ "M1",
+             "S43" ~ "M2",
+             "p2" ~ "M1",
+             "p8" ~ "M2",
              .default = cardio_TE_polyomics$Background)
@@
   "P" = "P",
-  "ME_P1" = "P1",
-  "ME_P2" = "P2"
+  "ME_P1" = "M1",
+  "ME_P2" = "M2"
 )
@@
   facet_wrap(Viral.load ~ Sample_id, labeller = label_parsed) +
-  scale_fill_manual(values = cols2, labels = c("P1", "P2")) +
+  scale_fill_manual(values = cols2, labels = c("M1", "M2")) +
   ylab("Coverage") +
@@
   facet_wrap(Viral.load ~ Sample_id, labeller = label_parsed) +
-  scale_fill_manual(values = cols2, labels = c("P1", "P2")) +
+  scale_fill_manual(values = cols2, labels = c("M1", "M2")) +
   theme_few() +

### 2025-03-29 manuscript_figures.Rmd
update notebook

@@
   ) +
-  scale_linetype_manual(values=c(2, 1)) +
-  ylab(" Log (Viral Reads)") +
   scale_linetype_manual(values=c("dotted", "solid")) + 
@@
   ) +
-  scale_linetype_manual(values=c(3, 1)) +
+  scale_linetype_manual(values=c("dotted", "solid")) + 
   ylab("Log (viral reads/cleaned reads)") +
@@
   ) +
-  scale_linetype_manual(values=c(3, 1)) +
+  scale_linetype_manual(values=c("dotted", "solid")) + 
   ylab("Log (viral reads/cleaned reads/genome length)") +
@@
   ) +
-  scale_linetype_manual(values=c(3, 1)) +
+  scale_linetype_manual(values=c(2, 1)) +
   ylab("Log (mean read depth)") +

### 2025-04-13 manuscript_figures.Rmd
Update Figures S1 and S2

@@
 ### --------------------
-
 ## Figure S1
-
-### Read counts
-```{r}
-### Read-in data
-
-#ATCC genomes updated version
-#October 2024
-
-#viral load vs read count normalised using different methods, comparing deduplicated and non-deduplicated
-#use bowtie2 data
-
-####read counts/normalised read counts####
-
-#metadata
-
-metadata <- read.csv("metadata/sampleIDs_TESpikeIn.csv", header = TRUE)
-
-metadata2 <- metadata |>
-  select(Sample.ID, Number.of.read.pairs..quality.adaptor.trimmed.) |>
-  rename(Sample_id = Sample.ID) |>
-  rename(QC_reads = Number.of.read.pairs..quality.adaptor.trimmed.)
-
-#import and combine read count files
-#deduplicated and non-deduplicated
-
-counts_dedup_bt <-
-  read.table(
-    "data_TE/TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_dedup_atcc_ref.tsv",
-    sep = "\t",
-    header = TRUE
-  )
-
-counts_nodedup_bt <-
-  read.table(
-    "data_TE/TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv",
-    sep = "\t",
-    header = TRUE
-  )
-
-counts_bt_all <- rbind(counts_dedup_bt, counts_nodedup_bt)
-
-counts_reads <- left_join(counts_bt_all, metadata2, by = "Sample_id")
-
-#normalise by both raw read count and genome length - should be the same as mean read depth
-
-counts_reads_norm <- counts_reads |>
-  mutate(norm_counts1 = matched / QC_reads) |>
-  mutate(norm_counts2 = matched / QC_reads / length) |>
-  mutate(norm_counts3 = matched / length) |>
-  mutate(genome_structure = case_when((
-    virus == "Human_adenovirus_40" |
-      virus == "Human_betaherpesvirus"
-  ) ~ "DNA",
-  .default = "RNA"
-  ))
-
-```
-
-### Read depths
-```{r}
-####read depths####
-
-#import and combine read depth files
-
-depth_dedup_bt <-
-  read_tsv(
-    "data_TE/TE_sequencing_experiment_readdepth_per_sample_and_virus_bowtie2_dedup_atcc_ref.tsv"
-  )
-
-depth_nodedup_bt <-
-  read_tsv(
-    "data_TE/TE_sequencing_experiment_readdepth_per_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv"
-  )
-
-depths_bt_all <- rbind(depth_dedup_bt, depth_nodedup_bt)
-
-depths_reads <- 
-  depths_bt_all |>
-  left_join(metadata2, by = "Sample_id") |>
-  mutate(genome_structure = case_when((
-    virus == "Human_adenovirus_40" |
-      virus == "Human_betaherpesvirus"
-  ) ~ "DNA",
-  .default = "RNA"
-  )) |>
-  rename(Viral.load = `Viral load`)
-
-```
-
-### Format data
-```{r}
-#change labels in facet plots
-
-facet_names <- c("dedup_TE" = "Deduplicated",
-                 "nodedup_TE" = "Non-Deduplicated")
-
-cols <- c("#4477AA",
-          "#66CCEE",
-          "#228833",
-          "#CCBB44",
-          "#EE6677",
-          "#AA3377")
-```
-
-### Make plots
-```{r}
-
-#plot viral read counts (non normalised)
-
-read_count <- 
-  counts_reads_norm |>
-  filter(Background != "p6") |>
-  filter(Background != "control") |>
-  ggplot(aes(
-    x = Viral.load,
-    y = log10(matched),
-    colour = virus
-  )) +
-  geom_point() +
-  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
-  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
-  theme_few() +
-  theme(
-    axis.title.x = element_text(size = 12),
-    axis.text.x = element_text(angle = 0, hjust = 1),
-    axis.title.y = element_text(size = 12),
-    legend.title = element_blank(),
-    legend.position = "top", 
-    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
-    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
-  ) +
-  scale_color_manual(
-    values = cols,
-    labels = c(
-      "Human adenovirus 40",
-      "Human betaherpesvirus",
-      "Human respiratory syncytial virus",
-      "Influenza B virus",
-      "Mammalian orthoreovirus 3",
-      "Zika virus"
-    )
-  ) +
-  scale_linetype_manual(values=c("dotted", "solid")) + 
-  ylab("Log (Viral Reads)") +
-  xlab("Genome copies") + 
-  scale_x_log10(labels = scales::trans_format("log10", scales::label_math()))
-
-read_count_norm1 <- 
-  counts_reads_norm |>
-  filter(Background != "p6") |>
-  filter(Background != "control") |>
-  ggplot(aes(
-    x = Viral.load,
-    y = log10(norm_counts1),
-    colour = virus
-  )) +
-  geom_point() +
-  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
-  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
-  theme_few() +
-  theme(
-    axis.title.x = element_text(size = 12),
-    axis.text.x = element_text(angle = 0, hjust = 1),
-    axis.title.y = element_text(size = 12),
-    legend.title = element_blank(),
-    legend.position = "top", 
-    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
-    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
-  ) +
-  scale_color_manual(
-    values = cols,
-    labels = c(
-      "Human adenovirus 40",
-      "Human betaherpesvirus",
-      "Human respiratory syncytial virus",
-      "Influenza B virus",
-      "Mammalian orthoreovirus 3",
-      "Zika virus"
-    )
-  ) +
-  scale_linetype_manual(values=c("dotted", "solid")) + 
-  ylab("Log (viral reads/cleaned reads)") +
-  xlab("Genome copies") + 
-  scale_x_log10(labels = scales::trans_format("log10", scales::label_math()))
-
-#normalise by raw read count and genome length
-
-read_count_norm2 <- 
-  counts_reads_norm |>
-  filter(Background != "p6") |>
-  filter(Background != "control") |>
-  ggplot(aes(
-    x = (Viral.load),
-    y = log10(norm_counts2),
-    colour = virus
-  )) +
-  geom_point() +
-  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
-  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
-  theme_few() +
-  theme(
-    axis.title.x = element_text(size = 12),
-    axis.text.x = element_text(angle = 0, hjust = 1),
-    axis.title.y = element_text(size = 12),
-    legend.title = element_blank(),
-    legend.position = "top", 
-    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
-    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
-  ) +
-  scale_color_manual(
-    values = cols,
-    labels = c(
-      "Human adenovirus 40",
-      "Human betaherpesvirus",
-      "Human respiratory syncytial virus",
-      "Influenza B virus",
-      "Mammalian orthoreovirus 3",
-      "Zika virus"
-    )
-  ) +
-  scale_linetype_manual(values=c("dotted", "solid")) + 
-  ylab("Log (viral reads/cleaned reads/genome length)") +
-  xlab("Genome copies") + 
-  scale_x_log10(labels = scales::trans_format("log10", scales::label_math()))
-
-read_depths <- 
-  depths_reads |>
-  filter(Background != "p6") |>
-  filter(Background != "control") |>
-  ggplot(aes(
-    x = Viral.load,
-    y = log10(mean_depth),
-    colour = virus
-  )) +
-  geom_point() +
-  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
-  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
-  theme_few() +
-  theme(
-    axis.title.x = element_text(size = 12),
-    axis.text.x = element_text(angle = 0, hjust = 1),
-    axis.title.y = element_text(size = 12),
-    legend.title = element_blank(),
-    legend.position = "top", 
-    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
-    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
-  ) +
-  scale_color_manual(
-    values = cols,
-    labels = c(
-      "Human adenovirus 40",
-      "Human betaherpesvirus",
-      "Human respiratory syncytial virus",
-      "Influenza B virus",
-      "Mammalian orthoreovirus 3",
-      "Zika virus"
-    )
-  ) +
-  scale_linetype_manual(values=c(2, 1)) +
-  ylab("Log (mean read depth)") +
-  xlab("Genome copies") + 
-  scale_x_log10(labels = scales::trans_format("log10", scales::label_math()))
-
-#ggsave("figures/compare_spike_ins_atcc/mean_depth.pdf",width=8,height=6)
-
-```
-
-### ***Final plot***
-```{r, warning = FALSE, echo = FALSE, message = FALSE}
-
-ggarrange(
-  read_count,
-  read_count_norm1,
-  read_count_norm2,
-  read_depths,
-  nrow = 2,
-  ncol = 2,
-  common.legend = TRUE,
-  align = "hv"
-)
-
-#ggsave("figures/compare_spike_ins_atcc/log_read_count_depth.png",width=10,height=7)
-
-#ggsave("figures/manuscript_figures_pdf/FigureS1.pdf",width=10,height=7)
-
-#apparently normalised read count using method 2 should not actually be the same
-
-ggsave("figures/manuscript_figure_2025/PDF/Figure_S1.pdf",
-       width = 12,
-       height = 8)
-
-ggsave("figures/manuscript_figure_2025/PNG/Figure_S1.png",
-       width = 12,
-       height = 8)
-
-```
-
-### --------------------
-
-## Figure S2
 
@@
     x = as.character(Viral.load),
-    y = log10(matched),
+    y = log10(norm_counts1),
     colour = Pool
@@
     x = as.character(Viral.load),
-    y = log10(matched),
+    y = norm_counts2,
     colour = Pool
@@
   pool_counts_sum,
-  pool_norm1_sum,
-  pool_norm2_sum,
   pool_depths_sum,
+  nrow = 1,
+  ncol = 2,
+  common.legend = TRUE, 
+  align = "hv"
+)
+
+#ggsave("figures/compare_spike_ins_atcc/pool_compare_viruses_sum.png",width=10,height=7)
+
+ggsave("figures/manuscript_figure_2025/PDF/Figure_S1.pdf",
+       width = 12,
+       height = 4)
+
+ggsave("figures/manuscript_figure_2025/PNG/Figure_S1.png",
+       width = 12,
+       height = 4)
+
+```
+
+### --------------------
+
+## Figure S2
+
+### Read counts
+```{r}
+### Read-in data
+
+#ATCC genomes updated version
+#October 2024
+
+#viral load vs read count normalised using different methods, comparing deduplicated and non-deduplicated
+#use bowtie2 data
+
+####read counts/normalised read counts####
+
+#metadata
+
+metadata <- read.csv("metadata/sampleIDs_TESpikeIn.csv", header = TRUE)
+
+metadata2 <- metadata |>
+  select(Sample.ID, Number.of.read.pairs..quality.adaptor.trimmed.) |>
+  rename(Sample_id = Sample.ID) |>
+  rename(QC_reads = Number.of.read.pairs..quality.adaptor.trimmed.)
+
+#import and combine read count files
+#deduplicated and non-deduplicated
+
+counts_dedup_bt <-
+  read.table(
+    "data_TE/TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_dedup_atcc_ref.tsv",
+    sep = "\t",
+    header = TRUE
+  )
+
+counts_nodedup_bt <-
+  read.table(
+    "data_TE/TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv",
+    sep = "\t",
+    header = TRUE
+  )
+
+counts_bt_all <- rbind(counts_dedup_bt, counts_nodedup_bt)
+
+counts_reads <- left_join(counts_bt_all, metadata2, by = "Sample_id")
+
+#normalise by both raw read count and genome length - should be the same as mean read depth
+
+counts_reads_norm <- counts_reads |>
+  mutate(norm_counts1 = matched / QC_reads) |>
+  mutate(norm_counts2 = matched / QC_reads / length) |>
+  mutate(norm_counts3 = matched / length) |>
+  mutate(genome_structure = case_when((
+    virus == "Human_adenovirus_40" |
+      virus == "Human_betaherpesvirus"
+  ) ~ "DNA",
+  .default = "RNA"
+  ))
+
+```
+
+### Read depths
+```{r}
+####read depths####
+
+#import and combine read depth files
+
+depth_dedup_bt <-
+  read_tsv(
+    "data_TE/TE_sequencing_experiment_readdepth_per_sample_and_virus_bowtie2_dedup_atcc_ref.tsv"
+  )
+
+depth_nodedup_bt <-
+  read_tsv(
+    "data_TE/TE_sequencing_experiment_readdepth_per_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv"
+  )
+
+depths_bt_all <- rbind(depth_dedup_bt, depth_nodedup_bt)
+
+depths_reads <- 
+  depths_bt_all |>
+  left_join(metadata2, by = "Sample_id") |>
+  mutate(genome_structure = case_when((
+    virus == "Human_adenovirus_40" |
+      virus == "Human_betaherpesvirus"
+  ) ~ "DNA",
+  .default = "RNA"
+  )) |>
+  rename(Viral.load = `Viral load`)
+
+```
+
+### Format data
+```{r}
+#change labels in facet plots
+
+facet_names <- c("dedup_TE" = "Deduplicated",
+                 "nodedup_TE" = "Non-Deduplicated")
+
+cols <- c("#4477AA",
+          "#66CCEE",
+          "#228833",
+          "#CCBB44",
+          "#EE6677",
+          "#AA3377")
+```
+
+### Make plots
+```{r}
+
+#plot viral read counts (non normalised)
+
+read_count <- 
+  counts_reads_norm |>
+  filter(Background != "p6") |>
+  filter(Background != "control") |>
+  ggplot(aes(
+    x = Viral.load,
+    y = log10(matched),
+    colour = virus
+  )) +
+  geom_point() +
+  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
+  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
+  theme_few() +
+  theme(
+    axis.title.x = element_text(size = 12),
+    axis.text.x = element_text(angle = 0, hjust = 1),
+    axis.title.y = element_text(size = 12),
+    legend.title = element_blank(),
+    legend.position = "top", 
+    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
+    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
+  ) +
+  scale_color_manual(
+    values = cols,
+    labels = c(
+      "Human adenovirus 40",
+      "Human betaherpesvirus",
+      "Human respiratory syncytial virus",
+      "Influenza B virus",
+      "Mammalian orthoreovirus 3",
+      "Zika virus"
+    )
+  ) +
+  scale_linetype_manual(values=c("dotted", "solid")) + 
+  ylab("Log (Viral Reads)") +
+  xlab("Genome copies") + 
+  scale_x_log10(labels = scales::trans_format("log10", scales::label_math()))
+
+read_count_norm1 <- 
+  counts_reads_norm |>
+  filter(Background != "p6") |>
+  filter(Background != "control") |>
+  ggplot(aes(
+    x = Viral.load,
+    y = log10(norm_counts1),
+    colour = virus
+  )) +
+  geom_point() +
+  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
+  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
+  theme_few() +
+  theme(
+    axis.title.x = element_text(size = 12),
+    axis.text.x = element_text(angle = 0, hjust = 1),
+    axis.title.y = element_text(size = 12),
+    legend.title = element_blank(),
+    legend.position = "top", 
+    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
+    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
+  ) +
+  scale_color_manual(
+    values = cols,
+    labels = c(
+      "Human adenovirus 40",
+      "Human betaherpesvirus",
+      "Human respiratory syncytial virus",
+      "Influenza B virus",
+      "Mammalian orthoreovirus 3",
+      "Zika virus"
+    )
+  ) +
+  scale_linetype_manual(values=c("dotted", "solid")) + 
+  ylab("Log (viral reads/cleaned reads)") +
+  xlab("Genome copies") + 
+  scale_x_log10(labels = scales::trans_format("log10", scales::label_math()))
+
+#normalise by raw read count and genome length
+
+read_count_norm2 <- 
+  counts_reads_norm |>
+  filter(Background != "p6") |>
+  filter(Background != "control") |>
+  ggplot(aes(
+    x = (Viral.load),
+    y = log10(norm_counts2),
+    colour = virus
+  )) +
+  geom_point() +
+  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
+  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
+  theme_few() +
+  theme(
+    axis.title.x = element_text(size = 12),
+    axis.text.x = element_text(angle = 0, hjust = 1),
+    axis.title.y = element_text(size = 12),
+    legend.title = element_blank(),
+    legend.position = "top", 
+    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
+    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
+  ) +
+  scale_color_manual(
+    values = cols,
+    labels = c(
+      "Human adenovirus 40",
+      "Human betaherpesvirus",
+      "Human respiratory syncytial virus",
+      "Influenza B virus",
+      "Mammalian orthoreovirus 3",
+      "Zika virus"
+    )
+  ) +
+  scale_linetype_manual(values=c("dotted", "solid")) + 
+  ylab("Log (viral reads/cleaned reads/genome length)") +
+  xlab("Genome copies") + 
+  scale_x_log10(labels = scales::trans_format("log10", scales::label_math()))
+
+read_depths <- 
+  depths_reads |>
+  filter(Background != "p6") |>
+  filter(Background != "control") |>
+  ggplot(aes(
+    x = Viral.load,
+    y = log10(mean_depth),
+    colour = virus
+  )) +
+  geom_point() +
+  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
+  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
+  theme_few() +
+  theme(
+    axis.title.x = element_text(size = 12),
+    axis.text.x = element_text(angle = 0, hjust = 1),
+    axis.title.y = element_text(size = 12),
+    legend.title = element_blank(),
+    legend.position = "top", 
+    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
+    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
+  ) +
+  scale_color_manual(
+    values = cols,
+    labels = c(
+      "Human adenovirus 40",
+      "Human betaherpesvirus",
+      "Human respiratory syncytial virus",
+      "Influenza B virus",
+      "Mammalian orthoreovirus 3",
+      "Zika virus"
+    )
+  ) +
+  scale_linetype_manual(values=c(2, 1)) +
+  scale_linetype_manual(values=c("dotted", "solid")) + 
+  ylab("Log (mean read depth)") +
+  xlab("Genome copies") + 
+  scale_x_log10(labels = scales::trans_format("log10", scales::label_math()))
+
+#ggsave("figures/compare_spike_ins_atcc/mean_depth.pdf",width=8,height=6)
+
+```
+
+### ***Final plot***
+```{r, warning = FALSE, echo = FALSE, message = FALSE}
+
+ggarrange(
+  read_count,
+  read_count_norm1,
+  read_count_norm2,
+  read_depths,
   nrow = 2,
@@
 
-#ggsave("figures/compare_spike_ins_atcc/pool_compare_viruses_sum.png",width=10,height=7)
+#ggsave("figures/compare_spike_ins_atcc/log_read_count_depth.png",width=10,height=7)
+
+#ggsave("figures/manuscript_figures_pdf/FigureS1.pdf",width=10,height=7)
+
+#apparently normalised read count using method 2 should not actually be the same
 
@@
   ) |>
-  mutate(genome_structure = case_when((genus == "Mastadenovirus" |
-                                         genus == "Cytomegalovirus") ~ "DNA",
+  mutate(genome_structure = 
+           case_when((genus == "Mastadenovirus" | 
+                        genus == "Cytomegalovirus" | 
+                        genus == "Cyclovirus" ~ "DNA"),
                      .default = "RNA"

### 2025-04-13 manuscript_figures.Rmd
Update figure S8

@@
   filter(type == "dedup_TE") |>
-  filter(Background == "p2") |>
+  filter(Background == "p8") |>
   rename(Viral.load = `Viral load`) |>

### 2025-04-14 manuscript_figures.Rmd
updating figure 1 to have dashed and solid lines
figure 1 update colors

@@
 
+lm_combined$virus_type =c('DNA','DNA','RNA','RNA','RNA','RNA')
 ```
@@
              mapping = aes(Viral.load, mean_depth, color = virus)) +
-  geom_abline(data = lm_combined, aes(intercept = `(Intercept)`, slope = log_viral_load), linetype="dotted") +
+  geom_abline(data = lm_combined, aes(intercept = `(Intercept)`, slope = log_viral_load, linetype=virus_type, color = virus)) +
   geom_text(data = lm_combined,
@@
   guides(color = FALSE) +
+  scale_linetype_manual(values=c("dotted", "solid"))+
+  guides(linetype = FALSE)+
   scale_x_log10(labels = scales::trans_format("log10", scales::label_math())) + 
@@
     axis.title.x = element_text(size = 12),
-    axis.text.x = element_text(angle = 30, hjust = 1),
+    axis.text.x = element_text(angle = 20, hjust = 1),
     axis.title.y = element_text(size = 12),

### 2025-04-14 manuscript_figures.Rmd
changed lines to lm

@@
   ) +
-  scale_x_log10(labels = scales::label_log10()) +
+ scale_x_log10(labels = label_log()) +  # scales::label_log10()
   ylab("Log (Viral Reads)")
@@
   ) +
-  scale_x_log10(label = scales::label_log10()) +
+  scale_x_log10(label = label_log()) + # scales::label_log10()
   ylab("Log (viral reads/cleaned reads)")
@@
   ) +
-  scale_x_log10(labels = scales::label_log10()) +
+  scale_x_log10(labels = label_log()) +# scales::label_log10()
   ylab("Log (viral reads/cleaned reads/genome length)")
@@
   ) +
-  scale_x_log10(label = scales::label_log10()) +
+  scale_x_log10(label = label_log()) + # scales::label_log10()
   ylab("Log (mean read depth)")
@@
   geom_point() +
-  geom_smooth(aes(group = genus, linetype = genome_structure), se = FALSE) +
+  geom_smooth(aes(group = genus, linetype = genome_structure), method = lm, se = FALSE) +
   facet_grid(target ~ Background.sample, labeller = as_labeller(facet_names)) +

### 2025-07-08 manuscript_figures.Rmd
Swapped Figures S15 and 16

@@
 ### --------------------
-
 #### Figure S15
-
-```{r, echo=FALSE, warning=FALSE, message=FALSE}
-
-#per site genome coverage - Kobu
-
-persite_both$Viral.load <- factor(persite_both$Viral.load, labels = c("Shotgun",
-    "10^{2}", 
-    "10^{3}", 
-    "10^{5}"
-    ))
-
-persite_both$Viral.load <- factor(persite_both$Viral.load, levels = c(
-    "10^{2}", 
-    "10^{3}", 
-    "10^{5}",
-    "Shotgun"
-    ))
-
-persite_both |>
-  filter(virus == "k97_19971") |>
-  filter(mapper == "bowtie2") |>
-  filter(type == "dedup") |>
-  filter(Background != "p6") |>
-  ggplot(aes(
-    x = site,
-    y = (coverage),
-    fill = Background
-  )) +
-  geom_col() +
-  theme_few() +
-  theme(
-    axis.title.x = element_text(size = 12),
-    axis.title.y = element_text(size = 12),
-    legend.title = element_blank(),
-    legend.position = "right",
-    panel.grid.major = element_line(linewidth = 0.1, color = "gray60"),
-    panel.grid.minor = element_line(linewidth = 0.1, color = "lightgray")
-  ) +   
-  facet_wrap(Viral.load ~ Sample_id, labeller = label_parsed) +
-  scale_fill_manual(values = cols2, labels = c("M1", "M2")) +
-  ylab("Coverage") +
-  xlab("") +
-  ggtitle("Kobuvirus") +
-  scale_y_log10(labels = scales::trans_format("log10",
-                                              scales::label_math()))
-
-```
-
-#### *** Final plot ***
-```{r}
-#PDF
-ggsave("figures/manuscript_figure_2025/PDF/Figure_S15.pdf",
-       width=12,
-       height=7)
-
-#PNG
-ggsave("figures/manuscript_figure_2025/Png/Figure_S15.png",
-       width=12,
-       height=7)
-
-```
-
-#### Figure S16
 ```{r, echo=FALSE, warning=FALSE, message=FALSE}
@@
 #PDF
+ggsave("figures/manuscript_figure_2025/PDF/Figure_S15.pdf",
+       width=12,
+       height=7)
+
+#PNG
+ggsave("figures/manuscript_figure_2025/PNG/Figure_S15.png",
+       width=12,
+       height=7)
+```
+
+### --------------------
+
+#### Figure S16
+
+```{r, echo=FALSE, warning=FALSE, message=FALSE}
+
+#per site genome coverage - Kobu
+
+persite_both$Viral.load <- factor(persite_both$Viral.load, labels = c("Shotgun",
+    "10^{2}", 
+    "10^{3}", 
+    "10^{5}"
+    ))
+
+persite_both$Viral.load <- factor(persite_both$Viral.load, levels = c(
+    "10^{2}", 
+    "10^{3}", 
+    "10^{5}",
+    "Shotgun"
+    ))
+
+persite_both |>
+  filter(virus == "k97_19971") |>
+  filter(mapper == "bowtie2") |>
+  filter(type == "dedup") |>
+  filter(Background != "p6") |>
+  ggplot(aes(
+    x = site,
+    y = (coverage),
+    fill = Background
+  )) +
+  geom_col() +
+  theme_few() +
+  theme(
+    axis.title.x = element_text(size = 12),
+    axis.title.y = element_text(size = 12),
+    legend.title = element_blank(),
+    legend.position = "right",
+    panel.grid.major = element_line(linewidth = 0.1, color = "gray60"),
+    panel.grid.minor = element_line(linewidth = 0.1, color = "lightgray")
+  ) +   
+  facet_wrap(Viral.load ~ Sample_id, labeller = label_parsed) +
+  scale_fill_manual(values = cols2, labels = c("M1", "M2")) +
+  ylab("Coverage") +
+  xlab("") +
+  ggtitle("Kobuvirus") +
+  scale_y_log10(labels = scales::trans_format("log10",
+                                              scales::label_math()))
+
+```
+
+#### *** Final plot ***
+```{r}
+#PDF
 ggsave("figures/manuscript_figure_2025/PDF/Figure_S16.pdf",
@@
 #PNG
-ggsave("figures/manuscript_figure_2025/PNG/Figure_S16.png",
+ggsave("figures/manuscript_figure_2025/Png/Figure_S16.png",
        width=12,
        height=7)
-```
+
+```

### 2025-07-08 manuscript_figures.Rmd
Update Figure S15

@@
 
+persite_both$Viral.load <- factor(persite_both$Viral.load, levels = c(
+    "10^{2}", 
+    "10^{3}", 
+    "10^{5}",
+    "Shotgun"
+    ))
+
 #per site genome coverage - Cardio
@@
     x = site,
-    y = coverage,
+    y = (coverage),
     fill = Background
@@
   geom_col() +
-  facet_wrap(Viral.load ~ Sample_id, labeller = label_parsed) +
-  scale_fill_manual(values = cols2, labels = c("M1", "M2")) +
   theme_few() +
@@
   ) +   
+  facet_wrap(Viral.load ~ Sample_id, labeller = label_parsed) +
+  scale_fill_manual(values = cols2, labels = c("M1", "M2")) +
   ylab("Coverage") +
